%-----------------------------------------------------------------------
%;  Copyright (C) 1995-2018
%;  Associated Universities, Inc. Washington DC, USA.
%;
%;  This program is free software; you can redistribute it and/or
%;  modify it under the terms of the GNU General Public License as
%;  published by the Free Software Foundation; either version 2 of
%;  the License, or (at your option) any later version.
%;
%;  This program is distributed in the hope that it will be useful,
%;  but WITHOUT ANY WARRANTY; without even the implied warranty of
%;  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%;  GNU General Public License for more details.
%;
%;  You should have received a copy of the GNU General Public
%;  License along with this program; if not, write to the Free
%;  Software Foundation, Inc., 675 Massachusetts Ave, Cambridge,
%;  MA 02139, USA.
%;
%;  Correspondence concerning AIPS should be addressed as follows:
%;         Internet email: aipsmail@nrao.edu.
%;         Postal address: AIPS Project Office
%;                         National Radio Astronomy Observatory
%;                         520 Edgemont Road
%;                         Charlottesville, VA 22903-2475 USA
%-----------------------------------------------------------------------
\renewcommand{\titlea}{31-December-2018 and earlier}
\renewcommand{\Rheading}{\AIPS\ \cookbook:~\titlea\hfill}
\renewcommand{\Lheading}{\hfill \AIPS\ \cookbook:~\titlea}
\markboth{\Lheading}{\Rheading}

\vglue1pt
\vfill
\cbegint
\begin{center}
%\resizebox{!}{13cm}{\gname{monkey}}
\resizebox{!}{13cm}{\gbb{539,539}{monkey}}
\cbeginb
\label{chap:toc}
\end{center}
\vfill
\eject

\thispagestyle{empty}
\vglue75pt
\centerline{\Large Acknowledgments}
\label{\Myfolio}
\vspace{20pt}

     The \COOKBOOK\ cover design is by Pat Smiley of the NRAO Graphic
Arts Department.  It is based on a design suggested by John Bally of
Bell Labs.

     The image on the title page was converted from the \AIPS\
television-like display to Encapsulated PostScript by the \AIPS\ task
{\tt \tndx{TVCPS}}\@.  It was then included in this \TEX\ \iodx{TEX}
document and plotted on a Hewlett Packard PostScript printer.  It is
the green portion of the digitized image of a Mandrill which has
become a standard in the image-processing field.  PostScript is a
registered trademark of Adobe Systems Incorporated.

     The plot in Chapter~4 was generated with {\tt TVCPS} in a similar
manner.  The plots at the ends of Chapters 6, 7, 8, and 9 were
generated by various \AIPS\ plotting tasks with task {\tt
\tndx{LWPLA}} used to convert the device-independent plot files into
PostScript.  The data displayed in Chapters 6 and 7 were provided by
Bill Cotton for use with the \AIPS\ {\tt VLAC} test suite and by Alan
Bridle for use with the \AIPS\ {\tt DDT} test suite, while the data in
Chapter~8 were provided by Don Wells for use in testing spectral-line
software.  The color plots in Chapter 6 are of data provided by Greg
Taylor and by Eric Greisen, Kristine Spekkens, and Gustaaf van
Moorsel.  The editors thank these people for providing their data for
our use.

     This \iodx{COOKBOOK}\COOKBOOK\ itself is based on an early users
guide written by Alan Bridle.  In 1983, it was typeset and edited by
Eric Greisen using the \TEX\ program, initially developed by Donald
E.~Knuth ({\it The \TEX book}, 1984, Addison-Wesley Publishing
Company, Reading, Massachusetts).  There were two editions of the
\COOKBOOK\ in 1983, one in 1985, and one in 1986.  The 1990 edition
edition was edited by Bill Junor.  For recent editions, Eric Greisen
has resumed his r\^ole as editor, while numerous individuals have
contributed to the text.  In particular, Glen Langston, Andrea Cox,
Lorant Sjouwerman, and Minnie Mao have submitted outline guides to VLA
continuum, spectral-line, high-frequency, and low frequency data
reduction which appear as Appendices A. B, D, and L, respectively.
The output of \TEX\ is now converted to PostScript by {\tt dvips}
(from Radical Eye Software).  The editors are grateful to Knuth for
this program and, especially, for his decision to place it in the
public domain.

{\large
This \Cookbook\ is now available on the Internet via the
\indx{World-Wide Web}.  The current Table of Contents together with a
revision history for the full {\tt 31DEC14} \Cookbook\ is available at
$$ \hbox{{\tt http://www.aips.nrao.edu/cook.html}.}$$
You should review this Web page occasionally to see if chapters
important to you have been altered and, if so, why.  You may use your
favorite Web browser to click on any chapter you wish to receive and
the \indx{PostScript} version will (eventually) appear on your
workstation.  Cross-linked html and pdf versions of the full
\Cookbook\ are available from this web page as well.  The html version
is not shipped with \AIPS, but the pdf version is shipped to all
\AIPS\ installations.}

\vfill\eject
\renewcommand{\titlea}{31-December-2018 (revised 6-September-2018)}
\renewcommand{\Rheading}{\AIPS\ \cookbook:~\titlea\hfill}
\renewcommand{\Lheading}{\hfill \AIPS\ \cookbook:~\titlea}
\markboth{\Lheading}{\Rheading}

\renewcommand{\titleb}{Table of Contents}
\renewcommand{\titlec}{ }
\tableofcontents

\vfill\eject
\renewcommand{\titlec}{List of Figures}
\listoffigures

\vfill\eject
\renewcommand{\titlec}{List of Recipes}
\listoftables   %\tableofrecipes

