%-----------------------------------------------------------------------
%;  Copyright (C) 1995-1998, 2000-2018
%;  Associated Universities, Inc. Washington DC, USA.
%;
%;  This program is free software; you can redistribute it and/or
%;  modify it under the terms of the GNU General Public License as
%;  published by the Free Software Foundation; either version 2 of
%;  the License, or (at your option) any later version.
%;
%;  This program is distributed in the hope that it will be useful,
%;  but WITHOUT ANY WARRANTY; without even the implied warranty of
%;  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%;  GNU General Public License for more details.
%;
%;  You should have received a copy of the GNU General Public
%;  License along with this program; if not, write to the Free
%;  Software Foundation, Inc., 675 Massachusetts Ave, Cambridge,
%;  MA 02139, USA.
%;
%;  Correspondence concerning AIPS should be addressed as follows:
%;         Internet email: aipsmail@nrao.edu.
%;         Postal address: AIPS Project Office
%;                         National Radio Astronomy Observatory
%;                         520 Edgemont Road
%;                         Charlottesville, VA 22903-2475 USA
%-----------------------------------------------------------------------
\setcounter{chapter}{13}
\chapts{RECIPES}{recipes}

\renewcommand{\titlea}{31-December-2013 (revised 15-August-2018)}
\renewcommand{\Rheading}{\AIPS\ \cookbook:~\titlea\hfill}
\renewcommand{\Lheading}{\hfill \AIPS\ \cookbook:~\titlea}
\markboth{\Lheading}{\Rheading}

%\sects{Additional recipe}

\sects{Unused recipes}
% chapter 13 *************************************************
\recipe{Hawaiian banana cream pie}

\bre
\Item {Preheat oven to \dgg{375}.}
\Item {In a bowl, combine 1 cup chopped cashew or macadamia {\bf
   nuts}, 1/2 cup  flaked {\bf coconut}, and 2 tablespoons {\bf brown
   sugar}.}
\Item {Beat 1 {\bf egg white} until stiff; fold into nut mixture.}
\Item {Press mixture evenly into an 8-inch pie plate, building up
   the sides slightly.  Bake for 7 minutes or until crust is lightly
   browned.  Crust will tighten as it cools (use a rack).}
\Item {In a medium-sized saucepan, beat 3 {\bf egg yolks}.  Mix in
   5 tablespoons {\bf cornstarch} and 3/4 cup granulated {\bf sugar}.
   Stir in 1.5 cups {\bf milk}, 1/4 teaspoon {\bf salt}, and 1
   tablespoon unsalted {\bf butter}.}
\Item {Cook mixture slowly over medium heat, stirring constantly,
   for 5 to 7 minutes.  Filling should be bubbling and thick.}
\Item {Remove from heat and stir in 1 teaspoon {\bf vanilla
   extract}.  Transfer this custard to a glass bowl, cover with
   plastic wrap, and refrigerate for 2 hours.}
\Item {Two hours before serving, whip 1/2 cup heavy {\bf whipping
   cream} to stiff peaks and fold into custard.  Peel and slice one
   {\bf banana}, arranging evenly on bottom of crust.  Spoon custard
   filling into crust.  Cover again with plastic wrap and chill for 2
   more hours.}
\Item {Sprinkle 1/2 cup finely chopped cashew or macadamia {\bf
   nuts} evenly over the filling.  Peel, slice and arrange a second
   {\bf banana} in a circular fashion around the outside top of the
   pie, placing a few slices decoratively in the
   center.}
\ere

% chapter  9 *************************************************
\recipe{Banana Bombay salad}

\bre
\Item {Puree 3 {\bf bananas}.}
\Item {Whisk with 1/4 cup {\bf lemon juice}, 1/4 cup {\bf mayonnaise},
     1/4 cup {\bf plain yogurt}, and 1/8 -- 1/4 ounce {\bf taragon}.
     Refrigerate at least 2 hours.}
\Item {Cut 2 pounds cooked {\bf turkey} or {\bf chicken breast} into
     bitesize pieces.}
\Item {Add 1/2 cup {\bf raisins}, 3 {\bf green apples} cut into
     pieces, and 1/2 cup chopped {\bf walnuts}.  Mix.}
\Item {Add banana puree and mix.  Cut 2 {\bf bananas} into thick
     chunks and add.  Serve chilled.}
\item[ ]{\hfill Thanks to Chiquita Bananas.  See {\tt
     http://www.jaetzel.de/tim/chiquit.htm}.}
\ere

\vfill\eject
% chapter  I  *************************************************
\recipe{Banana poundcake}

\bre
\Item {Mix in large bowl until blended:}
\vskip 4pt
\hbox to \hsize{\hskip 75pt\baselineskip 14pt\vbox{\halign{\rt{{$#$}}\quad%
&\lft{#}\cr
1 {1\over3}&cups mashed {\bf bananas} (4 medium)\cr
1          &pkg.~($18{1\over2}$ oz.) {\bf yellow cake mix}\cr
1          &pkg.~($3{3\over4}$ oz.) instant {\bf vanilla pudding mix}\cr
1\over3    &cup {\bf salad oil}\cr
1\over2    &cup {\bf water}\cr
1\over2    &teaspoon {\bf cinnamon}\cr
1\over2    &teaspoon {\bf nutmeg}\cr
4          &{\bf eggs} at room temperature\cr}}\hfil}
\Item {Beat at medium speed for 4 minutes.}
\Item {Turn batter into greased and lightly floured 10-inch tube pan.}
\Item {Bake in \dgg{350} oven for 1 hour or until cake tester inserted
in cake comes out clean.}
\Item {Cool in pan 10 minutes, then turn out onto rack and cool
completely.}
\Item {If desired, dust with confectioners sugar before serving.}
\item[ ]{Thanks to the United Fresh Fruit and Vegetable Association.}
\ere

% chapter E *************************************************
\recipe{Banana Dream Pizza}

\bre
\Item {Preheat oven to \dgg{400}.  In a large bowl, combine 2 1/2 cups
       {\bf all-purpose flour}, 2 tsp {\bf baking powder}, and a pinch
       of {\bf salt}.  Add 4 Tsp softened {\bf sweet cream butter} and
       blend.  Add 3/4 cup warm {\bf milk} and mix well.  If the dough
       is still sticky, add a small amount of flour.}
\Item {Form the dough into a ball.  Knead it on a floured surface
       until it is smooth.  Roll out the dough and place it in an
       oiled, 16-inch pizza pan.  Bake for 15-20 minutes, or until the
       crust is light brown.}
\Item {In a nonmetallic bowl, mash 4 {\bf bananans}.  Add 1 teaspoon
       {\bf lime or lemon juice} and 6 tablespoons {\bf honey}; mix
       well.}
\Item {Slice 2 {\bf bananas} horizontally and place the slices in
       water to cover.  Add 1 teaspoon {\bf lime or lemon juice} to
       prevent discoloration.}
\Item {Spread the banana mixture on the crust.}
\Item {Drain the sliced bananas and blot them with paper towels.
       Place them in a circular pattern on the banana mixture.  Baste
       the banana slices with  3 tablespoons {\bf melted butter}.}
\Item {Bake for 20--30 minutes at \dgg{400} until the crust is golden
       brown.}
\Item {Remove from the oven and top with 1 quart {\bf vanilla ice
       cream} and 1/2 cup chopped {\bf macadamia nuts} while still
       hot.  Serve immediately.}
\ere

\vfill\eject
% chapter E *************************************************
\recipe{Banana July cocktail}

\bre
\Item {Sprinkle 3 sliced {\bf bananas} with 1 tablespoon {\bf lemon
     juice}.}
\Item {Mix with 1 1/4 cans drained and flaked {\bf tuna}, 1/2 {\bf
     onion} chopped, and 2 tablespoons chopped {\bf gherkins} or {\bf
     olives}.}
\Item {Spoon into 7 cocktail shells.}
\Item {Melt 2 tablespoons {\bf butter} in a saucepan.  Add 2
     tablespoon {\bf cake flour} and salt and pepper to taste.}
\Item {Add 1/4 cup {\bf chicken stock} and 1/4 cup dry {\bf white
     wine}.  Simmer for one minute stirring constantly.}
\Item {Add 1/3 cup grated {\bf cheddar cheese} and allow to cool.}
\Item {Add 1/4 cup fresh cream to sauce and pour over banana-tuna
     mixture.}
\Item {Sprinkle with 1 tablespoon grated {\bf cheese} and {\bf
     paprika}.  Decorate with a slice of {\bf gherkin} or {\bf
     olive}.}
\Item {Bake 15--20 minutes at \dgg{350};  serve warm.}
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere

% chapter 7 *************************************************
\recipe{Frozen Push-Ups}

\bre
\Item {Peel 2 {\bf bananas} and slice into blender or
    food processor.}
\Item {Add 1 6-ounce can frozen {\bf orange juice} (thawed), 1/2
    cup instant non-fat {\bf dry milk}, 1/2 cup {\bf water}, and 1 cup
    plain low-fat {\bf yogurt}.}
\Item {Cover and blend until foamy.  Pour into small paper cups
    and freeze.}
\Item {To eat, squeeze bottom of cup.}
\item[ ]{\hfill Thanks to Ruthe Eshleman {\it The American Heart
     Association Cookbook}.}
\ere

% chapter L *************************************************
\recipe{Mexican bananas}

\bre
\Item {Mix together 1 cup {\bf sugar}, 1 teaspoon {\bf cinnamon}, 1/8
     teaspoon {\bf nutmeg}, and 1/8 teaspoon {\bf ginger}.}
\Item {Peel 6 firm {\bf bananas}, cut in half lengthwise, and brush
     with 1/4 cup {\bf lemon juice}.}
\Item {Place a banana half at end of each of 12 {\bf tortillas} and
     sprinkle with  sugar mixture.}
\Item {Roll tortillas, brush top and sides with 1/4 cup {\bf
     evaporated milk}, and then sprinkle with remaining sugar mixture.}
\ere


\vfill\eject
% chapter 3 *************************************************
\recipe{Banana curried chicken}

\bre
\Item {Fry 2 chopped {\bf onions} in 50 ml {\bf cooking oil} until
     light brown.}
\Item {Add 1/4 cup {\bf cake flour} and mix well. Add 1 (cup?) {\bf
     chicken stock} gradually while stirring.}
\Item {Add 1 cup {\bf raisins}, 1 teaspoon {\bf salt}, 2 pounds
     cooked, boned {\bf chicken}, 5 sliced {\bf bananas}, 2 grated
     {\bf apples}, 2 tablespoons grated {\bf lemon rind}, 1 tablespoon
     {\bf sugar}, 1 1/2 tablespoons {\bf curry powder}, 1 {\bf bay
     leaf}, 4 {\bf peppercorns}.}
\Item {Cover saucepan and simmer for 20 minutes.}
\Item {Remove bay leaf.  Add 1 cup {\bf cream} and heat just before
     serving.}
\Item {Serve on a bed of rice. Decorate with pineapples if preferred.}
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere


%                              USED RECIPES
%                              USED RECIPES
%                              USED RECIPES
\sects{Used recipes}
%                              USED RECIPES
%                              USED RECIPES
%                              USED RECIPES

% chapter 2 *************************************************
\recipe{Banana daiquiri}

\bre
\Item {Combine in an electric blender: 2 ounce {\bf light rum}, 0.5
     ounce {\bf banana liqueur}, 0.5 ounce {\bf lime juice}, 1/2 small
     {\bf banana} peeled and coarsely chopped, and 1/2 cup crushed
     {\bf ice}.}
\Item {Blend at high speed until smooth.}
\Item {Pour into large saucer champagne (or similar) glass.
           Serves one.}
\ere

% chapter 2 *************************************************
\recipe{Banana pick-me-up}

\bre
\Item {Slice ripe, peeled {\bf bananas} into 3 cm chunks.}
\Item {Wrap each chunk in strip blanched {\bf bacon}.}
\Item {Prepare mixture of {\bf brown sugar} and {\bf cinnamon} to
         taste.}
\Item {Sprinkle mixture over banana chunks.}
\Item {Bake at \dgg{350} until the bacon is crisp and the sugar
         slightly caramelized.}
\ere


% chapter  3  *************************************************
\recipe{Banana mandarin cheese pie}

\bre
\Item {In large mixer bowl, beat 8 ounces softened {\bf cream cheese}
     until fluffy.}
\Item {Gradually beat in 8 ounces {\bf sweetened condensed milk} until
     smooth.}
\Item {Stir in 1 teaspoon {\bf lemon juice} and 1 teaspoon {\bf
     vanilla extract}.}
\Item {Slice 2 medium {\bf bananas}, dip in lemon juice, and drain.}
\Item {Line 8(?)-inch {\bf graham cracker pie crust} with bananas and
     about 2/3 of an 11-ounce can (drained) {\bf mandarin oranges}.}
\Item {Pour filling over fruit and chill for 3 hours or until set.}
\Item {Garnish top with remaining orange segments and 1 medium {\bf
     banana} sliced and dipped in lemon juice.}
\ere

\vfill\eject

% chapter 4  *************************************************
\recipe{Columbian fresh banana cake with sea foam frosting}

\bre
\Item {Open an 18.5 oz {\bf yellow cake mix} into a large mixxing
   bowl; do not use a mix that contains pudding or requires oil.
   Combine with 1/8 teaspoon {\bf baking soda}.}
\Item {Stir 3/4 cup {\bf Coca-Cola} briskly until foaming stops.
   Add to cake mix and blend until just moistened.  Then beat at high
   speed for 3 minutes, scraping the bowl often.}
\Item {Combine 2 teaspoons {\bf lemon juice} with 1 cup mashed {\bf
   bananas} and then add to batter.  Add 1/3 cup finely chopped {\bf
   nuts} and beat for 1 minute at medium speed.}
\Item {Turn the batter into a well greased, lightly floured 9x13
   baking dish. Bake in a preheated \dgg{350} oven for about 40
   minutes or until the cake tests done.   Cool on a rack for 15
   minutes, remove cake from pan and turn right side up on a rack to
   finish cooling.}
\Item {In the top of a double boiler, combine 2 large {\bf egg
   whites}, 1 1/2 cups packed {\bf light brown sugar}, 1/8 teaspoon
   {\bf cream of tartar} (or 1 tablespoon corn syrup), and 1/3 cup
   {\bf Coca-Cola}.  Beat at high speed for 1 minute with an electric
   mixer.}
\Item {Place over boiling water --- the water should not touch the
   bottom of the top half of the double boiler.  Beat on high speed
   for about 7 minutes until the frosting forms peaks when the mixer
   is raised. Remove from boiling water.  Empty into a large bowl.}
\Item {Add 1 teaspoon {\bf vanilla extract} and continue beating on
   high speed until thick  enough to spread, about 2 minutes. Spread
   on the sides and top of the cold banana cake.}
\ere

% chapter 4 *************************************************
\recipe{Chewy banana split dessert}

\bre
\Item {Prepare and bake one package (19.8 0z) chewy fudge (or
      other favorite) {\bf brownie mix}.  Allow to cool thoroughly,
      four hours or more.}
\Item {Peel 2 large ripe {\bf bananas} and place very thin
      slices on top of brownie.}
\Item {Cover bananas evenly with one 12-oz.~container of {\bf
       whipped topping} (thawed) and drizzle 1/2 cup {\bf chocolate
       syrup} over that.}
\Item {Refrigerate to chill completely.  Cut into squares to
       serve.}
\ere

% chapter 6 *************************************************
\recipe{Banana mallow pie}

\bre
\Item {Combine 2 cups {\bf vanilla wafer} crumbs and 1/3 cup melted
     {\bf butter}.  Press into 9-inch pie plate and bake at \dgg{375}
     for 8 minutes.}
\Item {Prepare a 3 1/8 ounce package {\bf vanilla pie filling} using 1
     3/4 cup {\bf milk}.  Cover surface with transparent wrap and
     chill.}
\Item {Fold 1 1/2 cups {\bf mini-marshmallows} and 1 cup {\bf Cool
     Whip} into pie filling.}
\Item {Slice 2 {\bf bananas} into pie crust, pour filling over
     bananas, and chill several hours or overnight.}
\ere

\vfill\eject
% chapter 6 *************************************************
\recipe{Sopa de Pl\'atano}

\bre
\Item {Cook 10 whole red-skinned under-ripe {\bf
    bananas} in one quart of water over low heat.}
\Item {Peel and mash bananas with 1/4 teaspoon {\bf cloves}, 1/4
    teaspoon {\bf or\'egano}, and 1 teaspoon {\bf powdered cinnamon}.}
\Item {Knead the mixture, add a pinch of {\bf salt}, and fry in 4
    tablespoon {\bf shortening} until slightly browned.}
\Item {Chop 4 medium-sized {\bf tomatoes}, 2 {\bf green peppers},
    and 1 medium-sized {\bf onion}.}
\Item {Fry vegetables in 1/4 cup {\bf olive oil} about 5 minutes
    and then add 1 teaspoon {\bf salt}.}
\Item {Place banana mixture on serving dish and garnish with the
    hot vegetables.}
\item[ ]{\hfill Thanks to Ruth Mulvey and Luisa Alvarez {\it Good
     Food from Mexico}.}
\ere

% chapter 7 *************************************************
\recipe{Bananes r\^ oties}

\bre
\Item {Preheat oven to \dgg{375}.}
\Item {Place 6 (peeled) {\bf bananas} in a baking dish.}
\Item {Sprinkle bananas with juice of 1/2 {\bf lemon}.}
\Item {Pour 2 tablespoons melted {\bf butter} and 2 tablespoons
   {\bf dark rum} over the bananas.  Sprinkle with 2 tablespoons
   {\bf brown sugar}.}
\Item {Place in oven for 10 minutes.}
\Item {Pour on 2 more tablespoons {\bf melted butter} and 2 more
           tablespoons {\bf dark rum} and bake for 5 minutes more.}
\Item {Serve at once, spooning some sauce over each banana.}
\ere

% chapter 7 *************************************************
\recipe{Almond fudge banana cake}

\bre
\Item {Mash 3 extra-ripe {\bf bananas} to make 1 1/2 cups.}
\Item {Beat 1 1/2 cups {\bf sugar}, and 1/2 cup softened {\bf
     margarine} until light and fluffy.  Beat in 3 {\bf eggs}, 3
     tablespoons {\bf amaretto liqueur} (or 1/2---1 teaspoon {\bf
     almond extract}), and 1 teaspoon {\bf vanilla extract}.}
\Item {Combine 1 1/3 cups {\bf all-purpose flour}, 1/3 cup unsweetened
     {\bf cocoa powder}, 1 teaspoon {\bf baking soda}, 1/2 teaspoon
     {\bf salt}, and 1/2 cup toasted {\bf chopped almonds}.}
\Item {Add dry mixture and bananas alternately to beaten mixture.
     Beat well.}
\Item {Turn batter into greased 10-inch bundt pan.  Bake in \dgg{350}
     oven 45 to 50 minutes or until toothpick inserted in center comes
     out nearly clean and cake pulls away from sides of pan.  Cool 10
     minutes.  Remove cake from pan to wire rack to cool completely.}
\Item {Puree 1 small {\bf banana} and beat into 1 ounce (1 square)
     melted {\bf semisweet chocolate}.  Drizzle this glaze over top
     and down sides of cooled cake.}
\ere

\vfill\eject
% chapter 7 *************************************************
\recipe{Coriander banana nut bread}

\bre
\Item {Blend together in a large bowl $1 {2\over3}$ cups sifted
     all-purpose {\bf flour}, 3/4 cup {\bf sugar}, 1 tablespoon {\bf
     baking powder}, 1/2 teaspoon {\bf baking soda}, 1/2 teaspoon {\bf
     salt}, 2 teaspoons ground {\bf coriander}.}
\Item {Mix in $1$ cup chopped unblanched {\bf almonds} and set
     aside.}
\Item {Melt 1/3 cup {\bf shortening} and set aside to cool.}
\Item {Mix until well blended 1 large well-beaten {\bf egg}, 1/4
     cup {\bf buttermilk}, and 1 teaspoon {\bf vanilla extract}.}
\Item {Blend in $1 {1\over4}$ cups mashed ripe {\bf
     bananas} and the shortening.}
\Item {Make a well in center of dry ingredients and add banana
     mixture all at one time.  Stir only enough to moisten dry
     ingredients.}
\Item {Turn into greased $9\times5\times3$-inch loaf pan and
     spread to corners.}
\Item {Bake at \dgg{350} about 1 hour or until a wooden pick comes
     out clean when inserted in center of bread.  Immediately remove
     from pan and set on rack to cool.}
\ere

% chapter 7 *************************************************
\recipe{Easy banana bread}

\bre
\Item {Preheat oven to \dgg{350}.}
\Item {In a food processor cream 1/2 cup soft {\bf tofu}, 3/4 cup {\bf
     honey}, 1/4 cup {\bf sunflower or safflower oil}, 1 teaspoon {\bf
     vanilla extract}, {\bf egg substitute} for 1 egg, and 1 cup
     mashed ripe {\bf banana}.}
\Item {In a bowl combine 2 cups {\bf whole wheat pastry flour}, 1/2
     teaspoon {\bf baking powder}, and 1/2 teaspoon {\bf baking
     soda}.}
\Item {Add to food processor along with a dash {\bf salt} and process
     until creamy.  Pulse in 1 tablespoon {\bf poppy seeds}.}
\Item {Pour into an oiled 9 x 5 x 3-inch loaf pan.  Bake for 30 to 35
     minutes, or until toothpick inserted in center of bread comes out
     clean. Cool on a wire rack for 30 minutes before removing from
     pan.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
    tsculey@bigfoot.com}).}
\ere

% chapter 7 *************************************************
\recipe{Orange gingered bananas}

\bre
\Item {Combine in a small saucepan 1/4 cup {\bf orange juice}
   and 1/2 teaspoon {\bf cornstarch}.  Cook and stir over medium heat
   until boiling.}
\Item {Add 1/4 cup {\bf orange juice}, 1 1/2 teaspoons {\bf
   honey}, and 1 1/2 teaspoons chopped {\bf crystallized ginger} and
   cook, stirring, until thoroughly heated.}
\Item {Place 2 peeled, green-tipped {\bf bananas} in a shallow
   baking dish and cover with sauce.}
\Item {Bake at $350\deg$ about 15 minutes or until the bananas are
   tender (but not soft), basting with the sauce several times.}
\ere

\vfill\eject
% chapter 8 *************************************************
\recipe{Hot banana souffl\'e}

\bre
\Item {Preheat oven to \dgg{375}.}
\vskip 4pt
\Item {Select a 6-cup souffl\'e dish or other mold and grease it
     liberally with 1 tablespoon {\bf butter}.}
\Item {Place 6 {\bf eggs}, 1/2 cup {\bf cream}, juice of 1/2
     {\bf lemon}, 1 tablespoon {\bf kirsch}, and 1/4 cup {\bf sugar}
     in blender. Blend until the batter is smooth.}
\Item {Peel 2 large {\bf bananas}, removing any fibers and
     break into chunks.  With blender running, add the chunks one at a
     time.}
\Item {Break 11 ounces {\bf cream cheese} into chunks and add them
     to the blender.}
\vskip 4pt
\Item {When all the ingredients are thoroughly mixed, run the
     blender at high speed for a few seconds.}
\Item {Pour batter into prepared dish and place it in the hot
     oven.  Bake 45--50 minutes until the top is lightly browned and
     puffy.  You may quit when the center is still a bit soft or
     continue baking until the center is firm.}
\Item {Serve at once.  A whipped cream flavored with Grand
     Marnier makes a nice topping.}
\ere

% chapter 8 *************************************************
\recipe{Curried bananas}

\bre
\Item {Melt 2 tablespoons {\bf butter} in saucepan and cook 2
     tablespoons minced {\bf onion} in it for 2--3 minutes.}
\Item {Mix 1 tablespoon {\bf curry powder}, 1 teaspoon {\bf salt}, 1/4
     cup {\bf flour}, and a dash of {\bf cayenne pepper} with a little
     {\bf milk} to make a paste.}
\Item {Add paste to onion, cooking gently for 10 minutes.  Add balance
     of 2 cups {\bf milk} slowly, stirring until it boils.}
\Item {Slice 7 small green {\bf bananas}, and cook gently in the sauce
     until tender.}
\Item {Serve as a vegetable in a ring of hot cooked rice.}
\item[ ]{\hfill From {\it Everyday BANANA Recipes\/}, Banana
     Distributing Co., New Orleans, published by Bauerlein, Inc. New
     Orleans, 1927.}
\ere

\vfill\eject
% chapter 8 ************************************************
\recipe{Banana crunch cake}

\bre
\Item {Heat oven to \dgg{350}.  Grease and flour 10-inch tube (Bundt)
   pan.}
\Item {In medium bowl, combine 1/2 cup {\bf flour}, 1 cup {\bf
   cocnut}, 1 cup {\bf rolled oats}, 3/4 cup firmly packed {\bf brown
   sugar}, and 1/2 cup chopped {\tt pecans}.  Mix well.}
\Item {Using fork or pastry blender, cut in 1/2 cup {\bf margarine}
   until mixture is crumbly.  Set aside.}
\Item {In a large bowl, combine 1 1/2 cups sliced very ripe {\bf
   bananas}, 1/2 cup {\bf sour cream}, and 4 {\bf eggs}; blend until
   smooth.}
\Item {Add 1 package {\bf yellow cake mix}, Pilsbury Most Supreme is
   recommended.  Beat 2 minutes at high speed.}
\Item {Spread 1/3 of batter in tube pan, sprinkle with 1/3 of cocnut
   mixture.  Repeat layers twice more using remaining batter and
   coconut mixture, ending with coconut mixture.}
\Item {Bake at \dgg{350} for 50 to 60 minutes or until toothpick
   inserted near center comes out clean. Cool upright in pan 15
   minutes; remove from pan. Place on serving plate, coconut side up.
   Cool completely.}
\Item{HIGH ALTITUDE --- above 3500 Feet: Add 3 tablespoons flour
   to dry cake mix. Bake at \dgg{375} for 45 to 55 minutes.}
\ere

% chapter 8 *************************************************
\recipe{Panecillos de Pl\'atano}

\bre
\Item {Sift together 2 cups {\bf flour}, 1 teaspoon {\bf salt},
    and 3 teaspoon {\bf baking powder}.}
\Item {Add 4 tablespoons softened {\bf butter}, mix well, add 3/4
    cup {\bf milk}, and stir only until dampened.}
\Item {Roll to 1/2 inch thickness, cut into cookies about 2 inches
    in diameter, and place on greased cookie sheet.}
\Item {Slice 2 {\bf bananas} in 1/2 inch thicknesses and
    dip pieces in 2 tablespoons {\bf lemon juice} and then in 2
    tablespoons {\bf sugar}.  Place a slice on each cookie, pressing
    it down.}
\Item {Bake in a \dgg{425} oven for 12 minutes or until golden
    brown.}
\item[ ]{\hfill Thanks to Ruth Mulvey and Luisa Alvarez {\it Good
     Food from Mexico}.}
\ere

% chapter 8 *************************************************
\recipe{Banana-Rhubarb Crisp}

\bre
\Item {Slice 2 large {\tt bananas} into 1/4-inch rounds.
     Combine with $2 {1\over2}$ cups diced {\tt rhubarb}, 2 tablespoon
     {\tt sugar}, 1/4 teaspoon {\tt cinnamon}, and a generous dash
     {\tt nutmeg}.  Spoon the mixture into a well-greased 9-inch pie
     plate or shallow baking dish (preferably glass or ceramic).}
\Item {In a medium bowl, combine 1/2 cup white or whole-wheat
     pastry {\tt flour}, 1/2 cup {\tt graham cracker} crumbs, $1
     {1\over2}$ teaspoons {\tt baking powder}.  With a pastry blender
     or two knives worked in a crisscross fashion, cut in 1/4 cup {\tt
     butter} until the mixture is crumbly.}
\Item {Combine 1 {\tt egg} lightly beaten with 1/4 cup {\tt milk}
     and stir into the flour mixture.  Spoon the batter as evenly as
     possible over the fruit mixture.  Sprinkle with 2 tablespoons
     {\tt sugar}.}
\Item {Bake in a pre-heated \dgg{400} oven for 25--30 minutes.}
\item[ ]{\hfill Thanks to {\it Jane Brody's Good Food Book}.}
\ere

\vfill\eject
% chapter 9 *************************************************
\recipe{Banana relish}

\bre
\Item {Cut 12 {\bf bananas}, 1 pound {\bf dates}, and 2 pounds {\bf
     Bermuda onions} into small pieces.}
\Item {Add 2/3 cup {\bf molasses}, 1/2 teaspoon ground {\bf ginger}, 1
     teaspoon {\bf salt}, 1 teaspoon {\bf allspice}, 1 cup {\bf
     water}, and 2 cups {\bf vinegar}; mix well.}
\Item {Turn into a large stone jar or crock, bake in a slow oven till
     rich brown, seal in jars while hot.}
\ere

% chapter  9 *************************************************
\recipe{Cranberry banana bread}

\bre
\Item {In a large saucepan, bring 2 cups {\bf sugar} and 1 cup {\bf
     water} to a boil, stirring to dissolve the sugar.  Add 4 cups
     fresh {\bf cranberries} and simmer over low heat for 10 minutes
     or until berries pop open. Cool. Drain the berries, reserving the
     juice and measuring 1 cup of berries for use in the bread.}
\Item {Sift together 1 3/4 cup {\bf flour}, 1/2 teaspoon {\bf salt},
     2 teaspoon {\bf baking powder} and 1/4 teaspoon {\bf baking
     soda}.}
\Item {In a large bowl, combine 2/3 cup {\bf sugar}, 1/3 cup melted
     {\bf butter}, 2 beaten {\bf eggs}, 1/2 cup chopped {\bf walnuts},
     1 cup mashed {\bf banana}, and 1 cup cooked berries.}
\Item {Add the flour mixture to the berry mixture, stirring until
     blended. Pour the mixture into a greased and lightly floured 9 x
     5 x 3-inch loaf pan. Bake in a preheated, \dgg{350} oven for 1
     hour or until a toothpick inserted in the center comes out
     clean.}
\Item {For a topping (optional), combine 1/4 cup {\bf cranberry juice}
     from cooked berries, 2 tablespoons {\bf sugar} and 2 tablespoons
     {\bf Grand Marnier} in a small saucepan and stir over low heat
     until heated through. Poke a few holes in the baked loaf and pour
     on the topping.}
\Item {Cool 10 minutes in the pan. Turn the loaf out on a rack and
     cool completely. Wrap in foil and store one day before slicing.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
     tsculey@bigfoot.com}).}
\ere

% chapter  9*************************************************
\recipe{Churros de Pl\'atano}

\bre
\Item {Heat about 1 inch of salad (or part salad and part
   olive) {\bf oil} in a large frying pan.}
\Item {Peel and split 3 large, green-tipped {\bf bananas}
   lengthwise.  Then cut each piece in half and dip in {\bf lemon
   juice}.}
\Item {Separate 4 {\bf eggs}.  Beat the egg yolks until thick and
   light.  Then add 1/4 cup {\bf flour} and 1/2 teaspoon {\bf salt}.}
\Item {Beat the egg whites until stiff, but not dry, and fold into
   yolk mixture.}
\Item {Drop the drained banana pieces one at a time into the
   batter.  Pick up with a spoon and slide into the hot oil.}
\Item {Cook over medium heat, turning almost at once, until brown
   on both sides.  Drain on paper towels.}
\ere

\vfill\eject
% chapter 9 *************************************************
\recipe{Roasted turkey quesadillas with banana}

\bre
\Item {Place 6 corn or whole wheat flour {\bf tortillas} flat.}
\Item {Sprinkle with 6 ounces grated low-fat {\bf Jack} or {\bf
     cheddar cheese}, 2 tablespoons chopped fresh {\bf cilantro} or
     {\bf parsley}, 1/2 pound shredded roasted {\bf turkey} or {\bf
     chicken} meat, 2 seeded and minced {\bf jalape\~no peppers}, 1
     cup {\bf alfalfa sprouts}, and 2 medium {\bf bananas}, sliced
     into thin circles.}
\Item {Place 6 {\bf tortillas} on top and press firmly.}
\Item {Place on a lightly oiled cookie sheet; cover with another
      cookie sheet of similar size.  Bake in a pre-heated \dgg{350}
      oven for 15 minutes until soft and melted.  Cut into wedges and
      serve with hot sauce and salad.}
\item[ ]{\hfill Thanks to Chiquita Bananas.  See {\tt
     http://www.jaetzel.de/tim/chiquit.htm}.}
\ere

% chapter 11 *************************************************
\recipe{Banana coffeelate}

\bre
\Item {Peel and mash 2 ripe {\bf bananas}.}
\Item {Blend in 1/2 teaspoon {\bf vanilla extract}, a few grains
   {\bf salt}, 1/4 cup {\bf chocolate syrup}, 2 teaspoons {\bf sugar},
   and 2 teaspoons instant powdered {\bf coffee}.}
\Item {Add $1 {1\over2}$ cups {\bf milk}.}
\Item {Beat with rotary beater or electric mixer until smooth and
  creamy. Chill.}
\ere

% chapter 12 *************************************************
\recipe{Going bananas with bananas}

\bre
\Item {Garnish a baked ham or ham steak with bananas.}
\Item {Make a quick, rich desert with bananas and cream.}
\Item {Bananas are perfect for lunch boxes.  They come in their
   own wrapper, are easy to eat and mess-less.}
\Item {Slice a banana in half lengthwise, brush with melted butter
   and bake it until tender; serve it as a ``vegetable'' with roasted
   meats or fish.  Very Caribbean.}
\Item {Don't forget old favorites like bananas sliced over cereal,
   diced in pancake batter, or buried midst the ice cream in a banana
   split.}
\Item {Slice and stir-fry bananas with carrots, tomatoes and
  ground beef for a super-quick main dish.}
\ere

\vfill\eject
% chapter 12 *************************************************
\recipe{Banana-chocolate tea bread}

\bre
\Item {Cream 1/2 cup softened {\bf butter}, gradually add 1 cup {\bf
     sugar}, beating until light and fluffy.  Add 2 {\bf eggs}, one at
     a time, beating well after each addition.}
\Item {Combine 1 1/2 cups all-purpose {\bf flour}, 2 tablespoons {\bf
     cocoa}, 1 teaspoon {\bf baking soda}, 1 teaspoon {\bf salt}, and
     1/2 teaspoon {\tt cinnamon}; sift together.}
\Item {Stir flour mixture into egg mixture, blending well.}
\Item {Add 1 teaspoon {\bf vanilla extract}; stir in  1 cup mashed
     {\bf banana}, 1/2 cup {\bf sour cream}, 1/2 cup chopped {\bf
     walnuts}, and 1/3 cup miniature {\bf semi-sweet chocolate
     chips}.}
\Item {Spoon batter into two greased and floured 7-1/2 x 3 x 2-inch
     loaf pans. Bake at \dgg{350} for 55 minutes or until a wooden
     pick inserted in center comes out clean. Cool in pans 10 minutes,
     remove from pans and cool completely on a wire rack.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
     tsculey@bigfoot.com}).}
\ere

% chapter 12 ************************************************
\recipe{Banana caramel pie}

\bre
\Item {Mix 1 cup {\bf dark brown sugar}, 1/4 cup {\bf all-purpose
   flour}, 1/4 teaspoon {\bf salt} in a saucepan.  Stir in 1/4 cup
   {\bf cold water} and 2 {\bf egg yolks}.  Beat until smooth.}
\Item {Gradually stir in 1 cup {\bf boiling water}.  Then cook,
   stirring constantly, about 3 minutes until smoothly thickened.}
\Item {Stir in 1 tablespoon {\bf butter}, 1/2 teaspoon {\bf vanilla},
   and 1/4 cup {\bf evaporated milk}.  Cool slightly.}
\Item {Pour into pre-baked, cooled 8 or 9 inch {\bf pastry pie
   shell}.}
\Item {Slightly before serving, slice 4 ripe {\bf bananas} and arrange
   over filling.  Top with whipped cream or with a meringue made out
   of the 2 left-over egg whites.}
\ere

% chapter 12 ************************************************
\recipe{Curried shrimp}

\bre
\Item {Cook 2 1/2 pounds {\bf shrimp} for 3 minutes.  Peel and
     devein.}
\Item {Heat 1/3 pound {\bf butter} or margarine in large saucepan.
     Saute 4 chopped {\bf scallions} and 2 cups chopped, peeled
     {\bf apples} until tender.  Stir in 2 tablespoons {\bf curry
     poweder}, 1 tablespoon {\bf ground ginger}, and 1/3 cup {\bf
     flour}.  Stir for 2 minutes.  Remove from heat and blend in 3
     cups {\bf chicken broth}.  Return to heat, cook stirring until
     mixture boils and thickens.}
\Item {Add 1 pound roasted {\bf cashews}, 1 pound {\bf Turkish
     apricots}, and, if desired, 2 ounces diced {\bf crystalized
     ginger} and {\bf raisins}.  Cook over low heat for 15 minutes.}
\Item {Add shrimp and mix in.}
\Item {Cut 3 {\bf bananas} into thick slices and add to mixture.
     Serve over cooked white or curried rice.}
\item[ ]{\hfill Thanks to Chiquita Bananas.  See {\tt
     http://www.jaetzel.de/tim/chiquit.htm}.}
\ere

\vfill\eject
% chapter 13 *************************************************
\recipe{Saut\'ed sole tobago with bananas, pecans and lime}

\bre
\Item {Preheat 1/2 cup {\bf vegetable oil} in a heavy sauce pan over
     medium-high heat.}
\Item {Dredge 8 filets of {\bf sole} or {\bf flounder} lightly in {\bf
     flour}.}
\Item {Saut\'ee until golden brown, about 3 minutes each side.  Remove
     to warm platter.}
\Item {Pour off excess oil and wipe down sauce pan. Place pan back on
     stove over high head; add 1/4 cup {\bf butter}.}
\Item {When foamy and just starting to brown, add 2 cups diagonally
     sliced {\bf bananas} (1/2'' slices) and 1 cup {\bf pecan} halves.
     Toss and cook for 1 minute.}
\Item {Add 1/2 cup  fresh {\bf lime juice} and 1 cup dry white {\bf
     wine} (or light stock) . Cook for another 2 minutes.}
\Item {Add 1/4 cup {\bf fresh herbs} (mint, parsley, coriander, basil
     or tarragon).}
\Item {Pour sauce and bananas over fish. Garnish with additional
     banana slices and lime wedges. }
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere

% chapter 13 *************************************************
\recipe{Banana sweet potato puff casserole}

\bre
\Item {In a large bowl, combine 2 cups mashed {\bf sweet potatoes}, 1
     cup mashed ripe {\bf bananas} (3 medium), 3/4 teaspoon {\bf curry
     powder}, 1/3 cup {\bf sour cream}, 1/2 teaspoon {\bf salt}, and 1
     {\bf egg}.}
\Item {Beat with electric mixer until light and very fluffy. Turn into
     1 quart casserole dish.}
\Item {Bake at \dgg{350} for 20 minutes or until puffed and lightly
     browned.}
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere

% chapter 13 *************************************************
\recipe{Cream of banana soup}

\bre
\Item {Cook 1 quart green {\bf banana} pulp, 1 1/2 quarts {\bf chicken
     stock}, 1 small {\bf celery stalk}, 1/2 {\bf onion}, 1 {\bf
     carrot}, 1 small {\bf bay leaf}, 5 {\bf peppercorns}, and {\bf
     salt} to taste together for about 30 minutes until the mixture
     thickens.}
\Item {Strain over 1/4 cup {\bf flour} and 1/4 cup {\bf butter} which
     have been combined as for a white sauce.  Cook until thickened.}
\Item {Just before serving, add 2 cups {\bf cream} or {\bf milk} and
     heat.}
\Item {Serve with a slice of lemon on each plate as a garnish.}
\ere


\vfill\eject
% chapter A *************************************************
\recipe{Banana colada}

\bre
\Item  {Peel and slice 1 ripe {\bf  banana}.}
\Item {Place sliced banana in blender along with 6 ounces {\bf
     pineapple juice} (or crushed tinned pineapple in its own juice)
     and 1 ounce {\bf rum} plus 1 ounce {\bf coconut rum} {\it or\/} 2
     ounce {\bf rum} plus 1 teaspoon {\bf Coco Lopez}.}
\Item {Optionally add 1 ounce {\bf banana liqueur}.}
\Item {Blend until smooth.}
\Item {Add crushed ice, if so desired.}
\Item {If the mixture is too thick, add more juice (or more rum if
     you prefer!); if too thin, add more banana. This is a really easy
     recipe to adjust to one's taste.}
\ere

% chapter A *************************************************
\recipe{Breaded chicken and bananas}

\bre
\Item {In food processor, blend 1 can {\bf condensed milk}, 1/3 cup
     {\bf milk}, 1/2 cup flaked {\bf coconut}, and 1/4 cup {\bf lemon
     juice} until smooth. Pour into a bowl.}
\Item {Prepare 3 cups {\bf corn flake crumbs} in another bowl or plate.}
\Item {Cut  6 very firm {\bf bananas} lengthwise, dip in milk mixture,
     roll in corn flakes, and set aside.}
\Item {Cut 2 {\bf chickens} into pieces, dip in milk mixture, roll in
     corn flakes, and place in greased baking pans (2 13x9 pans may be
     required).}
\Item {Sprinkle chicken with 1/2 cup melted {\bf butter} and bake as
     \dgg{350} for one hour.}
\Item { Arrange bananas over the chicken. Sprinkle with 1/4 cup melted
     {\bf butter}. Bake 15 minutes longer or until chicken juices run
     clear.}
\Item {Garnish with sliced star and/or kiwi fruits if desired.}
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere

% chapter A *************************************************
\recipe{Banana-pineapple bread}

\bre
\Item {Mix together 1 cup chopped {\bf nuts}, 2-1/2 cups {\bf sugar},
     5 cups {\bf flour}, 1 teaspoon {\bf salt}, 1 teaspoon {\bf baking
     powder}, and 1 teaspoon {\bf cinnamon}.}
\Item {Mix together 1-1/2 cups {\bf vegetable oil}, 3 {\bf eggs}, 3
     mashed {\bf bananas}, 1 teaspoon {\bf lemon juice}, and 1 can
     {\bf crushed pineapple} (drained).}
\Item {Combine.  Bake at \dgg{350} for one hour.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
     tsculey@bigfoot.com}).}
\ere

\vfill\eject

% chapter B *************************************************
\recipe{Banana breeze pie}

\bre
\Item {In a small saucepan, melt 1/3 cup {\bf butter} or {\bf
     margarine}.  Add 1/4 cup {\bf  sugar} and 1/2 teaspoon {\bf
     cinnamon}.  Stir constantly over low heat until bubbles form
     around the edges of pan.}
\Item{Remove from heat, add 1 cup {\bf cornflake cereal} crumbs and
     mix well.  Press mixture evenly into a 9-inch pie pan to form
     crust. Chill.}
\Item {Beat 8 ounces softened {\bf cream cheese} until light and
     fluffy.  Add 1 15-ounce can {\bf condensed milk} and blend
     thoroughly.  Add 1/3 cup {\bf lemon juice} and 1 teaspoon {\bf
     vanilla}.  Stir until thickened.}
\Item {Slice 3 ripe {\bf bananas} and line crust. Pour filling into
     crust and refrigerate for 2--3 hours or until firm.  Do not
     freeze.}
\Item {Slice 2 ripe {\bf  bananas}, dip in lemon juice and arrange on
     top of pie.  Note. for a change of pace, use lime juice.}
\ere

% chapter B *************************************************
\recipe{Banana cutlets}

\bre
\Item {Peel 6 medium-ripe {\bf bananas} and halve them crosswise.}
\Item {Dip them in 1/3 cup {\bf lemon juice} and then roll in 1 cup
     crushed {\bf cornflake crumbs}.}
\Item {Saute them in 3 tablespoons {\bf butter} until a golden brown.}
\Item {Serve on lettuce.}
\ere

% chapter C *************************************************
\recipe{Golden mousse}

\bre
\Item {Combine 1 cup mashed ripe {\bf bananas}, 2
     tablespoons {\bf orange juice}, 1/4 cup shredded {\bf coconut}, 3
     tablespoons {\bf brown sugar}, a few grains {\bf salt}, and 1/8
     teaspoon grated {\bf orange rind}.}
\Item {Whip until stiff 1 cup {\bf heavy cream}.}
\Item {Fold whipped cream into fruit mixture and turn into
     freezing tray.  Freeze rapidly without stirring until firm.}
\ere

% chapter C *************************************************
\recipe{Mexican chicken vegetable soup with bananas}

\bre
\Item {In large, covered kettle, over medium-low heat, simmer 4 pounds
     cut up {\bf stewing chicken}, 1/c cup coarsely  chopped {\bf
     onion}, 1 teaspoon {\bf salt}, and 4 cups of hot {\bf water} for
     2 hours or until chicken is tender.}
\Item {Remove chicken to cutting board; cut meat from bones into
     chunks; discard bones. Skim any fat from surface of broth.}
\Item {Add chicken, 1/2 cup chopped {\bf celery}, 1 12-ounce can
     whole-kernel {\bf corn} and 1 16-ounce can {\bf tomatoes} to
     soup. Continue simmering, covered for 10 minutes. Season to
     taste.}
\Item {Five minutes before serving, peel 4 firm (green-tipped) {\bf
     bananas}, slice diagonally into 1-inch slices. }
\Item {Add sliced bananas to soup, continue cooking just until bananas
     are tender.  Serve immediately.}
\item[ ]{\hfill Thanks to Turbana Corporation ({\tt www.turbana.com}).}
\ere
\vfill\eject

% chapter C *************************************************
\recipe{Orange baked bananas}

\bre
\Item {Mix in a saucepan 1/2 cup firmly packed {\bf brown sugar},
         1 tablespoon {\bf cornstarch}, 1/8 teaspoon {\bf cinnamon},
         and a few grains {\bf salt}.}
\Item {Add gradually, blending in 3/4 cup boiling water.}
\Item {Bring rapidly to boiling and cook about 5 minutes or until
         sauce is thickened, stirring constantly.}
\Item {Remove from heat and blend in $1 {1\over2}$ teaspoons grated
         {\bf orange peel}, 1/4 cup {\bf orange juice}, 1 teaspoon
         {\bf lemon juice}, and 2 tablespoons {\bf butter}.}
\Item {Peel and cut into halves lengthwise 6 {\bf bananas} with
         all-yellow or green-tipped peel.}
\Item {Arrange halves cut side down in baking dish and brush with
         about 2 tablespoons melted {\bf butter}.}
\Item {Sprinkle 1/2 teaspoon {\bf salt} over bananas and then pour
         the orange sauce over bananas.}
\Item {Bake at \dgg{375} for 10 to 20 minutes.}
\ere

% chapter C *************************************************
\recipe{Delightful banana cheesecake}

\bre
\Item {Preheat oven to \dgg{350}.}
\Item {Combine 1.5 cups crushed {\bf cereal} (3 cups un-crushed
     Multi-Bran Chex suggested), 1/3 cup melted {\bf margarine} or
     butter, and 1/4 cup packed {\bf brown sugar}; mix well.}
\Item {Press firmly onto bottom and sides of greased 9-inch pie
     plate.  Bake 8--10 minutes, then cool completely.}
\Item {Arrange 1.5 cups sliced {\bf bananas} onto sides and
     bottom of cooled crust.}
\Item {Combine 16 oz.~softened light or regular {\bf cream
     cheese}, 1.5 cups {\bf powdered sugar}, and 3/4 teaspoon {\bf
     vanilla extract}.}
\Item {Mix well, then fold in 2 cups light or regular {\bf whipped
     topping}.  Pour over sliced bananas.}
\Item {Cover and refrigerate for 4 hours or until set.}
\Item {Garnish with 1/2 cup sliced {\bf bananas}.}
\item[ ]{\hfill Thanks to Ralston Purina Company.}
\ere

% chapter C *************************************************
\recipe{Chicken salad with banana mayonnaise and grapes}

\bre
\Item {Place 3 medium {\bf bananas} cut in chunks, 2 teaspoons chopped
     {\bf garlic}, 3/4 cup non-fat {\bf plain yogurt}, 1 tablespoon
     {\bf honey}, 2 teaspoons {\bf lemon juice}, and 1/4 teaspoon {\bf
     salt} in a blender or food processor.  Blend until creamy.}
\Item {Arrange 12 cups mixed {\bf lettuces} on six plates.}
\Item {Toss 6 {\bf chicken breasts} cooked and cubed with banana mayo;
     divide onto salads.}
\Item {Sprinkle with 2 bunchs ($\approx 48$) halved {\bf grapes} and
     1/2 cup {\bf walnut} or {\bf pecan} halves.}
\item[ ]{\hfill Thanks to Chiquita Bananas.  See {\tt
     http://www.jaetzel.de/tim/chiquit.htm}.}
\ere

\vfill\eject
% chapter D ***********************************************
\recipe{Dulce Zacateca\~no}

\bre
\Item {Peel 3 large not-too-ripe {\bf bananas} and slice
    lengthwise.  Saute in 5 tablespoons {\bf butter} until golden
    brown.  Drain on paper, place in a shallow baking dish, and
    sprinkle with a little {\bf sugar}.}
\Item {Whip 1/2 cup {\bf heavy sweet cream}.  Add 1/4 cup {\bf
    sugar}, 1/4 cup {\bf dry sherry wine}, and 1 teaspoon {\bf
    vanilla}.  Pour over bananas covering them completely.  Chill and
    serve very cold.}
\item[ ]{\hfill Thanks to Ruth Mulvey and Luisa Alvarez {\it Good
     Food from Mexico}.}
\ere

% chapter D **********************************************
\recipe{Virginia's instant banana pie}

\bre
\Item {Mix 1 cup {\bf sour cream}, 1 cup {\bf milk}, and 1 small
     package {\bf instant vanilla pudding} until mixture thickens.}
\Item {Slice 3 medium {\bf bananas} into the bottom of a 9-inch {\bf
     graham cracker pie crust}.}
\Item {Pour the pudding over the bananas and refrigerate at least 2 hours.}
\ere

% chapter D *************************************************
\recipe{Chocolate chip banana bread}

\bre
\Item {Blend 2 cups mashed {\bf bananas}, 1 tablespoon grated {\bf
     orange peel}, and 1/3 cup {\bf orange juice} in a bowl.  Beat in
     3 {\bf eggs}.  Stir in 1 cup packed {\bf brown sugar} and 1/3 cup
     {\bf vegetable oil}.}
\Item {Combine 2-1/2 cups {\bf all-purpose flour}, 1 cup {\bf
     chocolate chips} 2 teaspoons {\bf baking powder}, 1/2 teaspoon
     {\bf baking soda}, 1/2 teaspoon {\bf salt}, and 1/2 teaspoon {\bf
     nutmeg}.}
\Item {Stir dry ingredients into banana mixture just until blended.
     Pour into 4 greased 5-3/4 x 3-1/4-inch loaf pans.}
\Item {Bake in \dgg{350} oven for 45 to 55 minutes or until tester
     inserted comes out clean.  Let cool in pans on rack for 10
     minutes.  Remove from pan and let cool completely on rack.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
     tsculey@bigfoot.com}).}
\ere

% chapter D *************************************************
\recipe{Banana-pineapple rum bread}

\bre
\Item {Place 1/2 cup {\bf white rum} and 1/2 cup diced {\bf dried
     pineapple} in a bowl, cover, and let sit for at least one hour.}
\Item {In a mixing bowl, beat together 4 tablespoon {\bf butter} or
     margarine and 3/4 cup {\bf sugar}.  Add 1 extra large {\bf egg}
     and continue beating until light and fluffy.}
\Item {Add 2 large mashed ripe {\bf bananas} and mix well.  Beat in
     1/3 cup plain {\bf yogurt} --- curdling of the mixture is
     normal.}
\Item {In another mixing bowl, combine 2 cups {\bf all-purpose flour},
     1/2 tablespoon {\bf baking soda}, 1 teaspoon ground {\bf
     cinnamon}, 1 teaspoon ground {\bf nutmeg}, 1 teaspoon ground {\bf
     allspice}, and 1/2 teaspoon {\bf salt}.}
\Item {Add the wet ingredients and mix until well blended.  Drain the
     pineapple and add.  Fold in 1/2 cup coarsely chopped {\bf
     pecans}.}
\Item {Pour into liberally greased 9-inch loaf pan.  Bake at \dgg{350}
     for 45 to 55 minutes or until the bread passes the toothpick
     test.  Remove the pan from the oven and let it sit for 10
     minutes, before turning out on a rack to cool.}
\item[ ]{\hfill Thanks to Tim D. Culey, Baton Rouge, La. ({\tt
     tsculey@bigfoot.com}).}
\ere

\vfill\eject
% chapter D *************************************************
\recipe{Banana bran muffins}

\bre
\Item {Preheat oven to \dgg{400}.}
\Item {Grease 12 2.75-inch muffin cups.}
\Item {In bowl, combine 1/2 cup crushed {\bf cereal} (1.5 cups
   un-crushed Multi-Bran Chex recommended), 1.5 cups all-purpose {\bf
   flour}, 1/2 cup {\bf sugar}, 1/3 cup chopped {\bf nuts} (optional),
   2.5 teaspoons {\bf baking powder}, and 1/2 teaspoon {\bf baking
   soda}.}
\Item {In a separate bowl, combine 3 large mashed {\bf
   bananas} (1.5 cups), 1 {\bf egg} slightly beaten, 1/4 cup
   vegetable {\bf oil}, 2 tablespoons {\bf water}, and 1 teaspoon {\bf
   vanilla extract}.}
\Item {Add to cereal mixture and stir just until moistened.  Do
   not over-mix.}
\Item {Divide evenly among muffin cups.}
\Item {Bake 18--20 minutes, or until tester inserted in center
   comes out clean.}
\item[ ]{\hfill Thanks to Ralston Purina Company.}
\ere

%\vfill\eject

%\vfill\eject
% chapter F *************************************************
\recipe{Banana stuffing}

\bre
\Item {Pare and rub 4 {\bf bananas} through a sieve into bowl.}
\Item {Add 1/2 grated {\bf onion}, 1 {\bf green pepper} chopped fine,
     3 tablespoons finely  chopped {\bf parsley}, 4 slices cooked {\bf
     bacon} chopped fine, 1 1/4 cups {\bf bread crumbs}, pinch of {\bf
     thyme}, 1 teaspoon {\bf salt}, and 1 {\bf egg}.}
\Item {Mix thoroughly, fill 1 {\bf chicken}, and roast in the usual
     manner.}
\ere

% chapter F *************************************************
\recipe{Banana nut bread}

\bre
\Item {Cream 1 cup {\bf sugar} and 1/2 cup {\bf margarine}
      together.}
\Item {Add 2 {\bf eggs}, 2 cups {\bf flour}, 1/2 teaspoon {\bf
      salt}, and 1 teaspoon {\bf baking soda} and mix thoroughly.}
\Item {Add 1 cup chopped {\bf nuts} (walnuts or pecans), 3/4 cup
      mashed {\bf bananas}, and, lastly, 4 teaspoons {\bf sour
      milk} and mix well.}
\Item {Put in greased loaf pan.}
\Item {Bake in \dgg{350} oven for 1 hour.}
\ere

% chapter  L *************************************************
\recipe{Banana storage}

      Bananas ripen after harvesting.  They do it best at room
temperature.  Because of this there are three stages to banana
storage.
\par\bre
\Item {{\bf On the counter:} When you buy a bunch of
   bananas that are not exactly at the ripeness you want, you
   can keep them at room temperature until they are just right for
   you.  Be sure to keep them out of any plastic bags or containers.}
\Item {{\bf In the refrigerator:} If there are any bananas left,
   and they are at the ripeness you like, you can put them in the
   refrigerator.  The peel will get dusty brown and speckled, but the
   fruit inside will stay clear and fresh and at that stage of
   ripeness for 3 to 6 days.}
\Item {{\bf In the freezer:} If you want to keep your bananas even
   longer, you can freeze them.  Mash the bananas with a little lemon
   juice, put them in an air tight freezer container and freeze.  Once
   they're defrosted, you'll go bananas baking bread, muffins and a
   world of other banana yummies.  Or, you can freeze a whole banana
   on a Popsicle stick.  When it is frozen, dip it in chocolate sauce,
   maybe even roll it in nuts, then wrap it in aluminum foil and put
   it back in the freezer.  Talk about a scrumptious snack.}
\ere

%\vfill\eject
%\hphantom{.}
