%-----------------------------------------------------------------------
%;  Copyright (C) 1995
%;  Associated Universities, Inc. Washington DC, USA.
%;
%;  This program is free software; you can redistribute it and/or
%;  modify it under the terms of the GNU General Public License as
%;  published by the Free Software Foundation; either version 2 of
%;  the License, or (at your option) any later version.
%;
%;  This program is distributed in the hope that it will be useful,
%;  but WITHOUT ANY WARRANTY; without even the implied warranty of
%;  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%;  GNU General Public License for more details.
%;
%;  You should have received a copy of the GNU General Public
%;  License along with this program; if not, write to the Free
%;  Software Foundation, Inc., 675 Massachusetts Ave, Cambridge,
%;  MA 02139, USA.
%;
%;  Correspondence concerning AIPS should be addressed as follows:
%;          Internet email: aipsmail@nrao.edu.
%;          Postal address: AIPS Project Office
%;                          National Radio Astronomy Observatory
%;                          520 Edgemont Road
%;                          Charlottesville, VA 22903-2475 USA
%-----------------------------------------------------------------------
\def\SECRET{F}% internal consumption only?
\def\SHORT{F}%  short form for non-CPs?
\def\FINAL{F}%  draft or final? (final date:)
\if T\FINAL     \day=99\month=13\year=2001 \fi

\def\INP{\uppercase{aipssite}.ADR}%

% Macros for the AIPS Site Survey Directory.
% DCW, NRAO-CV, Mar-Apr86, 15Jul86.
% two-column format added by EWG, 15Jul86.
% Change "am" to "cm" (ComputerModern), DCW 07Jan87.
% minor \hfil fix, DCW 27Apr88.

\hsize=6.5in
%\vsize=9.1in	%\vsize=8.41in = old number from AA
\newdimen\paght	\paght=9.6in \vsize=\paght

\nopagenumbers

\font\SS=cmssi10      % sans-serif 10pt for value fields

\parskip=2.5pt
\def\V{\vskip6.0pt}
\def\up#1{\leavevmode \raise.16ex\hbox{#1}}
\font\giantbf=cmbx10 scaled\magstep3
\if T\FINAL{\global\tolerance=500\global\hfuzz=40pt\global\overfullrule=0pt}
	\else{}\fi

\def\BOXIT#1{\lower3pt\vbox{\hrule\hbox{\vrule\hskip3pt
	\vbox{\vskip3pt\hbox{#1}\vskip3pt}\hskip3pt\vrule}\hrule}}
\def\MONTH{\ \ifnum\month=1 January \else\ifnum\month=2 February
	\else\ifnum\month=3 March \else\ifnum\month=4 April
	\else\ifnum\month=5 May \else\ifnum\month=6 June
	\else\ifnum\month=7 July \else\ifnum\month=8 August
	\else\ifnum\month=9 September \else\ifnum\month=10 October
	\else\ifnum\month=11 November \else\ifnum\month=12 December
	\else Month?? \fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi}
\def\DATE{\if T\FINAL \the\day\MONTH\the\year
	\else{\it\the\year\MONTH\the\day:\the\time\ \up[Draft, from
	\hbox{\tt \INP}\up]}\fi}

\def\headtitlestring{AIPS Directory Listing}
\def\titlea{{\tenbf \headtitlestring}\if F\FINAL{ \tenrm\DATE}\fi}

\def\begindirectory{\begingroup	\let\CPfirstaddr=T
	\parindent=0pt	\parskip=0pt	\raggedbottom
	\everypar={\parindent=0pt\hangindent=30pt\hangafter=1}
	}
\def\enddirectory{\endgroup}

\def\CP#1{\vskip8pt plus5pt minus3pt\vbox{#1\vskip8pt\hrule}}
\def\CPaddr#1{\let\CPfirstaddr=T\let\CPfirstmach=T\noindent{#1}.\hfil}
\def\CPaddrline#1{\if F\CPfirstaddr{, }\else\global\let\CPfirstaddr=F\fi{#1}}
\def\CPtelephone#1{;\hfil\break\hbox to 9.3mm{}Telephone:~{\SS #1\/}}
\def\CPtelex#1{;\hbox to 20mm{}Telex:~{\SS #1\/}}
\def\CPnetaddr#1{;\hfil\break\hbox to 6.8mm{}Net address:~{\SS #1\/}}
\def\CPattribute#1{\if T\SECRET{; Attributes=(#1)}\fi}
\def\CPsitedata#1#2{\hfil\break\hbox to 6.8mm{\vbox to 15pt{}\hfil}Site-wide
                                                         values:~{\SS #2\/}}
\def\CPlastname#1{}
\def\CPcountry#1{}
\def\CPzip#1{}
\def\CPcity#1{}
\def\CPstreet#1{}
\def\CPinstitute#1{}
\def\N#1{}
\def\CPmachine#1#2{\if T\SHORT{
	\if T\CPfirstmach{\hfill\global\let\CPfirstmach=F}\fi\ \BOXIT{\bf #1}}
	\else{\raggedright\pretolerance=10000\adjdemerits=0
	\hyphenpenalty=10000\tolerance=10000\hbadness=10000
	\vskip4pt\noindent\hskip5pt\BOXIT{\bf #1}\ {\SS #2}\hfil\vskip0.1pt}
	\fi}

\def\begincrossref{\begingroup \let\XRfirst=T
	\parindent=20pt	\parskip=0pt	\raggedbottom
	\everypar={\hangindent=2\parindent \hangafter=1}
        \begintwocol
	\centerline{\bf Keyword-Value Cross-Reference Table}
	\vskip0.1in
	}
\def\endcrossref{\endtwocol\endgroup}
\def\XRkey#1#2{\vskip10pt plus10pt
{\noindent{\bf #1}\filbreak{#2}}}
\def\XRvalue#1#2{\vskip2pt\indent{\SS #1\/}\if F\SHORT{:\allowbreak\hskip
1em plus 1em minus 1 em}\fi
	\allowbreak\let\XRfirst=T{#2}\allowbreak\filbreak}
\def\XRmachine#1{\if F\SHORT\if F\XRfirst{, }\else\let\XRfirst=F\fi {#1}\fi}
%\def\XRnmach#1{\if T\SHORT{\leaders\hbox{ . }\hfill #1\hskip20pt minus20pt}
%	\else{\leaders\hbox{ . }\hfill \up[#1\up]\hskip20pt minus20pt}\fi}
\def\XRnmach#1{\if T\SHORT{\leaders\hbox{ . }\hfill #1}
	\else{\leaders\hbox{ . }\hfill \up[#1\up]}\fi}

% ===========================================================================
%             Eric AA macros

\newdimen\parthsize
\newdimen\colht
\def\colrst{\global\colht=\paght \global\advance\colht by -0.7in
	\global\vsize=\colht}
\colrst
\setbox5=\vbox{}

\nopagenumbers

\vsize=\paght
\def\twocol{F}
\def\firstcol{T}
\output{\if T\twocol{\if T\firstcol{\gdef\firstcol{F}\global\setbox1=\pagebody
                      \message{[Column 1 out]}}
                 \else{\gdef\firstcol{T}\global\setbox2=\hbox to 6.5in{\box1
                          \hfil\vrule\hfil\pagebody}}\fi}
    \else{\global\setbox2=\pagebody}\fi
%
    \if T\firstcol{\shipout\vbox{\vskip-.465in\moveleft1.37in\vbox to 10.3944in
  {\hbox{\vrule width 0pt\hbox to 8.079in{\vbox{
  \ifodd\count0{\moveright 1.480in\vbox{ %  to 9.8927 in
    \baselineskip0pt\lineskip0pt
    \ifnum\pageno>1{\vskip 0.32in\vbox to 0.15 in{\vfill
          \hbox to 6.5 in{\it \titlea\hss Page \folio}}
        \vskip 0.08in\hrule height 0.8pt width 6.5in\vskip 0.29 in}\fi
    \makeheadline\box2\makefootline
    \vfill}}
  \else{\moveright 1.192in\vbox { %  to 9.8927 in
    \baselineskip0pt\lineskip0pt
    \ifnum\pageno>1{\vskip 0.32in\vbox to 0.15 in{\vfill
          \hbox to 6.5 in{\it Page \folio\hss \titlea}}
        \vskip 0.08in\hrule height 0.8pt width 6.5in\vskip 0.29 in}\fi
    \makeheadline\box2\makefootline
    \vfill}}\fi
    }\hfil}\vrule width 0pt}\vfill}}
    \advancepageno
    \ifnum\outputpenalty>-20000 \else\dosupereject\fi\colrst}
    \else{\global\vsize=\colht}\fi}

\def\begintwocol{\vfill\eject
     \global\def\twocol{T}\global\tolerance=1600\global\hsize=3.12in}
\def\endtwocol{\if T\twocol\if T\firstcol \vfill\eject
     \hrule height 0pt width\hsize\fi\fi\vfill\eject
     \global\def\twocol{F}\global\hsize=6.5in\global\tolerance=200}

% TeX source for cover pages of the AIPS Site Directory.
% DCW, 29 April 1988 (for 1987 survey)
% NDW,  4 May 1989 (updated for 1988 and 1989 surveys)
% ==================================================================
\day=28\month=3\year=1991 % If set, drafts won't pick up current date!
\def\SECRET{F}
\def\FINAL{T} %commented out after 1988 survey; don't think it's needed
% but here's where to look if there are problems! -ndw, 15-May-89
%
\def\gboxit#1#2#3{\vbox{\hrule height#1\hbox{\vrule width#1\kern#2%
\vbox{\kern#2{#3}\kern#2}\kern#2\vrule width#1}\hrule height#1}}
\font\amfont=cmssi10 %at 14.4pt
\def\memnuma{69}
\def\memnumb{\phantom{9}}
\def\memnumc{\phantom{9}}
\centerline{\gboxit{2pt}{7.5pt}{\hbox{\amfont AIPS\quad MEMO\quad
        NO.\quad\memnuma\memnumb\memnumc\quad}}}
%
\vbox{\vskip0.7in}
\centerline{\giantbf 1991 AIPS Site Directory}
\def\headtitlestring{1991 AIPS Site Directory}
\V\V\V
\centerline{\bf Alan Bridle and Joanne Nance}
\V
\centerline{\bf National Radio Astronomy Observatory\footnote{$^{1}$}
{\rm operated by Associated Universities, Inc.,
under co-operative agreement with the U.~S.~National Science Foundation.}}
\centerline{\bf Edgemont Road, Charlottesville, Virginia 22903-2475, USA}
\vskip 0.1truein
\centerline{Tel +1-804-296-0211; TWX 910-997-0174}
\centerline{aipsmail@nrao.edu\hskip0.1in aipsmail@nrao.bitnet}
\centerline{6654::aipsmail\hskip0.1in ...!uunet!nrao.edu!aipsmail}
\V
\centerline{\DATE}
\vskip0.3in

\noindent{{\bf Summary:}}\hskip0.5in

We tabulate the most recent AIPS Site Survey data for 139 of the 142
active AIPS sites (``Active" sites are those whose last Site Survey
response reported that they have AIPS installed, have measurable AIPS
CPU use and have a non-zero number of AIPS users).  This listing of
``who has what and where'' will help AIPS Site Managers to find other
AIPS sites with hardware, use patterns and scientific interests similar
to their own.

\V\V

\noindent{{\bf About the Directory:}}\hskip0.5in

Forms for the 1990/91 AIPS Site Survey were mailed in November 1990 to
the designated Contact People at all 238 sites that have received AIPS
installation tapes and which we thought might be actively running AIPS.
Responses came back over the next few months, and we stopped entering
data early in March 1991.  This Directory lists the basic data for 204
machines or clusters of machines at 139 of the 142 ``active" AIPS sites.
(The managers of 3 sites asked us not to publish their data.) This
Directory's main listing and cross-reference tables should help AIPS
Site Managers to locate other sites with hardware and user environments
similar to their own.

The first part of this Directory lists the postal and network addresses
and telephone numbers for the Contact People, and, for each one, all of
the active AIPS machines or machine clusters that they support.  Detailed
configuration parameters are then shown for each machine or cluster.

The main listing is ordered first by country, and then by postal (ZIP)
code within each country.  Where postal codes are not available, the
listings are in order of the city name.  The NRAO sites are sorted as a
country called ``NRAO'', so they appear after the NETHERLANDS and before
SPAIN.

The second part of this Directory contains cross-reference tables for
some of the site and machine parameters, using the machine (or cluster)
sequence numbers shown in boxes in the main list.  These tables let you
quickly find all the machines with given types of hardware, given types
of AIPS use, etc.  The tables for Contact People (CP Lastname) and Site
Names provide alphabetical indices to the people and places that now
support active AIPS installations.  The DISK Space and MEMory Size
tables are included so that AIPS Managers can quickly see where their
site fits on the spectrum of these resources.

Recent trends in AIPS site hardware will be summarized and interpreted
in AIPS Memo No.  70.

We thank all the AIPS Contact People who took the time to fill in the
1990/91 Survey forms.  The previous survey results have been used
extensively by AIPS ``consumers" and Site Managers for planning purposes and
to provide ``contacts" outside the NRAO for help with hardware problems
relating to AIPS.  We hope that this new edition of the Site Directory
will allow this to continue.

\vfill\eject \noindent{{\bf Notes on the Cross-Reference Tables:}}
\vskip0.1in plus 0.1in

\par\noindent
Keyword cross-references appear in alphabetic order, and for each
keyword the values are sorted in some rational order.  The definitions
of each keyword are given below:

\noindent{{\tt AP}} is the ``array processor model''.  Ordinary scalar
floating point (FP) hardware is not considered to be an AP --- our aim
is to list only pipelined FP hardware here.  An exception is that some
machines have either a VLSI FP-coprocessor chip (usually a
Motorola~68881) or high-performance FP ``accelerator''.  The code
``FPA'' is used for these.

\noindent{{\tt AVUSERS}} is the average number of simultaneous AIPS
users on the machine or cluster.

\noindent{{\tt BAND}} classifies the site's use of AIPS for applications
other than radio astronomy.  The BAND cross-reference table may be
particularly useful if you wish to find other AIPS sites with similar
scientific interests to your own.

\noindent{{\tt Country}} is generally the country in which the site is
located, but this parameter also distinguishes ``NRAO'' from other U.S.
sites.

\noindent{{\tt CPU}} is the type of computer on which the active AIPS is
installed.

\noindent{{\tt DISK}} is the disk capacity for AIPS, in units of
{\it Gigabytes} (1000~Megabytes).  For clusters, this is summed across
the cluster.

\noindent{{\tt FRACT}} is the AIPS CPU use on the machine or cluster, as
a percentage of total CPU use on the system.

\noindent{{\tt MEM}} is the maximum memory size available for AIPS, in
units of {\it Megabytes}.  For clusters, this is the largest memory
available in the cluster, not the total.

\noindent{{\tt NONRADIO}} measures the non-radio AIPS use at the site,
as a percentage of the overall AIPS use.

\noindent{{\tt Net-Service}} information is tabulated so that AIPS sites
with service on only one or two of the major networks can discover which
other AIPS sites they can reach.  Network connectivity among the active
AIPS sites is now almost total: ``service-on-at-least-one-net'' is
available to 129 of the 139 sites in this Directory.

\noindent{{\tt OS}} is the computer's operating system and version if known.

\noindent{{\tt PLOT}} lists the plotting device(s).  A wide range of
plotters is attached to AIPS machines! Because the NRAO cannot develop
and support interface code for all of these models, this is an area
where cooperation among AIPS sites can be critical.  AIPS users and Site
Managers whose plotters are unsupported by the NRAO can inspect the PLOT
cross-reference table to find other sites which may be able to help
them.

\noindent{{\tt PROGS}} is the number of AIPS programmers at the
site.

\noindent{{\tt SITE}} is a terse descriptive string used to identify each
site in other AIPS documents and as an alternate index field for this one.

\noindent{{\tt TAPE}} describes the tape format(s) supported.

\noindent{{\tt TASKS}} is the number of `locally-coded' AIPS
tasks.

\noindent{{\tt TV}} is the TV display model (and in some cases,
windowing mode).  Site Managers with TV displays that do not yet
function fully may find the cross-reference listing helpful if they need
to consult people who have similar equipment elsewhere.

\noindent{{\tt USERS}} is the annual number of AIPS users at the site.

\noindent{{\tt VLA}} is the AIPS use for VLA processing on the machine
or cluster, expressed as a percent of the total AIPS use.

\vfill\eject



\begindirectory

\CP{
\CPaddr{
\CPaddrline{Dr. David F. Crawford}%
\CPaddrline{Univ. of Sydney}%
\CPaddrline{School of Physics}%
\CPaddrline{Sydney, NSW, 2006}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Crawford, Dr. David F.}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{Sydney, NSW, 2006}%
\CPinstitute{School of Physics}%
\CPstreet{Univ. of Sydney}%
\CPnetaddr{ crawford@suphys.physics.su.oz.au}%
\CPattribute{CP}%
}%
\CPsitedata{1}{%
USERS=10 PROGS=1 TASKS=5 BAND=IR,optical
}%
\CPmachine{1}{%
SITE=U.~Sydney
CPU=Sun~Clus.~(4/490~server;4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=0.2 MEM=16
PLOT=Apple~Laserwriter+
TV=Sun-4~(Sunview)
TAPE=800bpi,1600bpi,6250bpi,QIC24,Exabyte
AVUSERS=4 AIPS=15JUL90
FRACT=10 VLA=5 NONRADIO=5
}%
\CPmachine{2}{%
SITE=U.~Sydney
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3-1
DISK=0.1 MEM=4
PLOT=Apple~Laserwriter+
TV=AED~512
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL88
FRACT=10 VLA=0 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. A. Vaughan}%
\CPaddrline{School of Math \&\ Physics}%
\CPaddrline{Computing \&\ Electronics}%
\CPaddrline{MacQuarie Univ. NSW, 2109}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Vaughan, Dr. A.}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{MacQuarie Univ. NSW, 2109}%
\CPinstitute{School of Math \&\ Physics}%
\CPstreet{Computing \&\ Electronics}%
\CPtelephone{ +61-2-805-8943}%
\CPtelex{ 61-02-805-8983 UNIMACQ}%
\CPnetaddr{ alanv@macastro.mqes.mq.oz.au}%
\CPattribute{CP}%
}%
\CPsitedata{3}{%
USERS=1 PROGS=0 TASKS=0
}%
\CPmachine{3}{%
SITE=MacQuarie~U.
CPU=Sun-4/65
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=24
PLOT=?
TV=Sun-4~color
TAPE=Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=90 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Mark R. Calabretta}%
\CPaddrline{Australia Telescope}%
\CPaddrline{P.O. Box 76}%
\CPaddrline{Epping, NSW, 2121}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Calabretta, Dr. Mark R.}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{Epping, NSW, 2121}%
\CPinstitute{Australia Telescope}%
\CPstreet{P.O. Box 76}%
\CPtelephone{ +61-2-868-0412}%
\CPnetaddr{ aipsmngr@rpepping.oz.au}%
\CPnetaddr{ nssdca::psi\%rpepping::aipsmngr}%
\CPnetaddr{ (psi) 505228621001}%
\CPattribute{CP}%
}%
\CPsitedata{4}{%
USERS=60 PROGS=2 TASKS=11 BAND=IR,optical,biomed
}%
\CPmachine{4}{%
SITE=Australia~Telescope
CPU=Convex~C-220
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=4.4 MEM=128
PLOT=NEC~Silentwriter,Dataproducts~Laserwriter,Mitsubishi~Color
TV=?
TAPE=1600bpi,6250bpi,QIC150,Exabyte
AVUSERS=1.1 AIPS=15JUL90
FRACT=21 VLA=30 NONRADIO=10
}%
\CPmachine{5}{%
SITE=Australia~Telescope
CPU=Sun~Clus.~(3*4/60;4/60GX)
AP=none
OS=UNIX~Sun~4.1
DISK=? MEM=32
PLOT=NEC~Silentwriter,Dataproducts~Laserwriter,Mitsubishi~Color
TV=Sun-4
TAPE=Exabyte
AVUSERS=4 AIPS=15JUL90
FRACT=100 VLA=20 NONRADIO=0
}%
\CPmachine{6}{%
SITE=Australia~Telescope
CPU=Sun-4/65
AP=none
OS=UNIX~Sun~4.0.3c
DISK=2.4 MEM=16
PLOT=NEC~Silentwriter
TV=Sun-4
TAPE=1600bpi,6250bpi,QIC150,Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=95 VLA=0 NONRADIO=0
}%
\CPmachine{7}{%
SITE=Australia~Telescope
CPU=Sun-4/65GX
AP=none
OS=UNIX~Sun~x.x
DISK=2.4 MEM=16
PLOT=NEC~Silentwriter
TV=Sun-4
TAPE=1600bpi,6250bpi,QIC150,Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=95 VLA=0 NONRADIO=0
}%
\CPmachine{8}{%
SITE=Australia~Telescope
CPU=Sun-4/370
AP=none
OS=UNIX~Sun~4.1
DISK=? MEM=32
PLOT=NEC~Silentwriter,Dataproducts~Laserwriter,Mitsubishi~Color
TV=Sun-4
TAPE=QIC150
AVUSERS=1 AIPS=15JUL90
FRACT=50 VLA=0 NONRADIO=50
}%
\CPmachine{9}{%
SITE=Australia~Telescope
CPU=Sun-4/60GX
AP=none
OS=UNIX~Sun~4.0.3c
DISK=0.65 MEM=16
PLOT=NEC~Silentwriter
TV=Sun-4
TAPE=QIC150,Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=95 VLA=0 NONRADIO=10
}%
\CPmachine{10}{%
SITE=Australia~Telescope
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.0
DISK=4.4 MEM=8
PLOT=NEC~Silentwriter,Dataproducts~Laserwriter,Mitsubishi~Color
TV=Sun-3
TAPE=1600bpi,6250bpi,QIC150,Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=100 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bruce C. Cogan}%
\CPaddrline{Mount Stromlo Obs.}%
\CPaddrline{Private Bag}%
\CPaddrline{Weston Creek, ACT, 2611}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Cogan, Bruce C.}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{Weston Creek, ACT, 2611}%
\CPinstitute{Mount Stromlo Obs.}%
\CPstreet{Private Bag}%
\CPtelephone{ +61-06-249-0232}%
\CPtelex{ AA 62270}%
\CPnetaddr{ cogan\&mso.anu.edu.au}%
\CPnetaddr{ nssdca::psi\%mssso::cogan}%
\CPattribute{CP}%
}%
\CPsitedata{11}{%
USERS=5 PROGS=1 TASKS=4
}%
\CPmachine{11}{%
SITE=Mt.~Stromlo~Obs.
CPU=Sun~Clus.~(4/390~server;4/100;4/370;4/65;4*4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=2.3 MEM=32
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=800bpi,1600bpi,6250bpi,QIC24,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=1 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Marco Costa}%
\CPaddrline{Univ. of Western Australia}%
\CPaddrline{Physics Dept.}%
\CPaddrline{Nedlands, WA, 6009}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Costa, Marco}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{Nedlands, WA, 6009}%
\CPinstitute{Physics Dept.}%
\CPstreet{Univ. of Western Australia}%
\CPtelephone{ +61-09-380-2749}%
\CPnetaddr{ marco@earwax.pd.uwa.oz}%
\CPattribute{CP}%
}%
\CPsitedata{12}{%
USERS=1 PROGS=1 TASKS=0
}%
\CPmachine{12}{%
SITE=U.~W.~Aust.~Gravity~Lab
CPU=Sun-4/75
AP=none
OS=UNIX~Sun~4.1.1
DISK=1 MEM=16
PLOT=?
TV=Sun-4
TAPE=Exabyte
AVUSERS=1 AIPS=15OCT90
FRACT=20 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. P.A. (Pip) Hamilton}%
\CPaddrline{Univ. of Tasmania}%
\CPaddrline{Physics Dept., GPO Box 252C}%
\CPaddrline{Hobart, Tasmania, 7001}%
\CPaddrline{AUSTRALIA}%
\CPlastname{Hamilton, Dr. P.A. (Pip)}%
\CPcountry{AUSTRALIA}%
\CPzip{}%
\CPcity{Hobart, Tasmania, 7001}%
\CPinstitute{Physics Dept., GPO Box 252C}%
\CPstreet{Univ. of Tasmania}%
\CPtelephone{ +61-02-202401}%
\CPtelex{ +61-02-202410 (FAX)}%
\CPnetaddr{ aipsmanager@physvax.phys.utas.edu.au}%
\CPattribute{CP}%
}%
\CPsitedata{13}{%
USERS=3 PROGS=0 TASKS=1
}%
\CPmachine{13}{%
SITE=U.~Tasmania~Phys.
CPU=DEC~Clus.~(3500;2*750;2*2000)
AP=none
OS=VMS~5.2
DISK=0.5 MEM=25
PLOT=Apple~Laserwriter~II~NT
TV=DEC~(X-Windows)
TAPE=1600bpi
AVUSERS=1 AIPS=15JUL90
FRACT=3 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Heinz Andernach}%
\CPaddrline{INPE-DAS}%
\CPaddrline{C. P. 515}%
\CPaddrline{12201 Sao Jose dos Campos, S.P.}%
\CPaddrline{BRAZIL}%
\CPlastname{Andernach, Dr. Heinz}%
\CPcountry{BRAZIL}%
\CPzip{}%
\CPcity{12201 Sao Jose dos Campos, S.P.}%
\CPinstitute{(no institute?)}%
\CPstreet{INPE-DAS, C. P. 515}%
\CPtelephone{ +55-123-418977, x309}%
\CPtelex{ 1233530 INPE BR}%
\CPnetaddr{ heinz@dem.inpe.ansp.br}%
\CPnetaddr{ 47599::heinz}%
\CPattribute{CP}%
}%
\CPsitedata{14}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{14}{%
SITE=INPE-DAS~Brazil
CPU=DEC~VAX780
AP=FPS~5040
OS=VMS~4.6
DISK=0.2 MEM=6
PLOT=HP~7475B,Canon~LN03
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL88
FRACT=2 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{David O. Muchmore}%
\CPaddrline{Universite de Montreal}%
\CPaddrline{Dept. de Physique}%
\CPaddrline{C.P. 6128, Montreal PQ, H3C 3J7}%
\CPaddrline{CANADA}%
\CPlastname{Muchmore, David O.}%
\CPcountry{CANADA}%
\CPzip{PQ, H3C 3J7}%
\CPcity{C.P. 6128, Montreal}%
\CPinstitute{Dept. de Physique}%
\CPstreet{Universite de Montreal}%
\CPtelephone{ 514-343-6667}%
\CPnetaddr{ 176@umtlvr.bitnet}%
\CPnetaddr{ 176@cc.umontreal.ca}%
\CPattribute{CP}%
}%
\CPsitedata{15}{%
USERS=4 PROGS=0 TASKS=1 BAND=optical
}%
\CPmachine{15}{%
SITE=U.~Montreal
CPU=Sun~Clus.~(4/380S;4*4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=30
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=0.1 AIPS=15JUL90
FRACT=5 VLA=75 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Judith Irwin}%
\CPaddrline{Herzberg Inst. of Astrophysics}%
\CPaddrline{Nat. Research Council of Canada, 100 Sussex Dr.}%
\CPaddrline{Ottawa, ON, K1A 0R6}%
\CPaddrline{CANADA}%
\CPlastname{Irwin, Judith}%
\CPcountry{CANADA}%
\CPzip{K1A 0R6}%
\CPcity{Ottawa, ON,}%
\CPinstitute{Herzberg Inst. of Astrophysics}%
\CPstreet{Nat. Research Council of Canada, 100 Sussex Dr.}%
\CPtelephone{ 613-991-5839,613-993-6060}%
\CPnetaddr{ irwin@hiaras.nrc.ca}%
\CPattribute{CP}%
}%
\CPsitedata{16}{%
USERS=1 PROGS=1 TASKS=3 BAND=sub-mm
}%
\CPmachine{16}{%
SITE=HIA~Ottawa
CPU=Sun-3/180
AP=none
OS=UNIX~Sun~4.0.2
DISK=0.5 MEM=8
PLOT=Apple~Laserwriter
TV=Sun-3
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=98 NONRADIO=2
}%
}%

\CP{
\CPaddr{
\CPaddrline{V. A. Hughes}%
\CPaddrline{Queen's Univ.}%
\CPaddrline{Astronomy Group, Dept. of Physics}%
\CPaddrline{Kingston, ON, K7L 3N6}%
\CPaddrline{CANADA}%
\CPlastname{Hughes, V. A.}%
\CPcountry{CANADA}%
\CPzip{K7L 3N6}%
\CPcity{Kingston, ON,}%
\CPinstitute{Astronomy Group, Dept. of Physics}%
\CPstreet{Queen's Univ.}%
\CPtelephone{ 613-545-2713}%
\CPnetaddr{ hughesv@qucdn.bitnet}%
\CPnetaddr{ hughesv@qucdn.queensu.ca}%
\CPattribute{CP}%
}%
\CPsitedata{17}{%
USERS=2 PROGS=0.5 TASKS=0
}%
\CPmachine{17}{%
SITE=Queen's~U.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~x.x
DISK=0.9 MEM=16
PLOT=Apple~Laserwriter~II
TV=Sun-4
TAPE=QIC24
AVUSERS=1.5 AIPS=15OCT89
FRACT=100 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Susan Tarr}%
\CPaddrline{Univ. of Toronto}%
\CPaddrline{Dept. of Astronomy, 60 St. George St.}%
\CPaddrline{Toronto, ON, M5S 1A7}%
\CPaddrline{CANADA}%
\CPlastname{Tarr, Susan}%
\CPcountry{CANADA}%
\CPzip{M5S 1A7}%
\CPcity{Toronto, ON,}%
\CPinstitute{Dept. of Astronomy, 60 St. George St.}%
\CPstreet{Univ. of Toronto}%
\CPtelephone{ 416-978-3146}%
\CPnetaddr{ susan@radio.astro.utoronto.ca}%
\CPattribute{CP}%
}%
\CPsitedata{18}{%
USERS=12 PROGS=1.5 TASKS=5 BAND=IR,optical
}%
\CPmachine{18}{%
SITE=U.~Toronto
CPU=Sun-3/180
AP=none
OS=UNIX~Sun~x.x
DISK=2.3 MEM=20
PLOT=Imagen~Innovator
TV=Sun-3~color
TAPE=?
AVUSERS=? AIPS=?
FRACT=80 VLA=98 NONRADIO=2
}%
}%

\CP{
\CPaddr{
\CPaddrline{Michel Fich}%
\CPaddrline{Univ. of Waterloo}%
\CPaddrline{Physics Dept.}%
\CPaddrline{Waterloo, ON, N2L 3G1}%
\CPaddrline{CANADA}%
\CPlastname{Fich, Michel}%
\CPcountry{CANADA}%
\CPzip{N2L 3G1}%
\CPcity{Waterloo, ON,}%
\CPinstitute{Physics Dept.}%
\CPstreet{Univ. of Waterloo}%
\CPtelephone{ 519-885-1211,ext. 2725}%
\CPnetaddr{ fich@astro.waterloo.edu}%
\CPattribute{CP}%
}%
\CPsitedata{19}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{19}{%
SITE=U.~Waterloo~Phys.
CPU=SG~Iris~4D/20
AP=none
OS=UNIX~Irix~System~V~3.3.2
DISK=0.4 MEM=12
PLOT=QMS~PS-810
TV=SGI~4D/20
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Sun Kwok}%
\CPaddrline{Univ. of Calgary}%
\CPaddrline{Dept. of Physics, 2500 Univ. Dr.}%
\CPaddrline{Calgary, AB, T2N 1N4}%
\CPaddrline{CANADA}%
\CPlastname{Kwok, Sun}%
\CPcountry{CANADA}%
\CPzip{T2N 1N4}%
\CPcity{Calgary, AB,}%
\CPinstitute{Dept. of Physics, 2500 Univ. Dr.}%
\CPstreet{Univ. of Calgary}%
\CPtelephone{ 403-220-5414}%
\CPtelex{ 03821545}%
\CPnetaddr{ skwok@uncamult.bitnet}%
\CPnetaddr{ kwok@iras.ucalgary.ca}%
\CPattribute{CP}%
}%
\CPsitedata{20}{%
USERS=10 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{20}{%
SITE=U.~Calgary
CPU=Sun-4/260
AP=none
OS=UNIX~Sun~4.0.1
DISK=0.8 MEM=8
PLOT=QMS~800
TV=IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=2 AIPS=15APR90
FRACT=70 VLA=60 NONRADIO=40
}%
\CPmachine{21}{%
SITE=U.~Calgary
CPU=DG~4000~Server
AP=none
OS=UNIX~DG~4.3
DISK=0.8
PLOT=HP~LaserJet~III
TV=DG~System~Monitor~(Hitachi)
TAPE=?
AVUSERS=2 AIPS=15APR90
FRACT=60 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Anthony G. Willis}%
\CPaddrline{Dominion Radio Astrophysical Obs.}%
\CPaddrline{P.O. Box 248}%
\CPaddrline{Penticton, BC, V2A 6K3}%
\CPaddrline{CANADA}%
\CPlastname{Willis, Anthony G.}%
\CPcountry{CANADA}%
\CPzip{V2A 6K3}%
\CPcity{Penticton, BC,}%
\CPinstitute{Dominion Radio Astrophysical Obs.}%
\CPstreet{P.O. Box 248}%
\CPtelephone{ 604-497-5321}%
\CPtelex{ 048-88127}%
\CPnetaddr{ twillis@drao.nrc.ca}%
\CPattribute{CP}%
}%
\CPsitedata{22}{%
USERS=7 PROGS=0 TASKS=0 BAND=IR,optical
}%
\CPmachine{22}{%
SITE=DRAO~Penticton
CPU=DEC~VAX780
AP=FPS~120B
OS=VMS~4.7
DISK=0.5 MEM=16
PLOT=Versatec~1200,QMS~800
TV=IIS~75
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=6 VLA=85 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Phil Gregory}%
\CPaddrline{Univ. of British Columbia}%
\CPaddrline{Physics Dept., 6224 Agricultural Road}%
\CPaddrline{Vancouver, BC, V6T 1W5}%
\CPaddrline{CANADA}%
\CPlastname{Gregory, Phil}%
\CPcountry{CANADA}%
\CPzip{V6T 1W5}%
\CPcity{Vancouver, BC,}%
\CPinstitute{Physics Dept., 6224 Agricultural Road}%
\CPstreet{Univ. of British Columbia}%
\CPtelephone{ 604-228-6417}%
\CPnetaddr{ aspg@ubcmtsg.bitnet}%
\CPnetaddr{ p-greg@mtsg.ubc.ca}%
\CPattribute{CP}%
}%
\CPsitedata{23}{%
USERS=2 PROGS=1 TASKS=0
}%
\CPmachine{23}{%
SITE=UBC~Phys.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.2 MEM=16
PLOT=HP~7550
TV=Sun-4~color
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=25 VLA=50 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Gerry Grieve}%
\CPaddrline{Univ. of British Columbia}%
\CPaddrline{2219 Main Mall, Dept. of Geophys. \&\ Astron.}%
\CPaddrline{Vancouver, BC, V6T 1W5}%
\CPaddrline{CANADA}%
\CPlastname{Grieve, Gerry}%
\CPcountry{CANADA}%
\CPzip{V6T 1W5}%
\CPcity{Vancouver, BC,}%
\CPinstitute{2219 Main Mall, Dept. of Geophys. \&\ Astron.}%
\CPstreet{Univ. of British Columbia}%
\CPtelephone{ 604-228-4320}%
\CPnetaddr{ cbt@ubcmtsg.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{24}{%
USERS=1 PROGS=0 TASKS=3
}%
\CPmachine{24}{%
SITE=UBC~Astron.
CPU=DEC~VAX750
AP=FPS~5200
OS=VMS~x.x
DISK=0.4 MEM=6
PLOT=QMS~800
TV=IIS~70
TAPE=?
AVUSERS=? AIPS=?
FRACT=30 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Qiu Yuhai}%
\CPaddrline{Beijing Astronomical Observatory}%
\CPaddrline{Chinese Academy of Sciences}%
\CPaddrline{Beijing, 100080}%
\CPaddrline{CHINA}%
\CPlastname{Yuhai, Qiu}%
\CPcountry{CHINA}%
\CPzip{}%
\CPcity{Beijing, 100080}%
\CPinstitute{Beijing Astronomical Observatory}%
\CPstreet{Chinese Academy of Sciences}%
\CPtelephone{ +86-1-256-7182}%
\CPtelex{ 22040 BAOAS CN}%
\CPattribute{CP}%
}%
\CPsitedata{25}{%
USERS=4 PROGS=0.5 TASKS=3
}%
\CPmachine{25}{%
SITE=Beijing~Obs.
CPU=Sun~Clus.~(3*4/65)
AP=none
OS=UNIX~Sun~4.1
DISK=1.3 MEM=12
PLOT=Apple~Laserwriter~II
TV=Sun-4~color
TAPE=QIC150
AVUSERS=3 AIPS=15OCT90
FRACT=80 VLA=0 NONRADIO=0
}%
\CPmachine{26}{%
SITE=Beijing~Obs.
CPU=DEC~uVAX~II
AP=none
OS=VMS~4.2
DISK=0.6 MEM=4
PLOT=LBP-8II
TV=Sigma~ARGS~xxxx
TAPE=800bpi,1600bpi
AVUSERS=1 AIPS=15APR89
FRACT=90 VLA=35 NONRADIO=1
}%
}%

\CP{
\CPaddr{
\CPaddrline{AIPS Manager}%
\CPaddrline{Dept. of Geophysics}%
\CPaddrline{Beijing Univ.}%
\CPaddrline{Beijing, 100871}%
\CPaddrline{CHINA}%
\CPlastname{Manager, AIPS}%
\CPcountry{CHINA}%
\CPzip{}%
\CPcity{Beijing, 100871}%
\CPinstitute{Beijing Univ.}%
\CPstreet{Dept. of Geophysics}%
\CPtelephone{ +86-1-282741}%
\CPattribute{CP}%
}%
\CPsitedata{27}{%
USERS=10 PROGS=0 TASKS=0
}%
\CPmachine{27}{%
SITE=Beijing~U.~Geophys.
CPU=DEC~VAX780
AP=none
OS=VMS~x.x
DISK=0.18 MEM=8
PLOT=CANPB~II
TV=AGS
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=10 NONRADIO=0
}%
\CPmachine{28}{%
SITE=Beijing~U.~Geophys.
CPU=DEC~uVAX~II
AP=none
OS=VMS~x.x
DISK=0.36 MEM=8
PLOT=CANPB~II
TV=AGS
TAPE=?
AVUSERS=? AIPS=?
FRACT=90 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Zheng Xingwu}%
\CPaddrline{Nanjing University}%
\CPaddrline{Astronomy Dept.}%
\CPaddrline{Nanjing, 210008}%
\CPaddrline{CHINA}%
\CPlastname{Xingwu, Zheng}%
\CPcountry{CHINA}%
\CPzip{}%
\CPcity{Nanjing, 210008}%
\CPinstitute{Astronomy Dept.}%
\CPstreet{Nanjing University}%
\CPtelephone{ +86-25-637551-2882}%
\CPattribute{CP}%
}%
\CPsitedata{29}{%
USERS=10 PROGS=1 TASKS=16 BAND=optical
}%
\CPmachine{29}{%
SITE=Nanjing~U.~CSTDAC
CPU=Sun~Clus.~(4/280;3/60)
AP=none
OS=UNIX~Sun~4.0
DISK=8 MEM=16
PLOT=Omnilaser~2108
TV=Sun
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15APR90
FRACT=10 VLA=0 NONRADIO=50
}%
\CPmachine{30}{%
SITE=Nanjing~U.~CSTDAC
CPU=DEC~VAX8530
AP=none
OS=VMS~4.7
DISK=1.2 MEM=32
PLOT=KS~LP28XT
TV=none
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15APR90
FRACT=10 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{AIPS Manager}%
\CPaddrline{Bldg 348, Electromagnetics Inst.}%
\CPaddrline{Tech. Univ. of Denmark}%
\CPaddrline{DK 2800 Lyngby}%
\CPaddrline{DENMARK}%
\CPlastname{Manager, AIPS}%
\CPcountry{DENMARK}%
\CPzip{}%
\CPcity{DK 2800 Lyngby}%
\CPinstitute{Tech. Univ. of Denmark}%
\CPstreet{Bldg 348, Electromagnetics Inst.}%
\CPtelephone{ +45-2-881444}%
\CPnetaddr{ a601lpau@hasara11.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{31}{%
USERS=6 PROGS=2 TASKS=4 BAND=IR,optical
}%
\CPmachine{31}{%
SITE=Tech.~U.~Denmark
CPU=DEC~uVAX~II/GPX
AP=none
OS=VMS~x.x
DISK=0.227 MEM=12
PLOT=?
TV=Imaging~Technology~V200
TAPE=?
AVUSERS=? AIPS=?
FRACT=25 VLA=20 NONRADIO=80
}%
}%

\CP{
\CPaddr{
\CPaddrline{A. Bosma}%
\CPaddrline{Obs. de Marseille}%
\CPaddrline{2 Pl. Le Verrier}%
\CPaddrline{F-13248 Marseille Cedex}%
\CPaddrline{FRANCE}%
\CPlastname{Bosma, A.}%
\CPcountry{FRANCE}%
\CPzip{}%
\CPcity{F-13248 Marseille Cedex}%
\CPinstitute{Obs. de Marseille}%
\CPstreet{2 Pl. Le Verrier}%
\CPtelephone{ +33-9195-9088}%
\CPtelex{ 420241 F}%
\CPnetaddr{ bosma@fromrs51.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{32}{%
USERS=1 PROGS=1 TASKS=2 BAND=optical
}%
\CPmachine{32}{%
SITE=Marseille~Obs.
CPU=Sun-4/330
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.2 MEM=24
PLOT=DEC~LN03+~(Postscript),Digital~(across~network)
TV=Sun-4
TAPE=6250~bpi~across~network,QIC24,Exabyte
AVUSERS=1 AIPS=15OCT89
FRACT=0.3 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Didier Despois}%
\CPaddrline{Observatoire de Bordeaux}%
\CPaddrline{Universite de Bordeaux 1, B.P. 89}%
\CPaddrline{F-33270 Floirac}%
\CPaddrline{FRANCE}%
\CPlastname{Despois, Didier}%
\CPcountry{FRANCE}%
\CPzip{}%
\CPcity{F-33270 Floirac}%
\CPinstitute{Universite de Bordeaux 1, B.P. 89}%
\CPstreet{Observatoire de Bordeaux}%
\CPtelephone{ +33-5686-4330}%
\CPnetaddr{ despois@frobor51.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{33}{%
USERS=5 PROGS=1 TASKS=4 BAND=IR,optical
}%
\CPmachine{33}{%
SITE=U.~Bordeaux
CPU=DEC~Clus.~(u3900;3200)
AP=none
OS=VMS~5.4
DISK=0.4 MEM=32
PLOT=Benson~1113,Apple~Laserwriter~II
TV=DEC~VAXstation
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=1 VLA=50 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Simon Radford}%
\CPaddrline{Universite de Grenoble}%
\CPaddrline{I.R.A.M., Voie 10}%
\CPaddrline{F-38406 St. Martin d'Heres}%
\CPaddrline{FRANCE}%
\CPlastname{Radford, Simon}%
\CPcountry{FRANCE}%
\CPzip{}%
\CPcity{F-38406 St. Martin d'Heres}%
\CPinstitute{Universite de Grenoble}%
\CPstreet{I.R.A.M., Voie 10}%
\CPtelephone{ +33-7682-4900}%
\CPtelex{ 980753 IRAMF}%
\CPnetaddr{ radford@frgag51.bitnet}%
\CPnetaddr{ IRAM02::RADFORD}%
\CPattribute{CP}%
}%
\CPsitedata{34}{%
USERS=5 PROGS=0.2 TASKS=3
}%
\CPmachine{34}{%
SITE=IRAM
CPU=DEC~Clus.~(uVII;u3200)
AP=none
OS=VMS~4.7
DISK=0.5 MEM=8
PLOT=HP~7550A,Printronix~xxx,Apple~Laserwriter+
TV=Sigma~ARGS~7000,DEC~VAXstation
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Diego Cesarsky}%
\CPaddrline{Institut D'Astrophysique}%
\CPaddrline{98 Bis Blvd. Arago}%
\CPaddrline{F-75014 Paris}%
\CPaddrline{FRANCE}%
\CPlastname{Cesarsky, Diego}%
\CPcountry{FRANCE}%
\CPzip{}%
\CPcity{F-75014 Paris}%
\CPinstitute{Institut D'Astrophysique}%
\CPstreet{98 Bis Blvd. Arago}%
\CPtelephone{ +33-1-4320-1425}%
\CPnetaddr{ diego@friap51.bitnet}%
\CPnetaddr{ iapobs::diego}%
\CPattribute{CP}%
}%
\CPsitedata{35}{%
USERS=6 PROGS=1 TASKS=0 BAND=IR
}%
\CPmachine{35}{%
SITE=Inst.~Ap.~Paris
CPU=DEC~Clus.~(780;3*u2000)
AP=none
OS=VMS~x.x
DISK=0.4 MEM=8
PLOT=Versatec~V80
TV=Sigma~ARGS~xxxx
TAPE=?
AVUSERS=? AIPS=?
FRACT=5.5 VLA=60 NONRADIO=40
}%
}%

\CP{
\CPaddr{
\CPaddrline{Alain Kerdraon}%
\CPaddrline{Observatoire de Meudon}%
\CPaddrline{DASOP, 5 Place J. Jansen}%
\CPaddrline{F-92195 Meudon Cedex}%
\CPaddrline{FRANCE}%
\CPlastname{Kerdraon, Alain}%
\CPcountry{FRANCE}%
\CPzip{}%
\CPcity{F-92195 Meudon Cedex}%
\CPinstitute{Observatoire de Meudon}%
\CPstreet{DASOP, 5 Place J. Jansen}%
\CPtelephone{ +33-1-4507-7724}%
\CPnetaddr{ chef@frmeu51.bitnet}%
\CPnetaddr{ haddock@frmeu51.bitnet}%
\CPnetaddr{ 17670::haddock}%
\CPattribute{CP}%
}%
\CPsitedata{36}{%
USERS=15 PROGS=0 TASKS=0
}%
\CPmachine{36}{%
SITE=Obs.~de~Meudon
CPU=DEC~VAX8600
AP=none
OS=VMS~5.2
DISK=2 MEM=44
PLOT=DEC~LN03,Versatec~V80,Apple~NT..
TV=Sigmex~7000
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=2 AIPS=15APR90
FRACT=5 VLA=90 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Ralf Herbertz}%
\CPaddrline{I Physikalisches Institut}%
\CPaddrline{Zuelpicher Strasse 77}%
\CPaddrline{D-5000 Koeln 41}%
\CPaddrline{GERMANY}%
\CPlastname{Herbertz, Ralf}%
\CPcountry{GERMANY}%
\CPzip{}%
\CPcity{D-5000 Koeln 41}%
\CPinstitute{I Physikalisches Institut}%
\CPstreet{Zuelpicher Strasse 77}%
\CPtelephone{ 0221970-3558}%
\CPnetaddr{ HERBERTZ@HOST.PH1.UNI-KOELN.DE}%
\CPattribute{CP}%
}%
\CPsitedata{37}{%
USERS=20 PROGS=1 TASKS=5
}%
\CPmachine{37}{%
SITE=Phys.~Inst.~Koeln
CPU=DEC~Clus.~(2*uVAXII;uVAXIII)
AP=none
OS=VMS~5.3
DISK=1.5 MEM=16
PLOT=?
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=10 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Juergen Schmidt}%
\CPaddrline{Max-Planck-Inst. fuer Radioastronomie}%
\CPaddrline{Auf dem Hugel 69}%
\CPaddrline{D-5300 Bonn 1}%
\CPaddrline{GERMANY}%
\CPlastname{Schmidt, Juergen}%
\CPcountry{GERMANY}%
\CPzip{}%
\CPcity{D-5300 Bonn 1}%
\CPinstitute{Max-Planck-Inst. fuer Radioastronomie}%
\CPstreet{Auf dem Hugel 69}%
\CPnetaddr{ schmidtj\%mpirbn.uucp@unido.uucp}%
\CPnetaddr{ schmidtj@mpifr-bonn.mpg.de}%
\CPattribute{CP}%
}%
\CPsitedata{38}{%
USERS=30 PROGS=1 TASKS=0
}%
\CPmachine{38}{%
SITE=MPIfR~Bonn
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=1.8 MEM=32
PLOT=QMS~800
TV=Sun-4/65
TAPE=1600bpi,6250bpi
AVUSERS=3 AIPS=15APR90
FRACT=50 VLA=80 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Gustaaf van Moorsel}%
\CPaddrline{European Southern Obs.}%
\CPaddrline{Karl Schwarzschildstrasse 2}%
\CPaddrline{D-8046 Garching bei Munchen}%
\CPaddrline{GERMANY}%
\CPlastname{van Moorsel, Gustaaf}%
\CPcountry{GERMANY}%
\CPzip{}%
\CPcity{D-8046 Garching bei Munchen}%
\CPinstitute{European Southern Obs.}%
\CPstreet{Karl Schwarzschildstrasse 2}%
\CPtelephone{ +49-89-3200-6362}%
\CPnetaddr{ gustaaf@dgaeso51.bitnet}%
\CPnetaddr{ 28754::gustaaf}%
\CPattribute{CP}%
}%
\CPsitedata{39}{%
USERS=4 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{39}{%
SITE=ESO~Garching
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~x.x
DISK=0.5 MEM=16
PLOT=?
TV=Sun-4
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT90
FRACT=30 VLA=60 NONRADIO=40
}%
\CPmachine{40}{%
SITE=ESO~Garching
CPU=DEC~VAX8600
AP=none
OS=VMS~4.7
DISK=0.3 MEM=40
PLOT=Versatec~xxxx,QMS~800
TV=Gould~85
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=2 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Istvan Fejes}%
\CPaddrline{Fomi Satellite Geod. Obs.}%
\CPaddrline{Box 546}%
\CPaddrline{H-1373 Budapest}%
\CPaddrline{HUNGARY}%
\CPlastname{Fejes, Istvan}%
\CPcountry{HUNGARY}%
\CPzip{}%
\CPcity{H-1373 Budapest}%
\CPinstitute{Fomi Satellite Geod. Obs.}%
\CPstreet{Box 546}%
\CPtelephone{ 27-10980}%
\CPtelex{ 282241}%
\CPattribute{CP}%
}%
\CPsitedata{41}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{41}{%
SITE=Fomi~Sat.~Geod.~Obs.
CPU=DEC~uVAX~II
AP=none
OS=VMS~x.x
DISK=0.4 MEM=9
PLOT=HP~LaserJet~II
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=1 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Debi Prasad}%
\CPaddrline{Physical Research Lab.}%
\CPaddrline{Room No. 773, Navarangapura}%
\CPaddrline{Ahmedabad, 380009}%
\CPaddrline{INDIA}%
\CPlastname{Prasad, Debi}%
\CPcountry{INDIA}%
\CPzip{}%
\CPcity{Ahmedabad, 380009}%
\CPinstitute{Physical Research Lab.}%
\CPstreet{Room No. 773, Navarangapura}%
\CPtelephone{ 462129}%
\CPtelex{ 0121-397}%
\CPattribute{CP}%
}%
\CPsitedata{42}{%
USERS=10 PROGS=5 TASKS=0 BAND=optical
}%
\CPmachine{42}{%
SITE=PRL~Ahmedabad
CPU=DEC~uVAX~II
AP=none
OS=VMS~x.x
DISK=0.06 MEM=10
PLOT=?
TV=Vibhyor-2001
TAPE=?
AVUSERS=? AIPS=?
FRACT=20 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. T. N. Rengarajan}%
\CPaddrline{Infrared Group, TIFR}%
\CPaddrline{Homi Bhabha Road}%
\CPaddrline{Bombay, 400005}%
\CPaddrline{INDIA}%
\CPlastname{Rengarajan, Dr. T. N.}%
\CPcountry{INDIA}%
\CPzip{}%
\CPcity{Bombay, 400005}%
\CPinstitute{Infrared Group, TIFR}%
\CPstreet{Homi Bhabha Road}%
\CPtelephone{ +91-22-495-2311}%
\CPtelex{ 011-3009 COVE TIFR IN}%
\CPattribute{CP}%
}%
\CPsitedata{43}{%
USERS=10 PROGS=0 TASKS=0
}%
\CPmachine{43}{%
SITE=Bombay~IR
CPU=DEC~VAX780
AP=none
OS=VMS~x.x
DISK=0.05 MEM=2
PLOT=Calcomp
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{A. K. Pati}%
\CPaddrline{Indian Inst. of Astrophysics}%
\CPaddrline{Koramangala}%
\CPaddrline{Bangalore, 560034}%
\CPaddrline{INDIA}%
\CPlastname{Pati, A. K.}%
\CPcountry{INDIA}%
\CPzip{}%
\CPcity{Bangalore, 560034}%
\CPinstitute{Indian Inst. of Astrophysics}%
\CPstreet{Koramangala}%
\CPtelephone{ 0812-530672}%
\CPtelex{ 0845-2763}%
\CPnetaddr{ iiap@vigyan.ernet.in}%
\CPattribute{CP}%
}%
\CPsitedata{44}{%
USERS=4 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{44}{%
SITE=IIA~Bangalore
CPU=DEC~VAX780
AP=none
OS=VMS~4.7
DISK=0.3 MEM=4
PLOT=Printronix~600
TV=COMTAL~Vision~One
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15JAN89
FRACT=15 VLA=0 NONRADIO=55
}%
}%

\CP{
\CPaddr{
\CPaddrline{K. Anantharamaiah}%
\CPaddrline{Raman Research Inst.}%
\CPaddrline{AIPS Site Director}%
\CPaddrline{Bangalore, 560080}%
\CPaddrline{INDIA}%
\CPlastname{Anantharamaiah, K.}%
\CPcountry{INDIA}%
\CPzip{}%
\CPcity{Bangalore, 560080}%
\CPinstitute{Raman Research Inst.}%
\CPstreet{AIPS Site Director}%
\CPtelephone{ 91-812-340122}%
\CPtelex{ 845-2671 (telex) 91-812-340492 (fax)}%
\CPnetaddr{ rri\%vigyan@shakti.emet.in}%
\CPattribute{CP}%
}%
\CPsitedata{45}{%
USERS=2 PROGS=1 TASKS=0
}%
\CPmachine{45}{%
SITE=Raman~Res.~Inst.
CPU=DEC~VAX780
AP=none
OS=VMS~4.2
DISK=0.6 MEM=16
PLOT=Printronix~600,HP~LaserJet~IIP
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=25 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. T. Velusamy}%
\CPaddrline{Radio Astronomy Center, TIFR}%
\CPaddrline{Post Box 8, Udhagamandalam}%
\CPaddrline{Ootacamund, 643001}%
\CPaddrline{INDIA}%
\CPlastname{Velusamy, Dr. T.}%
\CPcountry{INDIA}%
\CPzip{}%
\CPcity{Ootacamund, 643001}%
\CPinstitute{Radio Astronomy Center, TIFR}%
\CPstreet{Post Box 8, Udhagamandalam}%
\CPtelephone{ (0423)-2651,-2032,-2065}%
\CPtelex{ 0853-241 RAC-IN}%
\CPattribute{CP}%
}%
\CPsitedata{46}{%
USERS=20 PROGS=0 TASKS=2 BAND=optical,X-ray
}%
\CPmachine{46}{%
SITE=Ooty~Obs.
CPU=DEC~VAX750
AP=FPS~5205
OS=VMS~4.2
DISK=0.912 MEM=6
PLOT=Printronix~4160,QMS~800,HP~LaserJet~IIP
TV=IIS~75
TAPE=1600bpi
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=50 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Sara Beck}%
\CPaddrline{Tel Aviv University}%
\CPaddrline{Wise Observatory, Dept. of Astronomy}%
\CPaddrline{Ramat Aviv, 66978}%
\CPaddrline{ISRAEL}%
\CPlastname{Beck, Dr. Sara}%
\CPcountry{ISRAEL}%
\CPzip{}%
\CPcity{Ramat Aviv, 66978}%
\CPinstitute{Wise Observatory, Dept. of Astronomy}%
\CPstreet{Tel Aviv University}%
\CPnetaddr{ becks@taunivm.bitnet}%
\CPnetaddr{ fnbenj@weizmann.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{47}{%
USERS=3 PROGS=0 TASKS=0
}%
\CPmachine{47}{%
SITE=Tel~Aviv~U.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.2 MEM=21
PLOT=?
TV=Sun-4
TAPE=Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=10 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Marco Catarzi}%
\CPaddrline{Oss. Astrofisico di Arcetri}%
\CPaddrline{Largo Enrico Fermi 5}%
\CPaddrline{50125 Firenze}%
\CPaddrline{ITALY}%
\CPlastname{Catarzi, Marco}%
\CPcountry{ITALY}%
\CPzip{}%
\CPcity{50125 Firenze}%
\CPinstitute{(no institute?)}%
\CPstreet{Oss. Astrofisico di Arcetri, Largo Enrico Fermi 5}%
\CPtelephone{ +39-55-27521}%
\CPtelex{ 572268 ARCETR I}%
\CPnetaddr{ 38954::catarzi}%
\CPnetaddr{ catarzi@astrfi.infn.it}%
\CPattribute{CP}%
}%
\CPsitedata{48}{%
USERS=6 PROGS=0 TASKS=0 BAND=other
}%
\CPmachine{48}{%
SITE=Arcetri~Obs.
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.8 MEM=4
PLOT=Apple~Laserwriter
TV=Sun-3
TAPE=6250bpi
AVUSERS=2 AIPS=15APR90
FRACT=30 VLA=95 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Prof. Santo Catalano}%
\CPaddrline{Univ. of Catania}%
\CPaddrline{Institute of Astronomy, Viale A. Doria}%
\CPaddrline{95125 Catania}%
\CPaddrline{ITALY}%
\CPlastname{Catalano, Prof. Santo}%
\CPcountry{ITALY}%
\CPzip{}%
\CPcity{95125 Catania}%
\CPinstitute{Institute of Astronomy, Viale A. Doria}%
\CPstreet{Univ. of Catania}%
\CPtelephone{ +39-95 330533}%
\CPtelex{ 970359 ASTRCT}%
\CPnetaddr{ gz9cta61@icineca.bitnet}%
\CPnetaddr{ 40061::santo}%
\CPattribute{CP}%
}%
\CPsitedata{49}{%
USERS=5 PROGS=0 TASKS=0 BAND=optical,UV,other
}%
\CPmachine{49}{%
SITE=U.~Catania
CPU=DEC~VAX750
AP=none
OS=VMS~x.x
DISK=1.2 MEM=11
PLOT=Printronix~300,Apple~Laserwriter+
TV=Ramtek~9460*
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=70 NONRADIO=30
}%
}%

\CP{
\CPaddr{
\CPaddrline{M. Nanni}%
\CPaddrline{Instituto di Radioastronomia}%
\CPaddrline{Via Irnerio 46}%
\CPaddrline{Bologna 40126}%
\CPaddrline{ITALY}%
\CPlastname{Nanni, M.}%
\CPcountry{ITALY}%
\CPzip{}%
\CPcity{Bologna 40126}%
\CPinstitute{Instituto di Radioastronomia}%
\CPstreet{Via Irnerio 46}%
\CPtelephone{ +39-51-287832}%
\CPtelex{ 211664 INFN BO I (will change to 520634)}%
\CPnetaddr{ 38057::nanni}%
\CPnetaddr{ nanni@astbo1.infn.it}%
\CPnetaddr{ 137.204.51.1 Nanni}%
\CPattribute{CP}%
}%
\CPsitedata{50}{%
USERS=15 PROGS=3 TASKS=1 BAND=optical
}%
\CPmachine{50}{%
SITE=Bologna~IRA
CPU=Convex~C-201
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=1.5 MEM=64
PLOT=QMS~800
TV=IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=3 AIPS=15JUL90
FRACT=30 VLA=70 NONRADIO=10
}%
\CPmachine{51}{%
SITE=NOTO~IRA
CPU=DEC~uVAX3300
AP=none
OS=VMS~5.3
DISK=1 MEM=16
PLOT=Postscript
TV=none
TAPE=6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=60 VLA=50 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Prof. Masato Ishiguro}%
\CPaddrline{Nobeyama Radio Obs.}%
\CPaddrline{Tokyo Astronomical Obs.}%
\CPaddrline{Nobeyama, Minamisaku, Nagano 384-13}%
\CPaddrline{JAPAN}%
\CPlastname{Ishiguro, Prof. Masato}%
\CPcountry{JAPAN}%
\CPzip{}%
\CPcity{Nobeyama, Minamisaku, Nagano 384-13}%
\CPinstitute{Tokyo Astronomical Obs.}%
\CPstreet{Nobeyama Radio Obs.}%
\CPtelephone{ +81-0267-98-2831}%
\CPtelex{ TWX 3329005TAONROJ}%
\CPattribute{CP}%
}%
\CPsitedata{52}{%
USERS=36 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{52}{%
SITE=Nobeyama~Radio~Obs.
CPU=Fujitsu~VP50
AP=Vector~Regs
OS=Fujitsu
DISK=2.52 MEM=32
PLOT=?
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=58 VLA=8 NONRADIO=0
}%
\CPmachine{53}{%
SITE=Nobeyama~Radio~Obs.
CPU=Fujitsu~M380S
AP=none
OS=Fujitsu
DISK=2.52 MEM=24
PLOT=Facom~6715E
TV=Graphica~I5048
TAPE=?
AVUSERS=? AIPS=?
FRACT=16 VLA=8 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Yoshihiko Mizumoto}%
\CPaddrline{Department of Physics}%
\CPaddrline{Kobe University}%
\CPaddrline{Rokkodai, Nada-Ku,Kobe 657}%
\CPaddrline{JAPAN}%
\CPlastname{Mizumoto, Yoshihiko}%
\CPcountry{JAPAN}%
\CPzip{}%
\CPcity{Rokkodai, Nada-Ku,Kobe 657}%
\CPinstitute{Kobe University}%
\CPstreet{Department of Physics}%
\CPtelephone{ +81-078-881-1212}%
\CPtelex{ 078-881-7593}%
\CPnetaddr{ aips@gradis.scitec.kobe-u.ac.jp}%
\CPattribute{CP}%
}%
\CPsitedata{54}{%
USERS=2 PROGS=0 TASKS=0 BAND=X-ray
}%
\CPmachine{54}{%
SITE=Kobe~U.~Gamma~Ray
CPU=Sun-3/260C
AP=none
OS=UNIX~Sun~4.2
DISK=0.5 MEM=4
PLOT=Apple~Laserwriter~II
TV=Sun-3
TAPE=QIC24
AVUSERS=1 AIPS=15APR90
FRACT=3 VLA=0 NONRADIO=90
}%
}%

\CP{
\CPaddr{
\CPaddrline{Walter Jaffe}%
\CPaddrline{Sterrewacht Leiden}%
\CPaddrline{Wassenaarseweg 78, Postbus 9513}%
\CPaddrline{2300 RA Leiden}%
\CPaddrline{NETHERLANDS}%
\CPlastname{Jaffe, Walter}%
\CPcountry{NETHERLANDS}%
\CPzip{}%
\CPcity{2300 RA Leiden}%
\CPinstitute{Sterrewacht Leiden}%
\CPstreet{Wassenaarseweg 78, Postbus 9513}%
\CPtelephone{ +31-71-275883}%
\CPnetaddr{ jaffe@hlerul51.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{55}{%
USERS=15 PROGS=2 TASKS=3 BAND=optical
}%
\CPmachine{55}{%
SITE=Leiden~Obs.
CPU=Convex~C-220
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=2.5 MEM=32
PLOT=HP~LaserJet~III
TV=?
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=3 AIPS=15OCT89
FRACT=25 VLA=90 NONRADIO=10
}%
\CPmachine{56}{%
SITE=Leiden~Obs.
CPU=DECstation~3100
AP=none
OS=VMS~4.6
DISK=0.3 MEM=12
PLOT=?
TV=DECstation
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=2 AIPS=15OCT89
FRACT=40 VLA=80 NONRADIO=10
}%
\CPmachine{57}{%
SITE=Leiden~Obs.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3c
DISK=0.3 MEM=12
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=40 VLA=80 NONRADIO=10
}%
\CPmachine{58}{%
SITE=Leiden~Obs.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3c
DISK=0.3 MEM=12
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=40 VLA=80 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Tony Foley}%
\CPaddrline{Radiosterrenwacht Dwingeloo}%
\CPaddrline{NFRA, POSTBUS 2}%
\CPaddrline{7990 AA Dwingeloo}%
\CPaddrline{NETHERLANDS}%
\CPlastname{Foley, Tony}%
\CPcountry{NETHERLANDS}%
\CPzip{}%
\CPcity{7990 AA Dwingeloo}%
\CPinstitute{(no institute?)}%
\CPstreet{Radiosterrenwacht Dwingeloo, NFRA, POSTBUS 2}%
\CPtelephone{ +31-5219-7244}%
\CPnetaddr{ estcs1::dwingelo::foley}%
\CPnetaddr{ foley\%r2mvx1.r2msur.surfnet@hasara5.bitnet}%
\CPnetaddr{ foley@nfra.nl}%
\CPattribute{CP}%
}%
\CPsitedata{59}{%
USERS=10 PROGS=1 TASKS=1 BAND=optical
}%
\CPmachine{59}{%
SITE=NFRA~Dwingeloo
CPU=Alliant~FX/8-3CE
AP=Vector~Regs
OS=UNIX~Concentrix~5.5
DISK=0.8 MEM=24
PLOT=QMS~800
TV=DECstation~(XAS)
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=3 AIPS=15APR90
FRACT=40 VLA=75 NONRADIO=20
}%
\CPmachine{60}{%
SITE=NFRA~Dwingeloo
CPU=DEC~uVAX3400
AP=none
OS=VMS~5.3
DISK=0.4 MEM=12
PLOT=QMS~800,Calcomp~81
TV=Gould~85
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=30 VLA=50 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{J.M. van der Hulst}%
\CPaddrline{Univ. of Groningen}%
\CPaddrline{Kapteyn Lab., Postbus 800}%
\CPaddrline{9700 AV Groningen}%
\CPaddrline{NETHERLANDS}%
\CPlastname{van der Hulst, J.M.}%
\CPcountry{NETHERLANDS}%
\CPzip{}%
\CPcity{9700 AV Groningen}%
\CPinstitute{Kapteyn Lab., Postbus 800}%
\CPstreet{Univ. of Groningen}%
\CPtelephone{ +31-50-634054}%
\CPtelex{ 53572 stars nl}%
\CPnetaddr{ vdhulst@hgrrug5.bitnet}%
\CPnetaddr{ vdhulst@rug.nl}%
\CPattribute{CP}%
}%
\CPsitedata{61}{%
USERS=10 PROGS=0 TASKS=0 BAND=optical
}%
\CPmachine{61}{%
SITE=Kapteyn~Institute
CPU=Sun~Clus.~(4*4/65)
AP=none
OS=UNIX~Sun~4.1
DISK=2.5 MEM=16
PLOT=QMS~PS-810
TV=Sun-4
TAPE=?
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=80 NONRADIO=10
}%
\CPmachine{62}{%
SITE=Kapteyn~Institute
CPU=Alliant~FX/80
AP=Vector~Regs
OS=UNIX~Concentrix~5.5
DISK=2.5 MEM=32
PLOT=QMS~PS-810
TV=Sun-4,DECstations
TAPE=1600bpi,6250bpi,DAT
AVUSERS=2 AIPS=15APR90
FRACT=30 VLA=80 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Goran Pilbratt}%
\CPaddrline{ESA/ESTEC-SA}%
\CPaddrline{Space Science Dept./Astrophys. Div., Postbus 299}%
\CPaddrline{NL-2200 AG Noordwijk}%
\CPaddrline{NETHERLANDS}%
\CPlastname{Pilbratt, Goran}%
\CPcountry{NETHERLANDS}%
\CPzip{}%
\CPcity{NL-2200 AG Noordwijk}%
\CPinstitute{Space Science Dept./Astrophys. Div., Postbus 299}%
\CPstreet{ESA/ESTEC-SA}%
\CPtelephone{ +31-1719-83621}%
\CPtelex{ Telefax +31 1719 84690}%
\CPnetaddr{ gpilbrat@estec.bitnet}%
\CPnetaddr{ estcs1::gpilbrat}%
\CPattribute{CP}%
}%
\CPsitedata{63}{%
USERS=1 PROGS=0 TASKS=0
}%
\CPmachine{63}{%
SITE=ESA/ESTEC-SA
CPU=DEC~Clus.~(3600;3100;2000;5000)
AP=none
OS=VMS~5.4,UNIX~Ultrix~4.0
DISK=1 MEM=32
PLOT=DEC~Scriptprinter
TV=Gould
TAPE=1600bpi,6250bpi,TK70
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bill Cotton}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Edgemont Road}%
\CPaddrline{Charlottesville, VA, 22903-2475}%
\CPaddrline{NRAO}%
\CPlastname{Cotton, Bill}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Charlottesville, VA, 22903-2475}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{Edgemont Road}%
\CPtelephone{ 804-296-0319}%
\CPtelex{ 910-997-0174}%
\CPnetaddr{ bcotton@nrao.edu}%
\CPnetaddr{ bcotton@nrao.bitnet}%
\CPnetaddr{ 6654::bcotton}%
\CPattribute{CP}%
}%
\CPsitedata{64}{%
USERS=8 PROGS=4 TASKS=230 BAND=optical,X-ray
}%
\CPmachine{64}{%
SITE=NRAO~CVAX
CPU=DEC~VAX780
AP=FPS~120B
OS=VMS~4.6
DISK=1.3 MEM=8
PLOT=QMS~800,Versatec~1200A
TV=none
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=0.3 AIPS=15APR89
FRACT=0.3 VLA=50 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Brian Glendenning}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Edgemont Rd.}%
\CPaddrline{Charlottesville, VA, 22903-2475}%
\CPaddrline{NRAO}%
\CPlastname{Glendenning, Brian}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Charlottesville, VA, 22903-2475}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{Edgemont Rd.}%
\CPtelephone{ 804-296-0286}%
\CPnetaddr{ bglenden@nrao.edu}%
\CPnetaddr{ bglenden@nrao.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{65}{%
USERS=25 PROGS=4 TASKS=230
}%
\CPmachine{65}{%
SITE=NRAO~LEMUR
CPU=IBM~RS-6000
AP=none
OS=UNIX~AIX
DISK=0.8 MEM=32
PLOT=QMS~800,QMS~PS-800
TV=IBM~color~(8-bit)
TAPE=800bpi,1600bpi,6250bpi,Exabyte
AVUSERS=1.5 AIPS=15APR91
FRACT=100 VLA=80 NONRADIO=0
}%
\CPmachine{66}{%
SITE=NRAO~COLOBUS
CPU=DECstation~3100
AP=none
OS=UNIX~Ultrix~4.0
DISK=0.3 MEM=16
PLOT=QMS~800,QMS~PS-800
TV=DECstation~color~(8-bit)
TAPE=TK50
AVUSERS=0.5 AIPS=15APR91
FRACT=100 VLA=0 NONRADIO=0
}%
\CPmachine{67}{%
SITE=NRAO~MANDRILL
CPU=Sun-4/110
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=8
PLOT=QMS~800,QMS~PS-800
TV=Sun-4~color
TAPE=QIC24,QIC11
AVUSERS=1 AIPS=15APR91
FRACT=10 VLA=80 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Pat Murphy}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Edgemont Road}%
\CPaddrline{Charlottesville, VA, 22903-2475}%
\CPaddrline{NRAO}%
\CPlastname{Murphy, Pat}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Charlottesville, VA, 22903-2475}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{Edgemont Road}%
\CPtelephone{ 804-296-0372}%
\CPnetaddr{ pmurphy@nrao.edu}%
\CPnetaddr{ pmurphy@nrao.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{68}{%
USERS=25 PROGS=4 TASKS=230 BAND=optical,X-ray,IR
}%
\CPmachine{68}{%
SITE=NRAO~NRAO1
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=5.8 MEM=32
PLOT=QMS~800,QMS~PS-800,Versatec~1200A
TV=IIS~70E,IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=6 AIPS=15APR91
FRACT=90 VLA=80 NONRADIO=1
}%
\CPmachine{69}{%
SITE=NRAO~CV~AIPS~GROUP
CPU=Sun~Clus.~(3/50;3/60G;3/110)
AP=none
OS=UNIX~Sun~4.1
DISK=0.63 MEM=12
PLOT=QMS~800,QMS~PS-800,Versatec~1200A
TV=Sun-3~mono,Sun-3~color
TAPE=QIC11,QIC24
AVUSERS=3 AIPS=15APR91
FRACT=100 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Frank Ghigo}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Jansky Lab.}%
\CPaddrline{Green Bank, WV, 24944}%
\CPaddrline{NRAO}%
\CPlastname{Ghigo, Frank}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Green Bank, WV, 24944}%
\CPinstitute{Jansky Lab.}%
\CPstreet{Nat. Radio Astronomy Obs.}%
\CPtelephone{ 304-456-2210}%
\CPtelex{ 710-938-1530}%
\CPnetaddr{ fghigo@nrao.edu}%
\CPnetaddr{ fghigo@nrao.bitnet}%
\CPnetaddr{ 6654::fghigo}%
\CPattribute{CP}%
}%
\CPsitedata{70}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{70}{%
SITE=NRAO~GB~SUN
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.3 MEM=8
PLOT=NEC~LC-800,Postscript
TV=Sun-3~color
TAPE=6250bpi,QIC24
AVUSERS=1 AIPS=15APR90
FRACT=0.1 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Darrel Emerson}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Campus Bldg. 65, 949 N. Cherry Ave.}%
\CPaddrline{Tucson, 85721}%
\CPaddrline{NRAO}%
\CPlastname{Emerson, Darrel}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Tucson, 85721}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{Campus Bldg. 65, 949 N. Cherry Ave.}%
\CPtelephone{ 602-882-8250}%
\CPnetaddr{ demerson@nrao.edu}%
\CPnetaddr{ 6654::DEMERSON}%
\CPattribute{CP}%
}%
\CPsitedata{71}{%
USERS=3 PROGS=0 TASKS=0
}%
\CPmachine{71}{%
SITE=NRAO~KILLIANS
CPU=Sun-4/60GX
AP=none
OS=UNIX~Sun~4.1
DISK=0.65 MEM=8
PLOT=QMS~800
TV=Sun-4~color
TAPE=QIC24,QIC150
AVUSERS=1 AIPS=15JUL90
FRACT=10 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Timothy J. Cornwell}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{AOC}%
\CPaddrline{Socorro, NM, 87801}%
\CPaddrline{NRAO}%
\CPlastname{Cornwell, Timothy J.}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Socorro, NM, 87801}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{AOC}%
\CPtelephone{ 505-835-7333}%
\CPnetaddr{ tcornwel@nrao.edu}%
\CPnetaddr{ tcornwel@nrao.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{72}{%
USERS=1 PROGS=6 TASKS=7 BAND=radar,optical
}%
\CPmachine{72}{%
SITE=NRAO~AOC~SPARC2
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=4
PLOT=QMS~1500
TV=Sun-4
TAPE=?
AVUSERS=1 AIPS=15APR91
FRACT=10 VLA=50 NONRADIO=50
}%
}%

\CP{
\CPaddr{
\CPaddrline{Phil J. Diamond}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{AOC}%
\CPaddrline{Socorro, NM, 87801}%
\CPaddrline{NRAO}%
\CPlastname{Diamond, Phil J.}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Socorro, NM, 87801}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{AOC}%
\CPtelephone{ 505-835-7365}%
\CPnetaddr{ pdiamond@nrao.edu}%
\CPnetaddr{ pdiamond@nrao.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{73}{%
USERS=1 PROGS=1 TASKS=230
}%
\CPmachine{73}{%
SITE=NRAO~KONG
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.2 MEM=16
PLOT=?
TV=Sun-3~(X-Windows)
TAPE=?
AVUSERS=1 AIPS=15APR91
FRACT=90 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bill Junor}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{AOC}%
\CPaddrline{Socorro, NM, 87801}%
\CPaddrline{NRAO}%
\CPlastname{Junor, Bill}%
\CPcountry{NRAO}%
\CPzip{}%
\CPcity{Socorro, NM, 87801}%
\CPinstitute{Nat. Radio Astronomy Obs.}%
\CPstreet{AOC}%
\CPtelephone{ 505-835-7210}%
\CPnetaddr{ bjunor@nrao.edu}%
\CPnetaddr{ bjunor@nrao.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{74}{%
USERS=400 PROGS=6 TASKS=230 BAND=radar,optical
}%
\CPmachine{74}{%
SITE=NRAO~YUCCA
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=4.2 MEM=64
PLOT=QMS~1500,QMS~1500~PS,IBM~4019
TV=IIS~70
TAPE=1600bpi,6250bpi
AVUSERS=6 AIPS=15APR91
FRACT=90 VLA=95 NONRADIO=5
}%
\CPmachine{75}{%
SITE=NRAO~CHOLLA
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=3.5 MEM=64
PLOT=QMS~1500,QMS~1500~PS,IBM~4019
TV=IIS~70,IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=5.5 AIPS=15APR91
FRACT=75 VLA=99 NONRADIO=0.5
}%
\CPmachine{76}{%
SITE=NRAO~VAX1
CPU=DEC~VAX780
AP=FPS~120B
OS=VMS~4.7
DISK=2.2 MEM=4
PLOT=QMS~1500
TV=IIS~70
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=10 VLA=95 NONRADIO=2
}%
\CPmachine{77}{%
SITE=NRAO~OUTBAX
CPU=DEC~VAX750
AP=FPS~120B
OS=VMS~4.7
DISK=1.13 MEM=3
PLOT=QMS~800
TV=IIS~70
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=90 VLA=95 NONRADIO=5
}%
\CPmachine{78}{%
SITE=NRAO~NEGRA
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=16
PLOT=?
TV=Sun-3~color
TAPE=?
AVUSERS=1 AIPS=15APR91
FRACT=95 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Antonio Ruiz-Falco}%
\CPaddrline{Inst. de Astrofisica de Andalucia}%
\CPaddrline{Ap. Correos 2144}%
\CPaddrline{18080 Granada}%
\CPaddrline{SPAIN}%
\CPlastname{Ruiz-Falco, Antonio}%
\CPcountry{SPAIN}%
\CPzip{}%
\CPcity{18080 Granada}%
\CPinstitute{Inst. de Astrofisica de Andalucia}%
\CPstreet{Ap. Correos 2144}%
\CPtelephone{ +34-58-121311}%
\CPtelex{ 78573}%
\CPnetaddr{ system@iaa.es}%
\CPnetaddr{ RFALCO@IAA.ES}%
\CPnetaddr{ 16988::system}%
\CPattribute{CP}%
}%
\CPsitedata{79}{%
USERS=10 PROGS=1 TASKS=1
}%
\CPmachine{79}{%
SITE=Inst.~Ast.~Andalucia
CPU=DEC~VAX6310
AP=none
OS=VMS~5.4
DISK=3 MEM=32
PLOT=DEC~LN03+
TV=?
TAPE=6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=10 VLA=100 NONRADIO=0
}%
\CPmachine{80}{%
SITE=Inst.~Ast.~Andalucia
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.2
DISK=0.43 MEM=10
PLOT=DEC~LN03+
TV=?
TAPE=1600bpi
AVUSERS=1 AIPS=15APR90
FRACT=1 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Aage Sandqvist}%
\CPaddrline{Stockholm Obs.}%
\CPaddrline{AIPS Manager}%
\CPaddrline{S-133 36 Saltsjobaden}%
\CPaddrline{SWEDEN}%
\CPlastname{Sandqvist, Dr. Aage}%
\CPcountry{SWEDEN}%
\CPzip{}%
\CPcity{S-133 36 Saltsjobaden}%
\CPinstitute{Stockholm Obs.}%
\CPstreet{AIPS Manager}%
\CPtelephone{ +46-8-164447}%
\CPtelex{ 12972 SOBSERV S}%
\CPnetaddr{ sandqvis@astro.su.se}%
\CPnetaddr{ aips@astro.su.se}%
\CPattribute{CP}%
}%
\CPsitedata{81}{%
USERS=4 PROGS=1 TASKS=1
}%
\CPmachine{81}{%
SITE=Stockholm~Obs.
CPU=Alliant~FX/8-4CE
AP=Vector~Regs
OS=UNIX~Concentrix~x.x
DISK=1.2 MEM=48
PLOT=Apple~Laserwriter
TV=Raster~Technology~1/80
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT88
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Lars B. Baath}%
\CPaddrline{Onsala Space Obs.}%
\CPaddrline{Chalmers Inst. Technology}%
\CPaddrline{S-43900 Onsala}%
\CPaddrline{SWEDEN}%
\CPlastname{Baath, Lars B.}%
\CPcountry{SWEDEN}%
\CPzip{}%
\CPcity{S-43900 Onsala}%
\CPinstitute{Chalmers Inst. Technology}%
\CPstreet{Onsala Space Obs.}%
\CPtelephone{ +46-300-60650}%
\CPtelex{ Sweden 2400 ONSPACE}%
\CPnetaddr{ lbb@oso.chalmers.se}%
\CPattribute{CP}%
}%
\CPsitedata{82}{%
USERS=7 PROGS=1 TASKS=15 BAND=IR,other
}%
\CPmachine{82}{%
SITE=Onsala~Space~Obs.
CPU=Convex~C-1/XL
AP=Vector~Regs
OS=UNIX~Convex~x.x
DISK=1.7 MEM=32
PLOT=QMS~800,TI~2108
TV=IIS~75
TAPE=?
AVUSERS=? AIPS=?
FRACT=80 VLA=10 NONRADIO=1
}%
}%

\CP{
\CPaddr{
\CPaddrline{David J. Titterington}%
\CPaddrline{Cavendish Lab.}%
\CPaddrline{Mullard Radio Astronomy Obs., Madingley Road}%
\CPaddrline{Cambridge CB3 0HE}%
\CPaddrline{UK}%
\CPlastname{Titterington, David J.}%
\CPcountry{UK}%
\CPzip{}%
\CPcity{Cambridge CB3 0HE}%
\CPinstitute{Mullard Radio Astronomy Obs., Madingley Road}%
\CPstreet{Cavendish Lab.}%
\CPtelephone{ +44-223-337733}%
\CPtelex{ 81292}%
\CPnetaddr{ djt@uk.ac.cam.phy-ravx}%
\CPattribute{CP}%
}%
\CPsitedata{83}{%
USERS=10 PROGS=2 TASKS=0 BAND=optical
}%
\CPmachine{83}{%
SITE=MRAO
CPU=Sun~Clus.~(2*4/60)
AP=none
OS=UNIX~Sun~4.0
DISK=0.6 MEM=12
PLOT=?
TV=Sun-4~(Sunview)
TAPE=6250bpi,QIC24,Exabyte
AVUSERS=2 AIPS=15JUL90
FRACT=90 VLA=80 NONRADIO=10
}%
\CPmachine{84}{%
SITE=MRAO
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=0.1 MEM=9
PLOT=Omnilaser~2108
TV=Digisolve~Ikon
TAPE=6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT87
FRACT=0.1 VLA=0 NONRADIO=50
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Robert A. Laing}%
\CPaddrline{Royal Greenwich Obs.}%
\CPaddrline{Madingley Road}%
\CPaddrline{Cambridge CB3 OE2}%
\CPaddrline{UK}%
\CPlastname{Laing, Dr. Robert A.}%
\CPcountry{UK}%
\CPzip{}%
\CPcity{Cambridge CB3 OE2}%
\CPinstitute{Royal Greenwich Obs.}%
\CPstreet{Madingley Road}%
\CPtelephone{ +44-223-337733}%
\CPnetaddr{ rl@starlink.ast.cambridge.ac.uk}%
\CPattribute{CP}%
}%
\CPsitedata{85}{%
USERS=2 PROGS=0 TASKS=0 BAND=optical
}%
\CPmachine{85}{%
SITE=R.~Greenwich~Obs.
CPU=DEC~Clus.~(uVAX3400)
AP=none
OS=VMS~5.3-1
DISK=0.25 MEM=20
PLOT=Canon~laser,Printronix~xxx
TV=Digisolve~Ikon
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT88
FRACT=2 VLA=90 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Richard Prestage}%
\CPaddrline{Royal Observatory}%
\CPaddrline{Blackford Hill}%
\CPaddrline{Edinburgh EH9 3HJ}%
\CPaddrline{UK}%
\CPlastname{Prestage, Richard}%
\CPcountry{UK}%
\CPzip{}%
\CPcity{Edinburgh EH9 3HJ}%
\CPinstitute{Royal Observatory}%
\CPstreet{Blackford Hill}%
\CPtelephone{ +44-31-667-3321}%
\CPnetaddr{ rmp@star.roe.ac.uk}%
\CPattribute{CP}%
}%
\CPsitedata{86}{%
USERS=5 PROGS=0 TASKS=0
}%
\CPmachine{86}{%
SITE=R.~Obs.~Edinburgh
CPU=DEC~VAX780
AP=none
OS=VMS~x.x
DISK=0.256 MEM=8
PLOT=Versatec~xxxx,Canon~LP68~II
TV=Sigma~ARGS~xxxx
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dave Shone}%
\CPaddrline{Nuffield Radio Astron. Labs.}%
\CPaddrline{Jodrell Bank}%
\CPaddrline{Macclesfield, Cheshire SK11 9DL}%
\CPaddrline{UK}%
\CPlastname{Shone, Dave}%
\CPcountry{UK}%
\CPzip{}%
\CPcity{Macclesfield, Cheshire SK11 9DL}%
\CPinstitute{Nuffield Radio Astron. Labs.}%
\CPstreet{Jodrell Bank}%
\CPtelephone{ +44-477-71321}%
\CPtelex{ 36149 JODREL G}%
\CPnetaddr{ dls@star.jb.man.ac.uk}%
\CPnetaddr{ aipman@star.jb.man.ac.uk}%
\CPattribute{CP}%
}%
\CPsitedata{87}{%
USERS=25 PROGS=0.3 TASKS=6 BAND=optical
}%
\CPmachine{87}{%
SITE=Jodrell~Bank
CPU=Alliant~FX/8-4CE
AP=Vector~Regs
OS=UNIX~Concentrix~5.5.03
DISK=1.6 MEM=24
PLOT=Canon~laser
TV=Apollo~DN560,Digisolve~Ikon,Sun-4
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=60 VLA=45 NONRADIO=2
}%
\CPmachine{88}{%
SITE=Jodrell~Bank
CPU=DEC~uVAX3400
AP=PSAP
OS=VMS~5.3
DISK=1.5 MEM=20
PLOT=Canon~laser
TV=Sigma~ARGS~xxxx,Digisolve~Ikon
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=20 VLA=45 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Harry Lehto}%
\CPaddrline{Univ. of Oxford, Dept. of Astrophys.}%
\CPaddrline{Nucl. and Astrophys. Bldg., Keble Rd.}%
\CPaddrline{Oxford OX1 3RH}%
\CPaddrline{UK}%
\CPlastname{Lehto, Harry}%
\CPcountry{UK}%
\CPzip{}%
\CPcity{Oxford OX1 3RH}%
\CPinstitute{Univ. of Oxford, Dept. of Astrophys.}%
\CPstreet{Nucl. and Astrophys. Bldg., Keble Rd.}%
\CPnetaddr{ hl@uk.ac.ox.astro}%
\CPnetaddr{ RLVAD::OXVAD::HL}%
\CPattribute{CP}%
}%
\CPsitedata{89}{%
USERS=3 PROGS=1 TASKS=0
}%
\CPmachine{89}{%
SITE=U.~Oxford~Astrophys.
CPU=DEC~Clus.~(3800;uVAXII;3100)
AP=none
OS=VMS~5.3-1
DISK=0.1 MEM=2.1
PLOT=Canon~Mark~II
TV=Digisolve~Ikon
TAPE=1600bpi,6250bpi,CIFER,Exabyte
AVUSERS=1 AIPS=15OCT89
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Willem A. Baan}%
\CPaddrline{Arecibo Obs.}%
\CPaddrline{P.O. Box 995}%
\CPaddrline{Arecibo, PR, 00613}%
\CPaddrline{USA}%
\CPlastname{Baan, Willem A.}%
\CPcountry{USA}%
\CPzip{00613}%
\CPcity{Arecibo, PR,}%
\CPinstitute{Arecibo Obs.}%
\CPstreet{P.O. Box 995}%
\CPtelephone{ 809-878-2612}%
\CPtelex{ 385638}%
\CPnetaddr{ wbaan@naic.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{90}{%
USERS=10 PROGS=0.5 TASKS=2 BAND=IR,other,radar
}%
\CPmachine{90}{%
SITE=Arecibo~Obs.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.1
DISK=1 MEM=8
PLOT=QMS~800,QMS~PS-xxx
TV=Sun-4
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=30 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Lori Allen}%
\CPaddrline{Univ. of Massachusetts}%
\CPaddrline{Five College Radio Astron. Obs., Graduate Research}%
\CPaddrline{Amherst, MA, 01003}%
\CPaddrline{USA}%
\CPlastname{Allen, Lori}%
\CPcountry{USA}%
\CPzip{01003}%
\CPcity{Amherst, MA,}%
\CPinstitute{Five College Radio Astron. Obs., Graduate Research}%
\CPstreet{Univ. of Massachusetts}%
\CPtelephone{ 413-545-0731}%
\CPnetaddr{ allen@phast.umass.edu}%
\CPattribute{CP}%
}%
\CPsitedata{91}{%
USERS=7 PROGS=0 TASKS=4 BAND=IR,optical
}%
\CPmachine{91}{%
SITE=U.~Mass.
CPU=DEC~VAX750
AP=none
OS=VMS~4.7
DISK=0.5 MEM=8
PLOT=HP~7475A,Apple~Laserwriter~II
TV=Gould~FD5000~(dead)
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=50 NONRADIO=50
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Colin J. Lonsdale}%
\CPaddrline{Haystack Obs.}%
\CPaddrline{Route 40}%
\CPaddrline{Westford, MA, 01886}%
\CPaddrline{USA}%
\CPlastname{Lonsdale, Dr. Colin J.}%
\CPcountry{USA}%
\CPzip{01886}%
\CPcity{Westford, MA,}%
\CPinstitute{Haystack Obs.}%
\CPstreet{Route 40}%
\CPtelephone{ 617-692-4765}%
\CPnetaddr{ cjl@wells.haystack.edu}%
\CPattribute{CP}%
}%
\CPsitedata{92}{%
USERS=6 PROGS=1 TASKS=2 BAND=holography,optical
}%
\CPmachine{92}{%
SITE=Haystack~Obs.
CPU=Sun-4/65
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.4 MEM=8
PLOT=HP~LaserJet+,HP~LaserJet~II,HP~LaserJet~III
TV=Sun-4~(Sunview)
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=70 VLA=85 NONRADIO=15
}%
\CPmachine{93}{%
SITE=Haystack~Obs.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.4 MEM=16
PLOT=HP~LaserJet+,HP~LaserJet~II,HP~LaserJet~III
TV=Sun-4~(Sunview)
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=85 VLA=70 NONRADIO=30
}%
\CPmachine{94}{%
SITE=Haystack~Obs.
CPU=Sun-4/260
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.4 MEM=8
PLOT=HP~LaserJet+,HP~LaserJet~II,HP~LaserJet~III
TV=Sun-4~(Sunview)
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=70 VLA=85 NONRADIO=15
}%
}%

\CP{
\CPaddr{
\CPaddrline{Sam Conner}%
\CPaddrline{Massachusetts Inst. of Tech.}%
\CPaddrline{Room 26-331, Dept. of Physics}%
\CPaddrline{Cambridge, MA, 02116}%
\CPaddrline{USA}%
\CPlastname{Conner, Sam}%
\CPcountry{USA}%
\CPzip{02116}%
\CPcity{Cambridge, MA,}%
\CPinstitute{Room 26-331, Dept. of Physics}%
\CPstreet{Massachusetts Inst. of Tech.}%
\CPtelephone{ 617-253-3074}%
\CPnetaddr{ conner@ewt.mit.edu}%
\CPattribute{CP}%
}%
\CPsitedata{95}{%
USERS=4 PROGS=0 TASKS=0
}%
\CPmachine{95}{%
SITE=MIT~Phys.
CPU=Apollo~DN3000
AP=none
OS=UNIX~x.x
DISK=0.3 MEM=4
PLOT=HP~LaserJet
TV=Apollo~DN3000~(1-bit)
TAPE=?
AVUSERS=? AIPS=?
FRACT=20 VLA=80 NONRADIO=20
}%
\CPmachine{96}{%
SITE=MIT~Phys.
CPU=Apollo~DN320
AP=none
OS=UNIX~x.x
DISK=0.3 MEM=4
PLOT=HP~LaserJet
TV=Apollo~DN320~(1-bit)
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=80 NONRADIO=20
}%
\CPmachine{97}{%
SITE=MIT~Phys.
CPU=Apollo~DN660
AP=none
OS=UNIX~x.x
DISK=0.3 MEM=6
PLOT=HP~LaserJet
TV=Apollo~DN660~(8-bit)
TAPE=?
AVUSERS=? AIPS=?
FRACT=20 VLA=80 NONRADIO=20
}%
}%

\CP{
\CPaddr{
\CPaddrline{Shoshana Rosenthal}%
\CPaddrline{Smithsonian Astrophysical Obs.}%
\CPaddrline{60 Garden Street, M.S. 42}%
\CPaddrline{Cambridge, MA, 02138}%
\CPaddrline{USA}%
\CPlastname{Rosenthal, Shoshana}%
\CPcountry{USA}%
\CPzip{02138}%
\CPcity{Cambridge, MA,}%
\CPinstitute{Smithsonian Astrophysical Obs.}%
\CPstreet{60 Garden Street, M.S. 42}%
\CPtelephone{ 617-495-7267}%
\CPnetaddr{ shoshana@cfa2.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{98}{%
USERS=12 PROGS=1 TASKS=0 BAND=optical,X-ray
}%
\CPmachine{98}{%
SITE=SAO~Cambridge
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=2.5 MEM=32
PLOT=?
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=2.5 AIPS=15OCT90
FRACT=50 VLA=40 NONRADIO=15
}%
\CPmachine{99}{%
SITE=SAO~Cambridge
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=2.5 MEM=32
PLOT=QMS~800
TV=IIS
TAPE=1600bpi,6250bpi
AVUSERS=2.5 AIPS=15OCT90
FRACT=50 VLA=40 NONRADIO=15
}%
\CPmachine{100}{%
SITE=SAO~Cambridge
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=1.7 MEM=9
PLOT=QMS~800
TV=DEC~uVAX~GPX
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=20 VLA=40 NONRADIO=0
}%
\CPmachine{101}{%
SITE=SAO~Cambridge
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=0.75 MEM=10
PLOT=QMS~800
TV=DEC~uVAX~GPX
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=20 VLA=40 NONRADIO=0
}%
\CPmachine{102}{%
SITE=SAO~Cambridge
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=0.5 MEM=16
PLOT=QMS~800
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Jacqueline N. Hewitt}%
\CPaddrline{Dept. of Physics}%
\CPaddrline{Room 26-327, MIT}%
\CPaddrline{Cambridge, MA, 02139}%
\CPaddrline{USA}%
\CPlastname{Hewitt, Jacqueline N.}%
\CPcountry{USA}%
\CPzip{02139}%
\CPcity{Cambridge, MA,}%
\CPinstitute{Dept. of Physics}%
\CPstreet{Room 26-327, MIT}%
\CPtelephone{ 617-253-3071}%
\CPnetaddr{ jhewitt@athena.mit.edu}%
\CPattribute{CP}%
}%
\CPsitedata{103}{%
USERS=9 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{103}{%
SITE=MIT~Phys.
CPU=Sun~Clus.~(4/330;2*4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=1.3 MEM=16
PLOT=?
TV=Sun-4~(Sunview)
TAPE=Exabyte
AVUSERS=2 AIPS=15APR90
FRACT=90 VLA=95 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Joseph Lehar}%
\CPaddrline{Massachusetts Inst. of Tech.}%
\CPaddrline{Room 26-331, Dept. of Physics}%
\CPaddrline{Cambridge, MA, 02139}%
\CPaddrline{USA}%
\CPlastname{Lehar, Joseph}%
\CPcountry{USA}%
\CPzip{02139}%
\CPcity{Cambridge, MA,}%
\CPinstitute{Room 26-331, Dept. of Physics}%
\CPstreet{Massachusetts Inst. of Tech.}%
\CPtelephone{ 617-253-3074}%
\CPnetaddr{ lehar@ewt.mit.edu}%
\CPattribute{CP}%
}%
\CPsitedata{104}{%
USERS=5 PROGS=0 TASKS=0
}%
\CPmachine{104}{%
SITE=MIT~Radio~Astron.
CPU=DEC~Clus.~(2*3500;750)
AP=none
OS=VMS~4.5
DISK=0.1 MEM=24
PLOT=QMS~Lasergrafix
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=30 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Robert F. Willson}%
\CPaddrline{Tufts University}%
\CPaddrline{Dept. of Physics \&\ Astronomy, Robinson Hall}%
\CPaddrline{Medford, MA, 02155}%
\CPaddrline{USA}%
\CPlastname{Willson, Robert F.}%
\CPcountry{USA}%
\CPzip{02155}%
\CPcity{Medford, MA,}%
\CPinstitute{Dept. of Physics \&\ Astronomy, Robinson Hall}%
\CPstreet{Tufts University}%
\CPtelephone{ 617-628-5000}%
\CPattribute{CP}%
}%
\CPsitedata{105}{%
USERS=3 PROGS=0 TASKS=0
}%
\CPmachine{105}{%
SITE=Tufts~U.
CPU=Sun-X/xxx
AP=none
OS=UNIX~Sun~x.x
DISK=0.3
PLOT=HP~LaserJet~II
TV=Sun
TAPE=QIC24
AVUSERS=1 AIPS=15APR90
FRACT=50 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{David Bradford}%
\CPaddrline{Boston Univ.}%
\CPaddrline{Astronomy Dept., 725 Commonwealth Ave.}%
\CPaddrline{Boston, MA, 02215}%
\CPaddrline{USA}%
\CPlastname{Bradford, David}%
\CPcountry{USA}%
\CPzip{02215}%
\CPcity{Boston, MA,}%
\CPinstitute{Astronomy Dept., 725 Commonwealth Ave.}%
\CPstreet{Boston Univ.}%
\CPtelephone{ 617-353-4884}%
\CPnetaddr{ bradford@buasta.bitnet}%
\CPnetaddr{ buasta::bradford}%
\CPnetaddr{ bradford@buasta.bu.edu}%
\CPattribute{CP}%
}%
\CPsitedata{106}{%
USERS=6 PROGS=0 TASKS=2 BAND=IR,optical
}%
\CPmachine{106}{%
SITE=Boston~U.
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3-1
DISK=2 MEM=13
PLOT=QMS~PS-820,QMS~Talaris~1590-T
TV=Gould~IP8400
TAPE=1600bpi,6250bpi,TK50
AVUSERS=1 AIPS=15OCT88
FRACT=30 VLA=30 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Y. F. Zhang}%
\CPaddrline{Boston University}%
\CPaddrline{Astronomy Dept., 725 Commonwealth Ave.}%
\CPaddrline{Boston, MA, 02215}%
\CPaddrline{USA}%
\CPlastname{Zhang, Y. F.}%
\CPcountry{USA}%
\CPzip{02215}%
\CPcity{Boston, MA,}%
\CPinstitute{Astronomy Dept., 725 Commonwealth Ave.}%
\CPstreet{Boston University}%
\CPtelephone{ 617-353-8917}%
\CPnetaddr{ EAST::zhang@buast0.bu.edu}%
\CPnetaddr{ ZHANG@buast0.bu.edu}%
\CPnetaddr{ ZHANG@buasta.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{107}{%
USERS=5 PROGS=0 TASKS=0 BAND=radar
}%
\CPmachine{107}{%
SITE=Boston~U.
CPU=Sun~Clus.~(2*4/60)
AP=none
OS=UNIX~Sun~4.0.3c
DISK=1 MEM=16
PLOT=QMS~810,QMS~PS-820
TV=Sun-4,DECstation
TAPE=1600bpi,6250bpi
AVUSERS=3 AIPS=15JUL90
FRACT=30 VLA=30 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dave Roberts}%
\CPaddrline{Brandeis Univ.}%
\CPaddrline{Physics Dept., 415 South St.}%
\CPaddrline{Waltham, MA, 02254-9110}%
\CPaddrline{USA}%
\CPlastname{Roberts, Dave}%
\CPcountry{USA}%
\CPzip{02254-9110}%
\CPcity{Waltham, MA,}%
\CPinstitute{Physics Dept., 415 South St.}%
\CPstreet{Brandeis Univ.}%
\CPnetaddr{ roberts@brandeis.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{108}{%
USERS=3 PROGS=1 TASKS=1
}%
\CPmachine{108}{%
SITE=Brandeis~U.
CPU=DEC~VAX8650
AP=FPS~5105
OS=VMS~x.x
DISK=1 MEM=32
PLOT=Printronix~xxx,DEC~LN01
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{David Batuski}%
\CPaddrline{Univ. of Maine}%
\CPaddrline{Dept. of Physics and Astronomy, Bennett Hall}%
\CPaddrline{Orono, ME, 04469}%
\CPaddrline{USA}%
\CPlastname{Batuski, David}%
\CPcountry{USA}%
\CPzip{04469}%
\CPcity{Orono, ME,}%
\CPinstitute{Dept. of Physics and Astronomy, Bennett Hall}%
\CPstreet{Univ. of Maine}%
\CPtelephone{ 207-581-1036}%
\CPnetaddr{ batuski@maine.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{109}{%
USERS=3 PROGS=1 TASKS=0
}%
\CPmachine{109}{%
SITE=U.~Maine~Astron.
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.3 MEM=8
PLOT=HP~LaserJet~(w/PS)
TV=Sun-3
TAPE=QIC24
AVUSERS=1 AIPS=15JUL90
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Jack Carmody}%
\CPaddrline{AIPS Manager}%
\CPaddrline{291 High St.}%
\CPaddrline{Nutley, NJ, 07110}%
\CPaddrline{USA}%
\CPlastname{Carmody, Jack}%
\CPcountry{USA}%
\CPzip{07110}%
\CPcity{Nutley, NJ,}%
\CPinstitute{(no institute?)}%
\CPstreet{AIPS Manager, 291 High St.}%
\CPtelephone{ 201-284-2235}%
\CPnetaddr{ dcdwest!vaxf!jcc@sdcs.arpa}%
\CPattribute{CP}%
}%
\CPsitedata{110}{%
USERS=1 PROGS=1 TASKS=2 BAND=IR,other
}%
\CPmachine{110}{%
SITE=Jack~Carmody
CPU=DEC~VAX780
AP=none
OS=VMS~4.4
DISK=2 MEM=10
PLOT=?
TV=?
TAPE=6250bpi
AVUSERS=? AIPS=15OCT85
FRACT=10 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Tony Tyson}%
\CPaddrline{AT\&T Bell Labs.}%
\CPaddrline{1D316}%
\CPaddrline{Murray Hill, NJ, 07974}%
\CPaddrline{USA}%
\CPlastname{Tyson, Tony}%
\CPcountry{USA}%
\CPzip{07974}%
\CPcity{Murray Hill, NJ,}%
\CPinstitute{AT\&T Bell Labs.}%
\CPstreet{1D316}%
\CPtelephone{ 201-582-7278}%
\CPnetaddr{ raw@physics.att.com}%
\CPattribute{CP}%
}%
\CPsitedata{111}{%
USERS=3 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{111}{%
SITE=Bell~Labs.~Murray~Hill
CPU=DEC~VAX785
AP=none
OS=UNIX~4.x
DISK=0.1 MEM=8
PLOT=Imagen~x/yyy
TV=IIS~75
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL85
FRACT=1 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. G. R. Knapp}%
\CPaddrline{Princeton Univ.}%
\CPaddrline{Dept. of Astrophysical Sciences, Peyton Hall}%
\CPaddrline{Princeton, NJ, 08540}%
\CPaddrline{USA}%
\CPlastname{Knapp, Dr. G. R.}%
\CPcountry{USA}%
\CPzip{08540}%
\CPcity{Princeton, NJ,}%
\CPinstitute{Dept. of Astrophysical Sciences, Peyton Hall}%
\CPstreet{Princeton Univ.}%
\CPtelephone{ 609-452-3824}%
\CPnetaddr{ gk@astrovax.princeton.edu}%
\CPattribute{CP}%
}%
\CPsitedata{112}{%
USERS=6 PROGS=1 TASKS=1
}%
\CPmachine{112}{%
SITE=Princeton~U.~Obs.
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~6.2
DISK=0.8 MEM=128
PLOT=Printronix~xxx,Versatec~xxxx,Imagen~8/300
TV=Sun-3~color
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Claire Russell}%
\CPaddrline{Columbia Astrophysics Lab.}%
\CPaddrline{538 West 120th Street}%
\CPaddrline{New York, NY, 10027}%
\CPaddrline{USA}%
\CPlastname{Russell, Claire}%
\CPcountry{USA}%
\CPzip{10027}%
\CPcity{New York, NY,}%
\CPinstitute{Columbia Astrophysics Lab.}%
\CPstreet{538 West 120th Street}%
\CPtelephone{ 212-854-2819}%
\CPnetaddr{ car@cuphyd.bitnet}%
\CPnetaddr{ car@cuphyd.phys.columbia.edu}%
\CPattribute{CP}%
}%
\CPsitedata{113}{%
USERS=10 PROGS=1 TASKS=0 BAND=X-ray,IR,optical
}%
\CPmachine{113}{%
SITE=Columbia~U.~Astron.
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=1 MEM=32
PLOT=Apple~Laserwriter
TV=IIS~75
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT90
FRACT=50 VLA=75 NONRADIO=25
}%
}%

\CP{
\CPaddr{
\CPaddrline{Marc L. Kutner}%
\CPaddrline{Rensselaer Polytechnic Inst.}%
\CPaddrline{Dept. of Physics}%
\CPaddrline{Troy, NY, 12180-3590}%
\CPaddrline{USA}%
\CPlastname{Kutner, Marc L.}%
\CPcountry{USA}%
\CPzip{12180-3590}%
\CPcity{Troy, NY,}%
\CPinstitute{Dept. of Physics}%
\CPstreet{Rensselaer Polytechnic Inst.}%
\CPtelephone{ 518-276-6417}%
\CPnetaddr{ astron@orion.phys.rpi.edu}%
\CPattribute{CP}%
}%
\CPsitedata{114}{%
USERS=10 PROGS=1 TASKS=0 BAND=IR
}%
\CPmachine{114}{%
SITE=Rensselaer~P.~Inst.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.2 MEM=8
PLOT=TI~8250,Omnilaser~(Postscript)
TV=Sun-4~(Sunview)
TAPE=1600bpi,6250bpi,0.25"~cartridge
AVUSERS=2 AIPS=15APR89
FRACT=60 VLA=25 NONRADIO=50
}%
\CPmachine{115}{%
SITE=Rensselaer~P.~Inst.
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~3.5
DISK=0.28 MEM=4
PLOT=TI~2105~(Postscript)
TV=Sun-3
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR89
FRACT=50 VLA=25 NONRADIO=50
}%
}%

\CP{
\CPaddr{
\CPaddrline{Andrew Clegg}%
\CPaddrline{Cornell Univ.}%
\CPaddrline{Space Sciences Bldg.}%
\CPaddrline{Ithaca, NY, 14853}%
\CPaddrline{USA}%
\CPlastname{Clegg, Andrew}%
\CPcountry{USA}%
\CPzip{14853}%
\CPcity{Ithaca, NY,}%
\CPinstitute{Cornell Univ.}%
\CPstreet{Space Sciences Bldg.}%
\CPtelephone{ 607-255-6438}%
\CPnetaddr{ clegg@crnlastr.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{116}{%
USERS=5 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{116}{%
SITE=Cornell~U.
CPU=DEC~VAX750
AP=none
OS=VMS~x.x
DISK=0.1 MEM=5
PLOT=QMS~800
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=1 VLA=33 NONRADIO=33
}%
}%

\CP{
\CPaddr{
\CPaddrline{Frank Briggs}%
\CPaddrline{Univ. of Pittsburgh}%
\CPaddrline{Dept. of Physics, 3941 O'Hara St., 100 Allen Hall}%
\CPaddrline{Pittsburgh, PA, 15260}%
\CPaddrline{USA}%
\CPlastname{Briggs, Frank}%
\CPcountry{USA}%
\CPzip{15260}%
\CPcity{Pittsburgh, PA,}%
\CPinstitute{Dept. of Physics, 3941 O'Hara St., 100 Allen Hall}%
\CPstreet{Univ. of Pittsburgh}%
\CPtelephone{ 412-624-9026}%
\CPnetaddr{ FBRIGGS@VMS.Cis.pitt.edu}%
\CPnetaddr{ FBRIGGS@PITTVMS.bitnet}%
\CPnetaddr{ fbriggs@louie.phyast.pit.edu}%
\CPattribute{CP}%
}%
\CPsitedata{117}{%
USERS=4 PROGS=0 TASKS=0
}%
\CPmachine{117}{%
SITE=U.~Pittsburgh~Astron.
CPU=Sun~Clus.~(5*4/60)
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.7 MEM=12
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT90
FRACT=25 VLA=50 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{James Leous}%
\CPaddrline{Pennsylvania St. Univ.}%
\CPaddrline{Dept. of Astronomy \&\ Astrophysics}%
\CPaddrline{Univ. Park, PA, 16802}%
\CPaddrline{USA}%
\CPlastname{Leous, James}%
\CPcountry{USA}%
\CPzip{16802}%
\CPcity{Univ. Park, PA,}%
\CPinstitute{Dept. of Astronomy \&\ Astrophysics}%
\CPstreet{Pennsylvania St. Univ.}%
\CPtelephone{ 814-865-0418}%
\CPnetaddr{ leous@astro.psu.edu}%
\CPattribute{CP}%
}%
\CPsitedata{118}{%
USERS=6 PROGS=1 TASKS=3 BAND=IR,X-ray
}%
\CPmachine{118}{%
SITE=Penn~State~U.
CPU=Sun~Clus.~(4/65;4/60)
AP=none
OS=UNIX~Sun~4.1,UNIX~Sun~4.0.3c
DISK=1.2 MEM=16
PLOT=Apple~Laserwriter~II
TV=Sun-4~color
TAPE=800bpi,1600bpi,6250bpi,QIC24,Exabyte,DAT
AVUSERS=1 AIPS=15JAN89
FRACT=20 VLA=80 NONRADIO=20
}%
\CPmachine{119}{%
SITE=Penn~State~U.
CPU=Sun~Clus.~(4/110;4/20)
AP=none
OS=UNIX~Sun~4.0.1,UNIX~Sun~4.1
DISK=1 MEM=8
PLOT=Apple~Laserwriter~II
TV=Sun-4/110
TAPE=800bpi,1600bpi,6250bpi,QIC24,Exabyte,DAT
AVUSERS=2.5 AIPS=15JAN89
FRACT=60 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Lyle Hoffman}%
\CPaddrline{Lafayette College}%
\CPaddrline{Dept. of Physics}%
\CPaddrline{Easton, PA, 18042-1782}%
\CPaddrline{USA}%
\CPlastname{Hoffman, Lyle}%
\CPcountry{USA}%
\CPzip{18042-1782}%
\CPcity{Easton, PA,}%
\CPinstitute{Dept. of Physics}%
\CPstreet{Lafayette College}%
\CPtelephone{ 215-250-5211}%
\CPnetaddr{ hoffmang@lafayett.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{120}{%
USERS=3 PROGS=0.5 TASKS=2
}%
\CPmachine{120}{%
SITE=Lafayette~Coll.
CPU=Sun-4/65
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=6.9
PLOT=HP~LaserJet~II
TV=Sun-4
TAPE=QIC24,Exabyte
AVUSERS=1 AIPS=15JUL90
FRACT=30 VLA=100 NONRADIO=0
}%
\CPmachine{121}{%
SITE=Lafayette~Coll.
CPU=DEC~VAX750
AP=none
OS=VMS~5.1-1
DISK=0.5 MEM=14
PLOT=HP~LaserJet~II
TV=Amiga~500
TAPE=1600bpi
AVUSERS=1 AIPS=15OCT88
FRACT=30 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Ralph Gaume}%
\CPaddrline{Naval Research Lab}%
\CPaddrline{Code 4213.6, 4555 Overlook Ave. SW}%
\CPaddrline{Washington, DC, 20375-5000}%
\CPaddrline{USA}%
\CPlastname{Gaume, Ralph}%
\CPcountry{USA}%
\CPzip{20375-5000}%
\CPcity{Washington, DC,}%
\CPinstitute{Naval Research Lab}%
\CPstreet{Code 4213.6, 4555 Overlook Ave. SW}%
\CPtelephone{ 202-767-0670}%
\CPnetaddr{ gaume@rira1.nrl.navy.mil}%
\CPattribute{CP}%
}%
\CPsitedata{122}{%
USERS=15 PROGS=4 TASKS=7 BAND=IR,optical
}%
\CPmachine{122}{%
SITE=Nav.~Res.~Lab.~Radio
CPU=Alliant~FX/8-3CE
AP=Vector~Regs
OS=UNIX~Concentrix~5.5.02
DISK=4 MEM=48
PLOT=QMS~Talaris~1590-T
TV=Sun-4~(X-Windows),Alliant~(X-Windows),PC-386~(X-Windows)
TAPE=1600bpi,6250bpi
AVUSERS=4 AIPS=15APR90
FRACT=80 VLA=80 NONRADIO=20
}%
\CPmachine{123}{%
SITE=Nav.~Res.~Lab.~X-Ray
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3-1
DISK=0.3 MEM=8
PLOT=QMS~Talaris~1590-T
TV=Gould~IP9000
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=0.25 AIPS=15JAN89
FRACT=10 VLA=25 NONRADIO=75
}%
}%

\CP{
\CPaddr{
\CPaddrline{David Florkowski}%
\CPaddrline{U.S. Naval Obs.}%
\CPaddrline{34th St. \&\ Mass. Ave., N.W.}%
\CPaddrline{Washington, DC, 20392}%
\CPaddrline{USA}%
\CPlastname{Florkowski, David}%
\CPcountry{USA}%
\CPzip{20392}%
\CPcity{Washington, DC,}%
\CPinstitute{U.S. Naval Obs.}%
\CPstreet{34th St. \&\ Mass. Ave., N.W.}%
\CPtelephone{ 202-653-0067}%
\CPnetaddr{ drf@phobos.usno.navy.mil}%
\CPattribute{CP}%
}%
\CPsitedata{124}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{124}{%
SITE=U.S.~Naval~Obs.
CPU=DEC~VAX8530
AP=FPS~5300
OS=VMS~5.4
DISK=0.6 MEM=32
PLOT=QMS~Talaris~1590
TV=IIS~IVAS
TAPE=1600bpi,6250bpi,8mm
AVUSERS=1 AIPS=15APR89
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. George H. Kaplan}%
\CPaddrline{U.S. Naval Obs.}%
\CPaddrline{34th and Mass. Ave. NW}%
\CPaddrline{Washington, DC, 20392}%
\CPaddrline{USA}%
\CPlastname{Kaplan, Dr. George H.}%
\CPcountry{USA}%
\CPzip{20392}%
\CPcity{Washington, DC,}%
\CPinstitute{U.S. Naval Obs.}%
\CPstreet{34th and Mass. Ave. NW}%
\CPtelephone{ 202-653-1556}%
\CPnetaddr{ GHK@PHOBOS.USNO.NAVY.MIL}%
\CPattribute{CP}%
}%
\CPsitedata{125}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{125}{%
SITE=U.S.~Naval~Obs.
CPU=DEC~VAX8530
AP=FPS~5305
OS=VMS~4.7
DISK=2.1 MEM=32
PLOT=QMS~Talaris~1590-T
TV=IIS~IVAS,TEK~4014
TAPE=1600bpi,3200bpi,6250bpi
AVUSERS=1 AIPS=15APR89
FRACT=2 VLA=95 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Ramesh Sinha}%
\CPaddrline{STX (ST Systems Corp.)}%
\CPaddrline{7104 Kurth Lane}%
\CPaddrline{Seabrook, MD, 20706}%
\CPaddrline{USA}%
\CPlastname{Sinha, Ramesh}%
\CPcountry{USA}%
\CPzip{20706}%
\CPcity{Seabrook, MD,}%
\CPinstitute{STX (ST Systems Corp.)}%
\CPstreet{7104 Kurth Lane}%
\CPtelephone{ 301-794-5203}%
\CPnetaddr{ xrrps@csdr.span}%
\CPattribute{CP}%
}%
\CPsitedata{126}{%
USERS=10 PROGS=3 TASKS=10 BAND=IR
}%
\CPmachine{126}{%
SITE=ST~Systems~Corp.
CPU=DEC~VAX8200
AP=none
OS=VMS~x.x
DISK=0.05 MEM=20
PLOT=QMS~Talaris~800,Citoh
TV=IIS~IVAS
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Jan M. Hollis}%
\CPaddrline{NASA-Goddard Space Flight Ctr.}%
\CPaddrline{Code 930, Bldg. 28, Rm W205J}%
\CPaddrline{Greenbelt, MD, 20771}%
\CPaddrline{USA}%
\CPlastname{Hollis, Dr. Jan M.}%
\CPcountry{USA}%
\CPzip{20771}%
\CPcity{Greenbelt, MD,}%
\CPinstitute{NASA-Goddard Space Flight Ctr.}%
\CPstreet{Code 930, Bldg. 28, Rm W205J}%
\CPtelephone{ 301-286-7591}%
\CPnetaddr{ sdcdcl::hollis}%
\CPnetaddr{ hollis@dftbit.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{127}{%
USERS=15 PROGS=3 TASKS=3 BAND=optical
}%
\CPmachine{127}{%
SITE=NASA-GSFC~SDCDCL
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~x.x
DISK=0.981 MEM=12
PLOT=?
TV=Sun-3
TAPE=QIC11,QIC24,Exabyte
AVUSERS=1 AIPS=15JUL88
FRACT=80 VLA=80 NONRADIO=20
}%
}%

\CP{
\CPaddr{
\CPaddrline{Susan Neff}%
\CPaddrline{NASA-Goddard Space Flight Ctr.}%
\CPaddrline{Code 684.1, Lab. for Astronomy \&\ Solar Physics}%
\CPaddrline{Greenbelt, MD, 20771}%
\CPaddrline{USA}%
\CPlastname{Neff, Susan}%
\CPcountry{USA}%
\CPzip{20771}%
\CPcity{Greenbelt, MD,}%
\CPinstitute{Code 684.1, Lab. for Astronomy \&\ Solar Physics}%
\CPstreet{NASA-Goddard Space Flight Ctr.}%
\CPtelephone{ 301-286-5137}%
\CPnetaddr{ uit::neff}%
\CPnetaddr{ 6170::neff}%
\CPnetaddr{ neff@stars.gsfc.nasa.gov}%
\CPattribute{CP}%
}%
\CPsitedata{128}{%
USERS=40 PROGS=0.1 TASKS=4 BAND=IR,optical,UV
}%
\CPmachine{128}{%
SITE=NASA-GSFC~Ast/Solar
CPU=DEC~VAX6410
AP=FPS~5305
OS=VMS~5.3
DISK=1 MEM=32
PLOT=QMS~800,Postscript
TV=Gould~8500
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1.5 AIPS=15OCT88
FRACT=10 VLA=30 NONRADIO=70
}%
\CPmachine{129}{%
SITE=NASA-GSFC~Ast/Solar
CPU=DEC~VAX8350
AP=none
OS=VMS~5.3
DISK=? MEM=8
PLOT=QMS~2400
TV=Gould~9500
TAPE=6250bpi,Exabyte
AVUSERS=0.1 AIPS=15OCT88
FRACT=1 VLA=20 NONRADIO=80
}%
\CPmachine{130}{%
SITE=NASA-GSFC~Ast/Solar
CPU=DEC~VAX8350
AP=none
OS=VMS~5.3
DISK=? MEM=12
PLOT=QMS~Talaris~800
TV=Gould~8500
TAPE=800bpi,1600bpi,6250bpi,Exabyte
AVUSERS=0.1 AIPS=15OCT88
FRACT=1 VLA=0 NONRADIO=100
}%
\CPmachine{131}{%
SITE=NASA-GSFC~Ast/Solar
CPU=DEC~VAX750
AP=none
OS=VMS~5.3
DISK=1 MEM=8
PLOT=Versatec~V80,QMS~800
TV=Gould~8500,IIS~75
TAPE=800bpi,1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT88
FRACT=30 VLA=10 NONRADIO=90
}%
\CPmachine{132}{%
SITE=NASA-GSFC~Ast/Solar
CPU=DEC~VAX750
AP=none
OS=VMS~5.3
DISK=? MEM=15
PLOT=QMS~800,Postscript
TV=Gould~8500
TAPE=800bpi,1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT88
FRACT=20 VLA=40 NONRADIO=60
}%
}%

\CP{
\CPaddr{
\CPaddrline{Jeffrey A. Pedelty}%
\CPaddrline{NASA Goddard Space Flight Ctr.}%
\CPaddrline{SISC Computer Facility, Code 936}%
\CPaddrline{Greenbelt, MD, 20771}%
\CPaddrline{USA}%
\CPlastname{Pedelty, Jeffrey A.}%
\CPcountry{USA}%
\CPzip{20771}%
\CPcity{Greenbelt, MD,}%
\CPinstitute{NASA Goddard Space Flight Ctr.}%
\CPstreet{SISC Computer Facility, Code 936}%
\CPtelephone{ 301-286-3065}%
\CPnetaddr{ pedelty@amarna.gsfc.nasa.gov}%
\CPnetaddr{ SDCDCL::PEDELTY}%
\CPattribute{CP}%
}%
\CPsitedata{133}{%
USERS=15 PROGS=1 TASKS=2 BAND=IR,optical,UV
}%
\CPmachine{133}{%
SITE=NASA/GSFC~SDCD
CPU=DEC~Clus.~(6400;8820;8250;2*780;3*3200;3100;2000)
AP=none
OS=VMS~5.2
DISK=1.73 MEM=160
PLOT=QMS~Talaris~1590-T,Apple~Laserwriter
TV=IIS~75,IIS~IVAS,X-Windows
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=1 VLA=50 NONRADIO=50
}%
}%

\CP{
\CPaddr{
\CPaddrline{Harry Payne}%
\CPaddrline{Space Telescope Science Inst.}%
\CPaddrline{3700 San Martin Dr.}%
\CPaddrline{Baltimore, MD, 21218}%
\CPaddrline{USA}%
\CPlastname{Payne, Harry}%
\CPcountry{USA}%
\CPzip{21218}%
\CPcity{Baltimore, MD,}%
\CPinstitute{Space Telescope Science Inst.}%
\CPstreet{3700 San Martin Dr.}%
\CPtelephone{ 301-338-4539}%
\CPnetaddr{ payne@scivax.span}%
\CPnetaddr{ aips@scivax.span}%
\CPnetaddr{ payne@stsci.edu}%
\CPattribute{CP}%
}%
\CPsitedata{134}{%
USERS=12 PROGS=0 TASKS=0
}%
\CPmachine{134}{%
SITE=S.~T.~Sci.~Inst.
CPU=Sun~Clus.~(4/490~server;4/280;10*4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=0.4 MEM=24
PLOT=Apple~Laserwriter~II~NTX
TV=Sun-4~(Sunview),Sun-4~(Openlook~(XAS))
TAPE=6250bpi,QIC24,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=94 NONRADIO=6
}%
\CPmachine{135}{%
SITE=S.~T.~Sci.~Inst.
CPU=Sun~Clus.~(4/490~server;6*4/60)
AP=none
OS=UNIX~Sun~4.1
DISK=0.4 MEM=32
PLOT=Apple~Laserwriter~II~NTX
TV=Sun-4~(Sunview),Sun-4~(Openlook~(XAS))
TAPE=6250bpi,QIC24,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=94 NONRADIO=6
}%
\CPmachine{136}{%
SITE=S.~T.~Sci.~Inst.
CPU=Sun~Clus.~(4/490~server;3/280;8*3/60)
AP=none
OS=UNIX~Sun~4.1
DISK=0.4 MEM=32
PLOT=Apple~Laserwriter~II~NTX
TV=Sun-4~(Sunview),Sun-4~(Openlook~(XAS))
TAPE=6250bpi,QIC24,Exabyte
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=94 NONRADIO=6
}%
\CPmachine{137}{%
SITE=S.~T.~Sci.~Inst.
CPU=DEC~Clus.~(8800)
AP=FPS~5205
OS=VMS~5.3
DISK=0.4 MEM=32
PLOT=QMS~800,QMS~1200,DEC~LN03R
TV=Gould~IP8500
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=70 NONRADIO=30
}%
\CPmachine{138}{%
SITE=S.~T.~Sci.~Inst.
CPU=DEC~VAX785
AP=FPS~5205
OS=VMS~5.3
DISK=0.4 MEM=32
PLOT=QMS~800,QMS~1200,DEC~LN03R
TV=Gould~IP8500
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=70 NONRADIO=30
}%
}%

\CP{
\CPaddr{
\CPaddrline{Gregory S. Hennessy}%
\CPaddrline{Box 3818, Univ. Station}%
\CPaddrline{Astronomy Dept., Univ. of Virginia}%
\CPaddrline{Charlottesville, VA, 22903}%
\CPaddrline{USA}%
\CPlastname{Hennessy, Gregory S.}%
\CPcountry{USA}%
\CPzip{22903}%
\CPcity{Charlottesville, VA,}%
\CPinstitute{Astronomy Dept., Univ. of Virginia}%
\CPstreet{Box 3818, Univ. Station}%
\CPtelephone{ 804-924-7494}%
\CPnetaddr{ gsh7w@virginia.edu}%
\CPattribute{CP}%
}%
\CPsitedata{139}{%
USERS=10 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{139}{%
SITE=U.~Virginia
CPU=DEC~uVAX~II
AP=none
OS=VMS~x.x
DISK=0.7 MEM=13
PLOT=QMS~800C,Pericom~Video~Printer
TV=IIS~IVAS
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{John Simonetti}%
\CPaddrline{Physics Dept.}%
\CPaddrline{Virginia Tech}%
\CPaddrline{Blacksburg, VA, 24061}%
\CPaddrline{USA}%
\CPlastname{Simonetti, John}%
\CPcountry{USA}%
\CPzip{24061}%
\CPcity{Blacksburg, VA,}%
\CPinstitute{Physics Dept.}%
\CPstreet{Virginia Tech}%
\CPtelephone{ 703-231-8740}%
\CPnetaddr{ jhs@astro.phys.vt.edu}%
\CPattribute{CP}%
}%
\CPsitedata{140}{%
USERS=4 PROGS=1 TASKS=0 BAND=IR,X-ray
}%
\CPmachine{140}{%
SITE=Va.~Tech.~Astron.
CPU=Sun-3/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.4 MEM=4
PLOT=NEC~LC890~(Postscript)
TV=Sun-3
TAPE=QIC24,Exabyte
AVUSERS=1 AIPS=15JAN90
FRACT=40 VLA=50 NONRADIO=20
}%
}%

\CP{
\CPaddr{
\CPaddrline{Wayne Christiansen}%
\CPaddrline{Univ. of North Carolina}%
\CPaddrline{Dept. of Phys. \&\ Astronomy, 260 Phillips Hall 039A}%
\CPaddrline{Chapel Hill, NC, 27514}%
\CPaddrline{USA}%
\CPlastname{Christiansen, Wayne}%
\CPcountry{USA}%
\CPzip{27514}%
\CPcity{Chapel Hill, NC,}%
\CPinstitute{Dept. of Phys. \&\ Astronomy, 260 Phillips Hall 039A}%
\CPstreet{Univ. of North Carolina}%
\CPtelephone{ 919-962-3022}%
\CPnetaddr{ waco@unc.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{141}{%
USERS=3 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{141}{%
SITE=U.~N.~Carolina
CPU=Convex~C-240
AP=Vector~Regs
OS=UNIX~Convex~9.0
DISK=7 MEM=128
PLOT=?
TV=Sun-3/260
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Stephen P. Reynolds}%
\CPaddrline{North Carolina State Univ.}%
\CPaddrline{Dept. of Physics, Box 8202, 104 Cox Hall}%
\CPaddrline{Raleigh, NC, 27695-8202}%
\CPaddrline{USA}%
\CPlastname{Reynolds, Stephen P.}%
\CPcountry{USA}%
\CPzip{27695-8202}%
\CPcity{Raleigh, NC,}%
\CPinstitute{Dept. of Physics, Box 8202, 104 Cox Hall}%
\CPstreet{North Carolina State Univ.}%
\CPtelephone{ 919-737-7751}%
\CPnetaddr{ reynolds@ncsuphys.bitnet}%
\CPnetaddr{ aips@ncsuphys.bitnet}%
\CPnetaddr{ reynolds@pymvax.ncsu.edu}%
\CPattribute{CP}%
}%
\CPsitedata{142}{%
USERS=3 PROGS=0.5 TASKS=2 BAND=other,X-ray
}%
\CPmachine{142}{%
SITE=NC~St.~U.~Phys.
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=0.8 MEM=10
PLOT=DEC~LN03+~(Postscript)
TV=IIS~IVAS
TAPE=1600bpi,TK50
AVUSERS=1 AIPS=15OCT90
FRACT=15 VLA=20 NONRADIO=80
}%
}%

\CP{
\CPaddr{
\CPaddrline{Charles A. Taylor}%
\CPaddrline{Univ. of Florida}%
\CPaddrline{Dept. of Astronomy, 212 Space Sciences Bldg.}%
\CPaddrline{Gainesville, FL, 32611}%
\CPaddrline{USA}%
\CPlastname{Taylor, Charles A.}%
\CPcountry{USA}%
\CPzip{32611}%
\CPcity{Gainesville, FL,}%
\CPinstitute{Dept. of Astronomy, 212 Space Sciences Bldg.}%
\CPstreet{Univ. of Florida}%
\CPtelephone{ 904-392-2050/2052}%
\CPnetaddr{ taylor@astro.ufl.edu}%
\CPattribute{CP}%
}%
\CPsitedata{143}{%
USERS=4 PROGS=0 TASKS=0
}%
\CPmachine{143}{%
SITE=U.~Florida
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~x.x
DISK=0.6 MEM=12
PLOT=Postscript
TV=Sun-4
TAPE=QIC24
AVUSERS=1 AIPS=15JUL90
FRACT=50 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Jonathan Campbell}%
\CPaddrline{NASA-Marshall Space Flight Ctr.}%
\CPaddrline{ES52}%
\CPaddrline{Huntsville, AL, 35812}%
\CPaddrline{USA}%
\CPlastname{Campbell, Jonathan}%
\CPcountry{USA}%
\CPzip{35812}%
\CPcity{Huntsville, AL,}%
\CPinstitute{NASA-Marshall Space Flight Ctr.}%
\CPstreet{ES52}%
\CPtelephone{ 205-544-7076}%
\CPnetaddr{ SAM::CAMPBELL}%
\CPattribute{CP}%
}%
\CPsitedata{144}{%
USERS=6 PROGS=2 TASKS=2 BAND=X-ray
}%
\CPmachine{144}{%
SITE=NASA-Marshall~SFC
CPU=DEC~Clus.~(780;3100)
AP=none
OS=VMS~5.4
DISK=0.05 MEM=16
PLOT=QMS~800,QMS~810
TV=DEC~(X-Windows~planned)
TAPE=6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=0.1 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{James Trice}%
\CPaddrline{Univ. of Kentucky}%
\CPaddrline{Dept. of Physics \&\ Astronomy, Room 177}%
\CPaddrline{Lexington, KY, 40506}%
\CPaddrline{USA}%
\CPlastname{Trice, James}%
\CPcountry{USA}%
\CPzip{40506}%
\CPcity{Lexington, KY,}%
\CPinstitute{Dept. of Physics \&\ Astronomy, Room 177}%
\CPstreet{Univ. of Kentucky}%
\CPtelephone{ 606-257-1995}%
\CPnetaddr{ trice@asta.pa.uky}%
\CPnetaddr{ phy105@ukcc.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{145}{%
USERS=1 PROGS=0 TASKS=0
}%
\CPmachine{145}{%
SITE=U.~Ky.~Astron.
CPU=Sun-4/280
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=8
PLOT=Apple~Laserwriter~II
TV=Sun-4
TAPE=1600bpi,6250bpi
AVUSERS=? AIPS=15OCT90
FRACT=1 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Michele Kaufman}%
\CPaddrline{Ohio State University}%
\CPaddrline{Physics Dept., 174 W. 18th Ave.}%
\CPaddrline{Columbus, OH, 43210}%
\CPaddrline{USA}%
\CPlastname{Kaufman, Michele}%
\CPcountry{USA}%
\CPzip{43210}%
\CPcity{Columbus, OH,}%
\CPinstitute{Physics Dept., 174 W. 18th Ave.}%
\CPstreet{Ohio State University}%
\CPtelephone{ 614-292-4464}%
\CPnetaddr{ rallis@ohstpy.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{146}{%
USERS=1 PROGS=0 TASKS=0 BAND=UV,optical
}%
\CPmachine{146}{%
SITE=Ohio~St.~U.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.9 MEM=8
PLOT=Apple~Laserwriter~II
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=100 VLA=70 NONRADIO=30
}%
\CPmachine{147}{%
SITE=Ohio~St.~U.
CPU=DEC~VAX8600
AP=none
OS=VMS~5.3-1
DISK=0.1 MEM=12
PLOT=QMS~Talaris~1590-T
TV=?
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JAN87
FRACT=3 VLA=70 NONRADIO=30
}%
}%

\CP{
\CPaddr{
\CPaddrline{George Turner}%
\CPaddrline{Indiana Univ.}%
\CPaddrline{Astronomy Dept., Swain West 319}%
\CPaddrline{Bloomington, IN, 47405}%
\CPaddrline{USA}%
\CPlastname{Turner, George}%
\CPcountry{USA}%
\CPzip{47405}%
\CPcity{Bloomington, IN,}%
\CPinstitute{Astronomy Dept., Swain West 319}%
\CPstreet{Indiana Univ.}%
\CPtelephone{ 812-855-6782}%
\CPnetaddr{ turner@ucs.indiana.edu}%
\CPnetaddr{ turnerg@iubacs.bitnet}%
\CPnetaddr{ 47310::sol.turner}%
\CPattribute{CP}%
}%
\CPsitedata{148}{%
USERS=2 PROGS=1 TASKS=0
}%
\CPmachine{148}{%
SITE=Indiana~U.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.0.3
DISK=0.66 MEM=16
PLOT=Postscript,HP~GL,Zeta
TV=Sun-4~(X-Windows)
TAPE=QIC24
AVUSERS=1 AIPS=15JUL90
FRACT=5 VLA=5 NONRADIO=0
}%
\CPmachine{149}{%
SITE=Indiana~U.
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.4
DISK=0.7 MEM=9
PLOT=HP~7475,Postscript
TV=IAC~QD512,Data~Translation~AT2651,X-Windows
TAPE=QIC24
AVUSERS=1 AIPS=15JUL90
FRACT=5 VLA=5 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Hugh D. Aller}%
\CPaddrline{Univ. of Michigan}%
\CPaddrline{Dept. of Astronomy, 810 Dennison Bldg.}%
\CPaddrline{Ann Arbor, MI, 48109-1090}%
\CPaddrline{USA}%
\CPlastname{Aller, Hugh D.}%
\CPcountry{USA}%
\CPzip{48109-1090}%
\CPcity{Ann Arbor, MI,}%
\CPinstitute{Dept. of Astronomy, 810 Dennison Bldg.}%
\CPstreet{Univ. of Michigan}%
\CPtelephone{ 313-764-3466}%
\CPnetaddr{ hugh@umast.span}%
\CPnetaddr{ hugh\_d.\_aller@umichum.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{150}{%
USERS=10 PROGS=1 TASKS=10 BAND=IR,optical
}%
\CPmachine{150}{%
SITE=U.~Michigan
CPU=DEC~VAX750
AP=FPS~5205
OS=VMS~x.x
DISK=1.3 MEM=8
PLOT=Imagen~Innovator,HP~7550A
TV=RCI~TRAPIX~5500
TAPE=?
AVUSERS=? AIPS=?
FRACT=20 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Susan M. Simkin}%
\CPaddrline{Michigan St. Univ.}%
\CPaddrline{Dept. Physics \&\ Astronomy}%
\CPaddrline{E. Lansing, MI, 48824-1116}%
\CPaddrline{USA}%
\CPlastname{Simkin, Susan M.}%
\CPcountry{USA}%
\CPzip{48824-1116}%
\CPcity{E. Lansing, MI,}%
\CPinstitute{Dept. Physics \&\ Astronomy}%
\CPstreet{Michigan St. Univ.}%
\CPtelephone{ 517-353-4540 (leave a message)}%
\CPnetaddr{ simkin@msupa.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{151}{%
USERS=20 PROGS=0.5 TASKS=10
}%
\CPmachine{151}{%
SITE=Michigan~St.~U.
CPU=DEC~VAX750
AP=none
OS=VMS~x.x
DISK=1.4 MEM=8
PLOT=Printronix~xxx,QMS~Talaris~xxx
TV=Grinnell~GMRxxx
TAPE=?
AVUSERS=? AIPS=?
FRACT=2 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{John P. Basart}%
\CPaddrline{Iowa St. Univ.}%
\CPaddrline{Coover Hall}%
\CPaddrline{Ames, IA, 50011}%
\CPaddrline{USA}%
\CPlastname{Basart, John P.}%
\CPcountry{USA}%
\CPzip{50011}%
\CPcity{Ames, IA,}%
\CPinstitute{Iowa St. Univ.}%
\CPstreet{Coover Hall}%
\CPtelephone{ 515-294-4955}%
\CPnetaddr{ n1jpb@isumvs.iastate.edu}%
\CPattribute{CP}%
}%
\CPsitedata{152}{%
USERS=7 PROGS=1 TASKS=6 BAND=X-ray
}%
\CPmachine{152}{%
SITE=Iowa~St.~U.
CPU=Stellar~1000
AP=none
OS=UNIX~Stellix~System~V
DISK=0.5 MEM=32
PLOT=?
TV=Stellar
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=75 NONRADIO=25
}%
}%

\CP{
\CPaddr{
\CPaddrline{Robert Mutel}%
\CPaddrline{Univ. of Iowa}%
\CPaddrline{Dept. of Physics \&\ Astronomy}%
\CPaddrline{Iowa City, IA, 52242}%
\CPaddrline{USA}%
\CPlastname{Mutel, Robert}%
\CPcountry{USA}%
\CPzip{52242}%
\CPcity{Iowa City, IA,}%
\CPinstitute{Dept. of Physics \&\ Astronomy}%
\CPstreet{Univ. of Iowa}%
\CPtelephone{ 319-335-1950}%
\CPtelex{ TWX 910-525-1398}%
\CPnetaddr{ rlm@iowa.physics.uiowa.edu}%
\CPnetaddr{ iowa::rlm = 7232::rlm (7.64)}%
\CPattribute{CP}%
}%
\CPsitedata{153}{%
USERS=10 PROGS=2 TASKS=4 BAND=other,optical
}%
\CPmachine{153}{%
SITE=U.~Iowa
CPU=Sun~Clus.~(2*4/40)
AP=none
OS=UNIX~Sun~4.1
DISK=1.5 MEM=24
PLOT=HP~LaserJet~II
TV=Sun-4~(X-Windows)
TAPE=QIC24
AVUSERS=1 AIPS=15JUL90
FRACT=35 VLA=65 NONRADIO=0
}%
\CPmachine{154}{%
SITE=U.~Iowa
CPU=DEC~uVAX3200
AP=none
OS=VMS~5.3-1
DISK=1.5 MEM=16
PLOT=HP~LaserJet~II
TV=DEC~(X-Windows)
TAPE=TK50,Exabyte
AVUSERS=2 AIPS=15JUL90
FRACT=50 VLA=75 NONRADIO=5
}%
\CPmachine{155}{%
SITE=U.~Iowa
CPU=DEC~VAX4000
AP=none
OS=VMS~5.3
DISK=0.4 MEM=24
PLOT=HP~7475,HP~LaserJet~II
TV=IIS~75
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=15 VLA=90 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Edward B. Churchwell}%
\CPaddrline{Univ. of Wisconsin}%
\CPaddrline{Washburn Obs., 475 N. Charter Street}%
\CPaddrline{Madison, WI, 53706}%
\CPaddrline{USA}%
\CPlastname{Churchwell, Edward B.}%
\CPcountry{USA}%
\CPzip{53706}%
\CPcity{Madison, WI,}%
\CPinstitute{Washburn Obs., 475 N. Charter Street}%
\CPstreet{Univ. of Wisconsin}%
\CPtelephone{ 608-262-7857}%
\CPnetaddr{ madraf::churchwell}%
\CPnetaddr{ churchwell\%madraf.decnet@vms.macc.wisc.edu}%
\CPnetaddr{ churchwell@madraf.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{156}{%
USERS=3 PROGS=0 TASKS=0 BAND=optical,UV,X-ray
}%
\CPmachine{156}{%
SITE=U.~Wisc.~Washburn~Obs.
CPU=DEC~VAX780
AP=none
OS=VMS~5.3-1
DISK=1 MEM=32
PLOT=Printronix~300,DEC~LN03+,DEC~LN03R
TV=Grinnell~GMR270,DEC~VAXstation~3100
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=40 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dan Schultz}%
\CPaddrline{Space Phys. Lab., Dept. of Physics}%
\CPaddrline{U. of Wisconsin, 1150 University Ave.}%
\CPaddrline{Madison, WI, 53706}%
\CPaddrline{USA}%
\CPlastname{Schultz, Dan}%
\CPcountry{USA}%
\CPzip{53706}%
\CPcity{Madison, WI,}%
\CPinstitute{U. of Wisconsin, 1150 University Ave.}%
\CPstreet{Space Phys. Lab., Dept. of Physics}%
\CPnetaddr{ 47413::dan}%
\CPnetaddr{ dan\%wisp.decnet@vms.macc.isc.edu}%
\CPattribute{CP}%
}%
\CPsitedata{157}{%
USERS=10 PROGS=3 TASKS=30 BAND=optical,X-ray
}%
\CPmachine{157}{%
SITE=U.~Wisc.~Space~Phys.
CPU=DEC~uVAX~II
AP=none
OS=VMS~4.5
DISK=0.37 MEM=5
PLOT=?
TV=Peritek~VCK
TAPE=6250bpi
AVUSERS=1 AIPS=15OCT89
FRACT=95 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Larry Rudnick}%
\CPaddrline{Univ. of Minnesota}%
\CPaddrline{Dept. of Astronomy, 116 Church St. S.E.}%
\CPaddrline{Minneapolis, MN, 55455}%
\CPaddrline{USA}%
\CPlastname{Rudnick, Larry}%
\CPcountry{USA}%
\CPzip{55455}%
\CPcity{Minneapolis, MN,}%
\CPinstitute{Dept. of Astronomy, 116 Church St. S.E.}%
\CPstreet{Univ. of Minnesota}%
\CPtelephone{ 612-624-3396}%
\CPnetaddr{ larry@ast1.spd.umn.edu}%
\CPattribute{CP}%
}%
\CPsitedata{158}{%
USERS=15 PROGS=4 TASKS=5 BAND=IR
}%
\CPmachine{158}{%
SITE=U.~Minnesota
CPU=Cray~X-MP/4
AP=Vector~Regs
OS=UNIX~UniCOS~5.1
DISK=2 MEM=16
PLOT=Apple~Laserwriter
TV=Sun
TAPE=6250bpi,Exabyte
AVUSERS=2 AIPS=15JAN87
FRACT=0.2 VLA=100 NONRADIO=0
}%
\CPmachine{159}{%
SITE=U.~Minnesota
CPU=Sun~Clus.~(2*3/60;2*4/60)
AP=none
OS=UNIX~Sun~4.0.3c
DISK=2.5 MEM=16
PLOT=Apple~Laserwriter
TV=Sun
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=0.5 AIPS=15OCT89
FRACT=70 VLA=80 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Farhad Yusef-Zadeh}%
\CPaddrline{Northwestern Univ.}%
\CPaddrline{Dept. of Physics \&\ Astronomy, Dearborn Obs., 2145}%
\CPaddrline{Evanston, IL, 60201}%
\CPaddrline{USA}%
\CPlastname{Yusef-Zadeh, Farhad}%
\CPcountry{USA}%
\CPzip{60201}%
\CPcity{Evanston, IL,}%
\CPinstitute{Dept. of Physics \&\ Astronomy, Dearborn Obs., 2145}%
\CPstreet{Northwestern Univ.}%
\CPtelephone{ 708-491-8147}%
\CPnetaddr{ 11340::zadeh}%
\CPnetaddr{ zadeh@ossenu.astro.nwu.edu}%
\CPattribute{CP}%
}%
\CPsitedata{160}{%
USERS=6 PROGS=0 TASKS=0 BAND=IR,optical
}%
\CPmachine{160}{%
SITE=Northwestern~U.
CPU=Sun-4/40
AP=none
OS=UNIX~Sun~4.0.3
DISK=1.4 MEM=8
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=99 VLA=100 NONRADIO=0
}%
\CPmachine{161}{%
SITE=Northwestern~U.
CPU=Sun-4/280S
AP=none
OS=UNIX~Sun~4.0.3
DISK=1 MEM=8
PLOT=Apple~Laserwriter
TV=Sun-4
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=99 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{S. Sukumar}%
\CPaddrline{Univ. of Illinois}%
\CPaddrline{210 Astronomy Bldg., 1002 West Green St.}%
\CPaddrline{Urbana, IL, 61801}%
\CPaddrline{USA}%
\CPlastname{Sukumar, S.}%
\CPcountry{USA}%
\CPzip{61801}%
\CPcity{Urbana, IL,}%
\CPinstitute{Univ. of Illinois}%
\CPstreet{210 Astronomy Bldg., 1002 West Green St.}%
\CPtelephone{ 217-244-3664}%
\CPnetaddr{ sukumar@ncsa.uiuc.edu}%
\CPnetaddr{ sukumar@rigel.astro.uiuc.edu}%
\CPattribute{CP}%
}%
\CPsitedata{162}{%
USERS=15 PROGS=0 TASKS=3 BAND=optical,IR
}%
\CPmachine{162}{%
SITE=U.~Illinois~CSO
CPU=Convex~C-210
AP=Vector~Regs
OS=UNIX~Convex~8.1
DISK=0.4 MEM=128
PLOT=Apple~Laserwriter
TV=none
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=2 VLA=80 NONRADIO=20
}%
\CPmachine{163}{%
SITE=U.~Illinois~Astron.
CPU=Sun-4/490
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=256
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=0.25 AIPS=15OCT90
FRACT=10 VLA=80 NONRADIO=20
}%
\CPmachine{164}{%
SITE=U.~Illinois~Astron.
CPU=Sun-3/180
AP=none
OS=UNIX~Sun~x.x
DISK=0.4 MEM=12
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15OCT89
FRACT=80 VLA=80 NONRADIO=20
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Michael Vannier}%
\CPaddrline{Mallinckrodt Inst. of Radiology}%
\CPaddrline{510 S. Kings Hwy.}%
\CPaddrline{St. Louis, MO, 63110}%
\CPaddrline{USA}%
\CPlastname{Vannier, Dr. Michael}%
\CPcountry{USA}%
\CPzip{63110}%
\CPcity{St. Louis, MO,}%
\CPinstitute{Mallinckrodt Inst. of Radiology}%
\CPstreet{510 S. Kings Hwy.}%
\CPtelephone{ 314-362-8467}%
\CPnetaddr{ mvannier@davey.wustl.edu}%
\CPattribute{CP}%
}%
\CPsitedata{165}{%
USERS=3 PROGS=1 TASKS=0 BAND=X-ray,biomed
}%
\CPmachine{165}{%
SITE=Mallinckrodt~Inst.~Radiology
CPU=DEC~uVAX3600
AP=none
OS=VMS~5.3
DISK=1 MEM=32
PLOT=HI~DMP29,HP~7475A
TV=IIS~75,IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR89
FRACT=10 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bruce Wilking}%
\CPaddrline{Univ. of Missouri at St. Louis}%
\CPaddrline{Physics Dept., 8001 Natural Bridge Rd.}%
\CPaddrline{St. Louis, MO, 63121}%
\CPaddrline{USA}%
\CPlastname{Wilking, Bruce}%
\CPcountry{USA}%
\CPzip{63121}%
\CPcity{St. Louis, MO,}%
\CPinstitute{Physics Dept., 8001 Natural Bridge Rd.}%
\CPstreet{Univ. of Missouri at St. Louis}%
\CPtelephone{ 314-553-5023}%
\CPnetaddr{ c1796@umslvaxa.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{166}{%
USERS=10 PROGS=1 TASKS=5 BAND=optical,IR
}%
\CPmachine{166}{%
SITE=U.~Missouri
CPU=DEC~VAX8600
AP=none
OS=VMS~5.0
DISK=2 MEM=72
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi,Kennedy~9610
AVUSERS=1 AIPS=15OCT89
FRACT=20 VLA=50 NONRADIO=35
}%
}%

\CP{
\CPaddr{
\CPaddrline{Andy Feldt}%
\CPaddrline{Univ. of Oklahoma}%
\CPaddrline{Dept. of Physics and Astronomy, 440 W. Brooks}%
\CPaddrline{Norman, OK, 73019}%
\CPaddrline{USA}%
\CPlastname{Feldt, Andy}%
\CPcountry{USA}%
\CPzip{73019}%
\CPcity{Norman, OK,}%
\CPinstitute{Dept. of Physics and Astronomy, 440 W. Brooks}%
\CPstreet{Univ. of Oklahoma}%
\CPtelephone{ 405-325-3961}%
\CPnetaddr{ feldt@phyast.nhn.uoknor.edu}%
\CPnetaddr{ feldt@uokniels.bitnet}%
\CPnetaddr{ uokhep::atom1::feldt}%
\CPattribute{CP}%
}%
\CPsitedata{167}{%
USERS=4 PROGS=0 TASKS=1 BAND=optical,IR
}%
\CPmachine{167}{%
SITE=U.~Oklahoma
CPU=DEC~uVAX~II/GPX
AP=none
OS=VMS~5.3
DISK=0.7 MEM=9
PLOT=Apple~Laserwriter,QMS~PS-800~II
TV=IIS~IVAS
TAPE=1600bpi,TK50
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=80 NONRADIO=20
}%
}%

\CP{
\CPaddr{
\CPaddrline{David H. Hough}%
\CPaddrline{Trinity Univ.}%
\CPaddrline{Physics Dept., 715 Stadium Dr.}%
\CPaddrline{San Antonio, TX, 78212}%
\CPaddrline{USA}%
\CPlastname{Hough, David H.}%
\CPcountry{USA}%
\CPzip{78212}%
\CPcity{San Antonio, TX,}%
\CPinstitute{Physics Dept., 715 Stadium Dr.}%
\CPstreet{Trinity Univ.}%
\CPtelephone{ 512-736-7466}%
\CPnetaddr{ dhough@trinity.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{168}{%
USERS=6 PROGS=0 TASKS=0
}%
\CPmachine{168}{%
SITE=Trinity~U.
CPU=Sun-4/330
AP=none
OS=UNIX~Sun~4.1
DISK=0.2 MEM=8
PLOT=Apple~Laserwriter+
TV=Sun-4,Sony~Trinitron
TAPE=QIC24,QIC150
AVUSERS=2 AIPS=15JAN90
FRACT=50 VLA=70 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dean C. Hines}%
\CPaddrline{Univ. of Texas at Austin}%
\CPaddrline{Dept. of Astronomy}%
\CPaddrline{Austin, TX, 78712-1083}%
\CPaddrline{USA}%
\CPlastname{Hines, Dean C.}%
\CPcountry{USA}%
\CPzip{78712-1083}%
\CPcity{Austin, TX,}%
\CPinstitute{Dept. of Astronomy}%
\CPstreet{Univ. of Texas at Austin}%
\CPtelephone{ 512-471-8443}%
\CPnetaddr{ dhines@astro.as.utexas.edu}%
\CPattribute{CP}%
}%
\CPsitedata{169}{%
USERS=15 PROGS=2 TASKS=6
}%
\CPmachine{169}{%
SITE=U.~Texas~Austin
CPU=Sun-3/280S
AP=none
OS=UNIX~Sun~4.0
DISK=0.6 MEM=8
PLOT=Apple~Laserwriter~II
TV=Sun-3~color
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=60 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. George A. Dulk}%
\CPaddrline{Univ. of Colorado, Dept. of Astrophys., Plan. \&\ Atm. Sci.}%
\CPaddrline{Campus Box 391}%
\CPaddrline{Boulder, CO, 80309-0391}%
\CPaddrline{USA}%
\CPlastname{Dulk, Dr. George A.}%
\CPcountry{USA}%
\CPzip{80309-0391}%
\CPcity{Boulder, CO,}%
\CPinstitute{Univ. of Colorado, Dept. of Astrophys., Plan. \&\ Atm. Sci.}%
\CPstreet{Campus Box 391}%
\CPtelephone{ 303-492-8788}%
\CPnetaddr{ solarx::dulk}%
\CPattribute{CP}%
}%
\CPsitedata{170}{%
USERS=3 PROGS=1 TASKS=7 BAND=IR,optical,UV
}%
\CPmachine{170}{%
SITE=U.~Colorado~APAS
CPU=DEC~VAX750
AP=FPS~5310
OS=VMS~x.x
DISK=1.2 MEM=8
PLOT=QMS~800
TV=Grinnell~xxxx
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=95 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Terry A. Kucera}%
\CPaddrline{Univ. of Colorado, Dept. of Astrophys, Plan. \&\ Atm. Sci.}%
\CPaddrline{Campus Box 391}%
\CPaddrline{Boulder, CO, 80309-0391}%
\CPaddrline{USA}%
\CPlastname{Kucera, Terry A.}%
\CPcountry{USA}%
\CPzip{80309-0391}%
\CPcity{Boulder, CO,}%
\CPinstitute{Univ. of Colorado, Dept. of Astrophys, Plan. \&\ Atm. Sci.}%
\CPstreet{Campus Box 391}%
\CPtelephone{ 303-492-8788}%
\CPnetaddr{ solarx::kucera}%
\CPattribute{CP}%
}%
\CPsitedata{171}{%
USERS=3 PROGS=0 TASKS=0
}%
\CPmachine{171}{%
SITE=U.~Colorado~APAS
CPU=Sun~Clus.~(5*4/65GX)
AP=none
OS=UNIX~Sun~4.4.1
DISK=1 MEM=16
PLOT=SPARCprinter,HP~LaserJet~III
TV=Sony~Trinitron
TAPE=6250bpi,Exabyte,CD-ROM
AVUSERS=1 AIPS=15JUL90
FRACT=100 VLA=80 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Kris Mickus}%
\CPaddrline{Univ. of Colorado}%
\CPaddrline{CASA, Campus Box 391}%
\CPaddrline{Boulder, CO, 80309-0391}%
\CPaddrline{USA}%
\CPlastname{Mickus, Kris}%
\CPcountry{USA}%
\CPzip{80309-0391}%
\CPcity{Boulder, CO,}%
\CPinstitute{Univ. of Colorado}%
\CPstreet{CASA, Campus Box 391}%
\CPtelephone{ 303-492-6785}%
\CPnetaddr{ taurus::domik}%
\CPnetaddr{ cygnus::domik}%
\CPnetaddr{ cosmos::mickus}%
\CPattribute{CP}%
}%
\CPsitedata{172}{%
USERS=4 PROGS=0 TASKS=5
}%
\CPmachine{172}{%
SITE=U.~Colorado~CASA
CPU=DEC~uVAX3100
AP=none
OS=VMS~5.3
DISK=?
PLOT=QMS~Talaris~800,QMS~Talaris~xxx
TV=DEC~(X-Windows)
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=100 NONRADIO=0
}%
\CPmachine{173}{%
SITE=U.~Colorado~CASA
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.3
DISK=0.1 MEM=16
PLOT=QMS~Talaris~800,QMS~Talaris~xxx
TV=IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT90
FRACT=5 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Charles Woodward}%
\CPaddrline{Univ. of Wyoming}%
\CPaddrline{Dept. of Phys. \&\ Astron., P.O. Box 3905 Univ. Sta}%
\CPaddrline{Laramie, WY, 82071}%
\CPaddrline{USA}%
\CPlastname{Woodward, Charles}%
\CPcountry{USA}%
\CPzip{82071}%
\CPcity{Laramie, WY,}%
\CPinstitute{Dept. of Phys. \&\ Astron., P.O. Box 3905 Univ. Sta}%
\CPstreet{Univ. of Wyoming}%
\CPtelephone{ 307-766-3058;6150}%
\CPnetaddr{ chelsea@uwvax.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{174}{%
USERS=4 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{174}{%
SITE=U.~Wyoming~IR~Obs.
CPU=Masscomp~MC5500
AP=none
OS=UNIX~x.x
DISK=0.8 MEM=10
PLOT=HP~LaserJet+
TV=Masscomp~GA1000
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=3 NONRADIO=7
}%
}%

\CP{
\CPaddr{
\CPaddrline{Rogier A. Windhorst}%
\CPaddrline{Arizona St. Univ.}%
\CPaddrline{Dept. of Physics}%
\CPaddrline{Tempe, AZ, 85287-1504}%
\CPaddrline{USA}%
\CPlastname{Windhorst, Rogier A.}%
\CPcountry{USA}%
\CPzip{85287-1504}%
\CPcity{Tempe, AZ,}%
\CPinstitute{Dept. of Physics}%
\CPstreet{Arizona St. Univ.}%
\CPtelephone{ 602-965-7143}%
\CPnetaddr{ raw@cosmos.la.asu.edu}%
\CPnetaddr{ atraw@asuacvax.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{175}{%
USERS=2 PROGS=1 TASKS=0
}%
\CPmachine{175}{%
SITE=Arizona~St.~U.
CPU=DECstation~5000
AP=none
OS=UNIX~Ultrix~4.0
DISK=1 MEM=24
PLOT=?
TV=DECstation~(SAO~Image)
TAPE=800bpi,1600bpi,3200bpi,6250bpi
AVUSERS=1 AIPS=15JAN91
FRACT=10 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bill Peters}%
\CPaddrline{Univ. of Arizona}%
\CPaddrline{SMT Group, Steward Obs.}%
\CPaddrline{Tucson, AZ, 85721}%
\CPaddrline{USA}%
\CPlastname{Peters, Bill}%
\CPcountry{USA}%
\CPzip{85721}%
\CPcity{Tucson, AZ,}%
\CPinstitute{SMT Group, Steward Obs.}%
\CPstreet{Univ. of Arizona}%
\CPtelephone{ 602-621-5380}%
\CPnetaddr{ tucsmt.as.arizona.edu}%
\CPattribute{CP}%
}%
\CPsitedata{176}{%
USERS=2 PROGS=0 TASKS=0
}%
\CPmachine{176}{%
SITE=Steward~Obs.
CPU=DEC~VAX750
AP=none
OS=VMS~5.2
DISK=0.5 MEM=6.25
PLOT=QMS~Lasergrafix
TV=none
TAPE=1600bpi
AVUSERS=1 AIPS=15APR89
FRACT=1 VLA=1 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Erick Young}%
\CPaddrline{Univ. of Arizona}%
\CPaddrline{Steward Obs.}%
\CPaddrline{Tucson, AZ, 85721}%
\CPaddrline{USA}%
\CPlastname{Young, Erick}%
\CPcountry{USA}%
\CPzip{85721}%
\CPcity{Tucson, AZ,}%
\CPinstitute{Steward Obs.}%
\CPstreet{Univ. of Arizona}%
\CPtelephone{ 602-621-4119}%
\CPtelex{ 407175}%
\CPnetaddr{ eyoung@as.arizona.edu}%
\CPattribute{CP}%
}%
\CPsitedata{177}{%
USERS=5 PROGS=1 TASKS=0 BAND=IR,optical
}%
\CPmachine{177}{%
SITE=Steward~Obs.
CPU=Sun-3/160
AP=none
OS=UNIX~Sun~x.x
DISK=1.8 MEM=8
PLOT=TI~2108~(Postscript)
TV=Sun-3~color
TAPE=?
AVUSERS=? AIPS=?
FRACT=15 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Deidre Hunter}%
\CPaddrline{Lowell Observatory}%
\CPaddrline{1400 W. Mars Hill Road}%
\CPaddrline{Flagstaff, AZ, 86001}%
\CPaddrline{USA}%
\CPlastname{Hunter, Deidre}%
\CPcountry{USA}%
\CPzip{86001}%
\CPcity{Flagstaff, AZ,}%
\CPinstitute{Lowell Observatory}%
\CPstreet{1400 W. Mars Hill Road}%
\CPtelephone{ 602-774-3358}%
\CPnetaddr{ percy::dah}%
\CPattribute{CP}%
}%
\CPsitedata{178}{%
USERS=1 PROGS=0 TASKS=0
}%
\CPmachine{178}{%
SITE=Lowell~Obs.
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=16
PLOT=?
TV=Sun-4~(Sunview)
TAPE=6250bpi
AVUSERS=1 AIPS=15JUL90
FRACT=100 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Nebojsa Duric}%
\CPaddrline{Univ. of New Mexico}%
\CPaddrline{Dept. of Physics \&\ Astronomy}%
\CPaddrline{Albuquerque, NM, 87131}%
\CPaddrline{USA}%
\CPlastname{Duric, Nebojsa}%
\CPcountry{USA}%
\CPzip{87131}%
\CPcity{Albuquerque, NM,}%
\CPinstitute{Dept. of Physics \&\ Astronomy}%
\CPstreet{Univ. of New Mexico}%
\CPtelephone{ 505-277-5238/2617}%
\CPnetaddr{ aips@unmb.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{179}{%
USERS=14 PROGS=2 TASKS=6 BAND=IR,optical,X-ray
}%
\CPmachine{179}{%
SITE=U.~New~Mexico
CPU=DEC~VAX8650
AP=FPS~120B
OS=VMS~x.x
DISK=0.5 MEM=8
PLOT=QMS~800,HP~LaserJet~II
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=80 NONRADIO=20
}%
\CPmachine{180}{%
SITE=U.~New~Mexico
CPU=DEC~uVAX3200
AP=none
OS=VMS~x.x
DISK=0.5 MEM=8
PLOT=QMS~800,HP~LaserJet~II
TV=IIS~75
TAPE=?
AVUSERS=? AIPS=?
FRACT=90 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Abe Jacobson}%
\CPaddrline{Los Alamos National Laboratory}%
\CPaddrline{MS D-466}%
\CPaddrline{Los Alamos, NM, 87545}%
\CPaddrline{USA}%
\CPlastname{Jacobson, Abe}%
\CPcountry{USA}%
\CPzip{87545}%
\CPcity{Los Alamos, NM,}%
\CPinstitute{Los Alamos National Laboratory}%
\CPstreet{MS D-466}%
\CPtelephone{ 505-667-9656}%
\CPnetaddr{ 083044@essdp1.lanl.gov}%
\CPnetaddr{ 085044@essdp2.lanl.gov}%
\CPattribute{CP}%
}%
\CPsitedata{181}{%
USERS=1 PROGS=1 TASKS=0
}%
\CPmachine{181}{%
SITE=LANL
CPU=DEC~VAXxxxx
AP=none
OS=VMS~5.3-1
DISK=0.2 MEM=32
PLOT=?
TV=?
TAPE=6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=5 VLA=0 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Jean Eilek}%
\CPaddrline{Nat. Radio Astronomy Obs.}%
\CPaddrline{Array Operations Center}%
\CPaddrline{Socorro, NM, 87801}%
\CPaddrline{USA}%
\CPlastname{Eilek, Jean}%
\CPcountry{USA}%
\CPzip{87801}%
\CPcity{Socorro, NM,}%
\CPinstitute{Array Operations Center}%
\CPstreet{Nat. Radio Astronomy Obs.}%
\CPtelephone{ 505-835-7393}%
\CPnetaddr{ jeilek@nrao.edu}%
\CPattribute{CP}%
}%
\CPsitedata{182}{%
USERS=1 PROGS=0 TASKS=0 BAND=other
}%
\CPmachine{182}{%
SITE=New~Mexico~Tech~AOC
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~x.x
DISK=0.5 MEM=16
PLOT=?
TV=Sun-4
TAPE=?
AVUSERS=1 AIPS=15APR91
FRACT=50 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Rene Walterbos}%
\CPaddrline{New Mexico State Univ.}%
\CPaddrline{Dept. of Astronomy, Box 30001/Dept. 4500}%
\CPaddrline{Las Cruces, NM, 88003}%
\CPaddrline{USA}%
\CPlastname{Walterbos, Rene}%
\CPcountry{USA}%
\CPzip{88003}%
\CPcity{Las Cruces, NM,}%
\CPinstitute{Dept. of Astronomy, Box 30001/Dept. 4500}%
\CPstreet{New Mexico State Univ.}%
\CPtelephone{ 505-646-6522}%
\CPnetaddr{ rwalterb@nmsu.edu}%
\CPattribute{CP}%
}%
\CPsitedata{183}{%
USERS=47 PROGS=0 TASKS=1
}%
\CPmachine{183}{%
SITE=New~Mexico~St.~U.
CPU=Sun-4/390
AP=none
OS=UNIX~Sun~4.1
DISK=2 MEM=32
PLOT=Apple~Laserwriter~IIx
TV=Sun-4~color
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT89
FRACT=30 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Mark Morris}%
\CPaddrline{Dept. of Astronomy, UCLA}%
\CPaddrline{Math Sciences Building, 405 Hilgard Ave.}%
\CPaddrline{Los Angeles, CA, 90024}%
\CPaddrline{USA}%
\CPlastname{Morris, Mark}%
\CPcountry{USA}%
\CPzip{90024}%
\CPcity{Los Angeles, CA,}%
\CPinstitute{Dept. of Astronomy, UCLA}%
\CPstreet{Math Sciences Building, 405 Hilgard Ave.}%
\CPtelephone{ 213-825-3320}%
\CPnetaddr{ bonnie::aips (5.708)}%
\CPnetaddr{ aips@uclastro.bitnet}%
\CPnetaddr{ aips@astro.ucla.edu}%
\CPattribute{CP}%
}%
\CPsitedata{184}{%
USERS=16 PROGS=0 TASKS=0 BAND=IR,optical
}%
\CPmachine{184}{%
SITE=UC~Los~Angeles
CPU=Sun-3/260
AP=none
OS=UNIX~Sun~4.0.1
DISK=0.575 MEM=32
PLOT=Apple~Laserwriter
TV=Sun-3~color
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=90 NONRADIO=10
}%
\CPmachine{185}{%
SITE=UC~Los~Angeles
CPU=Sun-3/160
AP=none
OS=UNIX~Sun~4.0.1
DISK=0.575 MEM=16
PLOT=Apple~Laserwriter
TV=Sun-3
TAPE=1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=80 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{A. J. Annala}%
\CPaddrline{Univ. of Southern Calif.}%
\CPaddrline{Psychology Dept., Mail Code 1061}%
\CPaddrline{Los Angeles, CA, 90089-1061}%
\CPaddrline{USA}%
\CPlastname{Annala, A. J.}%
\CPcountry{USA}%
\CPzip{90089-1061}%
\CPcity{Los Angeles, CA,}%
\CPinstitute{Psychology Dept., Mail Code 1061}%
\CPstreet{Univ. of Southern Calif.}%
\CPtelephone{ 213-743-3278}%
\CPnetaddr{ annala@neuro.usc.edu}%
\CPattribute{CP}%
}%
\CPsitedata{186}{%
USERS=1 PROGS=1 TASKS=0 BAND=optical
}%
\CPmachine{186}{%
SITE=USC~Psychology
CPU=Sun-3/160
AP=none
OS=UNIX~Sun~x.x
DISK=0.28 MEM=8
PLOT=Apple~Laserwriter
TV=Sun-3
TAPE=?
AVUSERS=? AIPS=?
FRACT=5 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Braden E. Hines}%
\CPaddrline{Jet Propulsion Lab}%
\CPaddrline{MS 169-214, 4800 Oak Grove Drive}%
\CPaddrline{Pasadena, CA, 91109}%
\CPaddrline{USA}%
\CPlastname{Hines, Braden E.}%
\CPcountry{USA}%
\CPzip{91109}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Jet Propulsion Lab}%
\CPstreet{MS 169-214, 4800 Oak Grove Drive}%
\CPtelephone{ 818-354-2465}%
\CPnetaddr{ brad@huey.jpl.nasa.gov}%
\CPattribute{CP}%
}%
\CPsitedata{187}{%
USERS=3 PROGS=2 TASKS=0 BAND=optical
}%
\CPmachine{187}{%
SITE=JPL~Optical~Interferometry
CPU=Sun-4/65
AP=none
OS=UNIX~Sun~4.1
DISK=0.2 MEM=64
PLOT=HP~LaserJet~III~(w/PS)
TV=Sun-4~(X-Windows)
TAPE=?
AVUSERS=1 AIPS=15JUL90
FRACT=1 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dayton Jones}%
\CPaddrline{Jet Propulsion Lab.}%
\CPaddrline{Mail Code 238-700, 4800 Oak Grove Dr.}%
\CPaddrline{Pasadena, CA, 91109}%
\CPaddrline{USA}%
\CPlastname{Jones, Dayton}%
\CPcountry{USA}%
\CPzip{91109}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Jet Propulsion Lab.}%
\CPstreet{Mail Code 238-700, 4800 Oak Grove Dr.}%
\CPtelephone{ FTS 792-7774, 818-354-7774}%
\CPnetaddr{ xenos::dj}%
\CPnetaddr{ 6030::dj}%
\CPattribute{CP}%
}%
\CPsitedata{188}{%
USERS=7 PROGS=1 TASKS=1
}%
\CPmachine{188}{%
SITE=NASA-JPL~XENOS
CPU=DEC~VAX785
AP=none
OS=VMS~x.x
DISK=0.25 MEM=16
PLOT=Printronix~MVP,Imagen~8/300
TV=?
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=10 NONRADIO=5
}%
}%

\CP{
\CPaddr{
\CPaddrline{Thomas B. H. Kuiper}%
\CPaddrline{Jet Propulsion Lab.}%
\CPaddrline{169/506, 4800 Oak Grove Dr.}%
\CPaddrline{Pasadena, CA, 91109}%
\CPaddrline{USA}%
\CPlastname{Kuiper, Thomas B. H.}%
\CPcountry{USA}%
\CPzip{91109}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Jet Propulsion Lab.}%
\CPstreet{169/506, 4800 Oak Grove Dr.}%
\CPtelephone{ 818-354-5623; FTS 792-5623}%
\CPnetaddr{ kuiper\%jplrag.jpl.nasa.gov}%
\CPnetaddr{ jplrag::kuiper}%
\CPattribute{CP}%
}%
\CPsitedata{189}{%
USERS=5 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{189}{%
SITE=JPL~RAG
CPU=DEC~VAX750
AP=none
OS=VMS~5.0-2
DISK=0.9 MEM=12
PLOT=Tektronix~4010,Apple~Laserwriter+
TV=?
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15JUL86
FRACT=5 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{David L. Meier}%
\CPaddrline{California Inst. of Tech.}%
\CPaddrline{264-700, Jet Propulsion Lab., 4800 Oak Grove Dr.}%
\CPaddrline{Pasadena, CA, 91109}%
\CPaddrline{USA}%
\CPlastname{Meier, David L.}%
\CPcountry{USA}%
\CPzip{91109}%
\CPcity{Pasadena, CA,}%
\CPinstitute{264-700, Jet Propulsion Lab., 4800 Oak Grove Dr.}%
\CPstreet{California Inst. of Tech.}%
\CPtelephone{ 213-354-5062}%
\CPnetaddr{ dlm@cena.jpl.nasa.gov}%
\CPnetaddr{ 5332::"dlm"}%
\CPnetaddr{ CENA::"dlm"}%
\CPattribute{CP}%
}%
\CPsitedata{190}{%
USERS=6 PROGS=0 TASKS=1 BAND=optical
}%
\CPmachine{190}{%
SITE=JPL~Section~335
CPU=Sun~Clus.~(2*4/60;4/390)
AP=none
OS=UNIX~Sun~4.1
DISK=0.33 MEM=32
PLOT=Apple~Laserwriter~II
TV=Sun-4~(Sunview)
TAPE=1600bpi,6250bpi,QIC24,QIC150
AVUSERS=2 AIPS=15JUL90
FRACT=50 VLA=25 NONRADIO=25
}%
}%

\CP{
\CPaddr{
\CPaddrline{Don Rudy}%
\CPaddrline{Jet Propulsion Lab}%
\CPaddrline{Mail Code 183-501}%
\CPaddrline{Pasadena, CA, 91109}%
\CPaddrline{USA}%
\CPlastname{Rudy, Don}%
\CPcountry{USA}%
\CPzip{91109}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Jet Propulsion Lab}%
\CPstreet{Mail Code 183-501}%
\CPtelephone{ 818-354-1812}%
\CPnetaddr{ FLYBY::DRUDY}%
\CPnetaddr{ djr@venus1.gps.caltech.edu}%
\CPattribute{CP}%
}%
\CPsitedata{191}{%
USERS=2 PROGS=1 TASKS=4
}%
\CPmachine{191}{%
SITE=JPL~Planetary~Astronomy
CPU=DEC~VAXxxxx
AP=none
OS=VMS~x.x
DISK=0.5 MEM=13
PLOT=?
TV=Peritek~VCX
TAPE=6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=25 VLA=10 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bryan Butler}%
\CPaddrline{California Inst. of Tech.}%
\CPaddrline{Div. of Planetary Science, Mail Stop 170-25}%
\CPaddrline{Pasadena, CA, 91125}%
\CPaddrline{USA}%
\CPlastname{Butler, Bryan}%
\CPcountry{USA}%
\CPzip{91125}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Div. of Planetary Science, Mail Stop 170-25}%
\CPstreet{California Inst. of Tech.}%
\CPtelephone{ 818-356-6477}%
\CPnetaddr{ VENUS::BUTLER}%
\CPnetaddr{ butler@venus1.gps.caltech.edu (131.215.67.167)}%
\CPattribute{CP}%
}%
\CPsitedata{192}{%
USERS=7 PROGS=3 TASKS=10 BAND=radar,sub-mm
}%
\CPmachine{192}{%
SITE=CIT~Plan.~Sci.
CPU=DEC~uVAX~II
AP=none
OS=VMS~5.1
DISK=0.7 MEM=9
PLOT=TI~2115~laser,Apple~Laserwriter~II
TV=Peritek~VCH-Q
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR90
FRACT=50 VLA=50 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{John Conway}%
\CPaddrline{Owens Valley Radio Observatory}%
\CPaddrline{Mail Code 105-24, Caltech}%
\CPaddrline{Pasadena, CA, 91125}%
\CPaddrline{USA}%
\CPlastname{Conway, John}%
\CPcountry{USA}%
\CPzip{91125}%
\CPcity{Pasadena, CA,}%
\CPinstitute{Owens Valley Radio Observatory}%
\CPstreet{Mail Code 105-24, Caltech}%
\CPtelephone{ 818-356-6857}%
\CPnetaddr{ jec@deimos.caltech.edu}%
\CPattribute{CP}%
}%
\CPsitedata{193}{%
USERS=20 PROGS=0 TASKS=20
}%
\CPmachine{193}{%
SITE=CIT~Astronomy
CPU=Convex~C-1/XP
AP=Vector~Regs
OS=UNIX~Convex~7.1
DISK=0.6 MEM=32
PLOT=Printronix~xxx,Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=2 AIPS=15JUL90
FRACT=50 VLA=15 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Tom Fuller}%
\CPaddrline{Univ. of Calif. at Santa Barbara}%
\CPaddrline{Physics Dept.}%
\CPaddrline{Santa Barbara, CA, 93106}%
\CPaddrline{USA}%
\CPlastname{Fuller, Tom}%
\CPcountry{USA}%
\CPzip{93106}%
\CPcity{Santa Barbara, CA,}%
\CPinstitute{Physics Dept.}%
\CPstreet{Univ. of Calif. at Santa Barbara}%
\CPtelephone{ 805-893-8875 (or 893-4567 for messages)}%
\CPnetaddr{ fuller@voodoo.ucsb.edu}%
\CPnetaddr{ fuller@voodoo.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{194}{%
USERS=5 PROGS=1 TASKS=0
}%
\CPmachine{194}{%
SITE=UC~Santa~Barbara
CPU=Sun-4/330
AP=none
OS=UNIX~Sun~4.0.3
DISK=2
PLOT=Postscript
TV=Sun-4~(Sunview),Sun-4~(X-Windows)
TAPE=1600bpi,6250bpi,QIC24,QIC150
AVUSERS=2 AIPS=15OCT89
FRACT=75 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Butler Hine}%
\CPaddrline{NASA-Ames Res. Ctr.}%
\CPaddrline{MS 245-3}%
\CPaddrline{Moffett Field, CA, 94035}%
\CPaddrline{USA}%
\CPlastname{Hine, Butler}%
\CPcountry{USA}%
\CPzip{94035}%
\CPcity{Moffett Field, CA,}%
\CPinstitute{NASA-Ames Res. Ctr.}%
\CPstreet{MS 245-3}%
\CPtelephone{ 415-694-5910}%
\CPnetaddr{ hine@24609.span}%
\CPattribute{CP}%
}%
\CPsitedata{195}{%
USERS=3 PROGS=0 TASKS=0 BAND=IR,optical
}%
\CPmachine{195}{%
SITE=NASA-Ames
CPU=DEC~VAX8600
AP=none
OS=VMS~x.x
DISK=0.128 MEM=12
PLOT=Imagen~8/300,Versatec~xxxx
TV=IIS~IVAS
TAPE=?
AVUSERS=? AIPS=?
FRACT=10 VLA=99 NONRADIO=1
}%
\CPmachine{196}{%
SITE=NASA-Ames
CPU=DEC~VAX785
AP=none
OS=VMS~4.6
DISK=0.128 MEM=12
PLOT=Imagen~8/300,Versatec~xxxx
TV=IIS~IVAS
TAPE=?
AVUSERS=? AIPS=?
FRACT=1 VLA=100 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Robert H. Becker}%
\CPaddrline{LLNL}%
\CPaddrline{3192 Malvasia Court}%
\CPaddrline{Pleasanton, CA, 94566}%
\CPaddrline{USA}%
\CPlastname{Becker, Robert H.}%
\CPcountry{USA}%
\CPzip{94566}%
\CPcity{Pleasanton, CA,}%
\CPinstitute{(no institute?)}%
\CPstreet{LLNL, 3192 Malvasia Court}%
\CPtelephone{ 916-752-6921,1500; (h)415-462-3189; (LLNL)415-423-0664}%
\CPnetaddr{ bob@wyrd.llnl.gov}%
\CPattribute{CP}%
}%
\CPsitedata{197}{%
USERS=4 PROGS=0 TASKS=0 BAND=IR
}%
\CPmachine{197}{%
SITE=LLNL-IGPP
CPU=Sun-4/380
AP=FPA
OS=UNIX~Sun~x.x
DISK=1.8 MEM=8
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=90 VLA=90 NONRADIO=10
}%
\CPmachine{198}{%
SITE=LLNL-IGPP
CPU=Sun-4/390
AP=FPA
OS=UNIX~Sun~x.x
DISK=? MEM=8
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=25 VLA=90 NONRADIO=10
}%
\CPmachine{199}{%
SITE=LLNL-IGPP
CPU=Sun-4/390
AP=none
OS=UNIX~Sun~x.x
DISK=1.5 MEM=8
PLOT=Apple~Laserwriter
TV=IIS~IVAS
TAPE=1600bpi,6250bpi
AVUSERS=2 AIPS=15JUL90
FRACT=50 VLA=90 NONRADIO=10
}%
}%

\CP{
\CPaddr{
\CPaddrline{Wilson Hoffman}%
\CPaddrline{Univ. of Calif.}%
\CPaddrline{Astronomy Dept., 601 Campbell Hall}%
\CPaddrline{Berkeley, CA, 94720}%
\CPaddrline{USA}%
\CPlastname{Hoffman, Wilson}%
\CPcountry{USA}%
\CPzip{94720}%
\CPcity{Berkeley, CA,}%
\CPinstitute{Astronomy Dept., 601 Campbell Hall}%
\CPstreet{Univ. of Calif.}%
\CPtelephone{ 415-642-7768;6424}%
\CPnetaddr{ hoffman@bkyast.berkeley.edu}%
\CPnetaddr{ hoffman\%bkyast.hepnet@lbl.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{200}{%
USERS=12 PROGS=1 TASKS=1
}%
\CPmachine{200}{%
SITE=UC~Berkeley
CPU=Sun-4/60
AP=none
OS=UNIX~Sun~4.1
DISK=0.6 MEM=8
PLOT=QMS~810
TV=Sun-4~color
TAPE=1600bpi,6250bpi,Exabyte
AVUSERS=1 AIPS=15OCT90
FRACT=20 VLA=95 NONRADIO=0
}%
\CPmachine{201}{%
SITE=UC~Berkeley
CPU=DEC~uVAX3200
AP=none
OS=VMS~5.2
DISK=0.6 MEM=12
PLOT=Imagen~2308S
TV=IIS~IVAS
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15APR89
FRACT=15 VLA=95 NONRADIO=0
}%
}%

\CP{
\CPaddr{
\CPaddrline{Bruce Balick}%
\CPaddrline{Univ. of Washington}%
\CPaddrline{Astronomy Dept. FM-20}%
\CPaddrline{Seattle, WA, 98195}%
\CPaddrline{USA}%
\CPlastname{Balick, Bruce}%
\CPcountry{USA}%
\CPzip{98195}%
\CPcity{Seattle, WA,}%
\CPinstitute{Astronomy Dept. FM-20}%
\CPstreet{Univ. of Washington}%
\CPtelephone{ 206-543-7683}%
\CPnetaddr{ balick@uwaphast.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{202}{%
USERS=14 PROGS=0 TASKS=0 BAND=IR,optical
}%
\CPmachine{202}{%
SITE=U.~Washington
CPU=DEC~Clus.~(4*u3200)
AP=none
OS=VMS~5.3a
DISK=0.8 MEM=16
PLOT=QMS~Talaris~1200,Apple~Laserwriter
TV=IIS~75
TAPE=800bpi,1600bpi,6250bpi
AVUSERS=1 AIPS=15OCT88
FRACT=1 VLA=0 NONRADIO=100
}%
\CPmachine{203}{%
SITE=U.~Washington
CPU=DEC~uVAX3200
AP=none
OS=VMS~5.3a
DISK=0.6 MEM=9
PLOT=QMS~Talaris~1200,Apple~Laserwriter
TV=IIS~75
TAPE=?
AVUSERS=1 AIPS=15OCT88
FRACT=30 VLA=0 NONRADIO=100
}%
}%

\CP{
\CPaddr{
\CPaddrline{Dr. Julie Lutz}%
\CPaddrline{Washington State Univ.}%
\CPaddrline{Prog. in Astronomy}%
\CPaddrline{Pullman, WA, 99164-2930}%
\CPaddrline{USA}%
\CPlastname{Lutz, Dr. Julie}%
\CPcountry{USA}%
\CPzip{99164-2930}%
\CPcity{Pullman, WA,}%
\CPinstitute{Washington State Univ.}%
\CPstreet{Prog. in Astronomy}%
\CPnetaddr{ jlutz@wsumath.bitnet}%
\CPattribute{CP}%
}%
\CPsitedata{204}{%
USERS=2 PROGS=0 TASKS=0 BAND=optical
}%
\CPmachine{204}{%
SITE=Washington~St.~U.
CPU=DEC~VAX785
AP=none
OS=VMS~4.4
DISK=0.1 MEM=12
PLOT=?
TV=IIS~70E
TAPE=1600bpi
AVUSERS=? AIPS=15JUL86
FRACT=1 VLA=0 NONRADIO=0
}%
}%

\enddirectory

\begincrossref

\XRkey{AP}{%
\XRvalue{FPA}{%
\XRmachine{197}%
\XRmachine{198}%
\XRnmach{2}}%
\XRvalue{FPS~120B}{%
\XRmachine{22}%
\XRmachine{64}%
\XRmachine{76}%
\XRmachine{77}%
\XRmachine{179}%
\XRnmach{5}}%
\XRvalue{FPS~5040}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{FPS~5105}{%
\XRmachine{108}%
\XRnmach{1}}%
\XRvalue{FPS~5200}{%
\XRmachine{24}%
\XRnmach{1}}%
\XRvalue{FPS~5205}{%
\XRmachine{46}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{150}%
\XRnmach{4}}%
\XRvalue{FPS~5300}{%
\XRmachine{124}%
\XRnmach{1}}%
\XRvalue{FPS~5305}{%
\XRmachine{125}%
\XRmachine{128}%
\XRnmach{2}}%
\XRvalue{FPS~5310}{%
\XRmachine{170}%
\XRnmach{1}}%
\XRvalue{PSAP}{%
\XRmachine{88}%
\XRnmach{1}}%
\XRvalue{Vector~Regs}{%
\XRmachine{4}%
\XRmachine{38}%
\XRmachine{50}%
\XRmachine{52}%
\XRmachine{55}%
\XRmachine{59}%
\XRmachine{62}%
\XRmachine{68}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{81}%
\XRmachine{82}%
\XRmachine{87}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{122}%
\XRmachine{141}%
\XRmachine{158}%
\XRmachine{162}%
\XRmachine{193}%
\XRnmach{22}}%
\XRvalue{none}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{3}%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{10}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{13}%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{17}%
\XRmachine{18}%
\XRmachine{19}%
\XRmachine{20}%
\XRmachine{21}%
\XRmachine{23}%
\XRmachine{25}%
\XRmachine{26}%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{29}%
\XRmachine{30}%
\XRmachine{31}%
\XRmachine{32}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{35}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{39}%
\XRmachine{40}%
\XRmachine{41}%
\XRmachine{42}%
\XRmachine{43}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{47}%
\XRmachine{48}%
\XRmachine{49}%
\XRmachine{51}%
\XRmachine{53}%
\XRmachine{54}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{60}%
\XRmachine{61}%
\XRmachine{63}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{67}%
\XRmachine{69}%
\XRmachine{70}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{79}%
\XRmachine{80}%
\XRmachine{83}%
\XRmachine{84}%
\XRmachine{85}%
\XRmachine{86}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{103}%
\XRmachine{104}%
\XRmachine{105}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{109}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{116}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{120}%
\XRmachine{121}%
\XRmachine{123}%
\XRmachine{126}%
\XRmachine{127}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{133}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{139}%
\XRmachine{140}%
\XRmachine{142}%
\XRmachine{143}%
\XRmachine{144}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{148}%
\XRmachine{149}%
\XRmachine{151}%
\XRmachine{152}%
\XRmachine{153}%
\XRmachine{154}%
\XRmachine{155}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{159}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{174}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{177}%
\XRmachine{178}%
\XRmachine{180}%
\XRmachine{181}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{188}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{194}%
\XRmachine{195}%
\XRmachine{196}%
\XRmachine{199}%
\XRmachine{200}%
\XRmachine{201}%
\XRmachine{202}%
\XRmachine{203}%
\XRmachine{204}%
\XRnmach{163}}%
}%

\XRkey{BAND}{%
\XRvalue{IR}{%
\XRmachine{1}%
\XRmachine{4}%
\XRmachine{18}%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{31}%
\XRmachine{33}%
\XRmachine{35}%
\XRmachine{52}%
\XRmachine{68}%
\XRmachine{82}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{106}%
\XRmachine{110}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{116}%
\XRmachine{118}%
\XRmachine{122}%
\XRmachine{126}%
\XRmachine{128}%
\XRmachine{133}%
\XRmachine{140}%
\XRmachine{150}%
\XRmachine{158}%
\XRmachine{160}%
\XRmachine{162}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{170}%
\XRmachine{174}%
\XRmachine{177}%
\XRmachine{179}%
\XRmachine{184}%
\XRmachine{189}%
\XRmachine{195}%
\XRmachine{197}%
\XRmachine{202}%
\XRnmach{39}}%
\XRvalue{UV}{%
\XRmachine{49}%
\XRmachine{128}%
\XRmachine{133}%
\XRmachine{146}%
\XRmachine{156}%
\XRmachine{170}%
\XRnmach{6}}%
\XRvalue{X-ray}{%
\XRmachine{46}%
\XRmachine{54}%
\XRmachine{64}%
\XRmachine{68}%
\XRmachine{98}%
\XRmachine{113}%
\XRmachine{118}%
\XRmachine{140}%
\XRmachine{142}%
\XRmachine{144}%
\XRmachine{152}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{165}%
\XRmachine{179}%
\XRnmach{15}}%
\XRvalue{biomed}{%
\XRmachine{4}%
\XRmachine{165}%
\XRnmach{2}}%
\XRvalue{holography}{%
\XRmachine{92}%
\XRnmach{1}}%
\XRvalue{optical}{%
\XRmachine{1}%
\XRmachine{4}%
\XRmachine{15}%
\XRmachine{18}%
\XRmachine{22}%
\XRmachine{29}%
\XRmachine{31}%
\XRmachine{32}%
\XRmachine{33}%
\XRmachine{39}%
\XRmachine{42}%
\XRmachine{44}%
\XRmachine{46}%
\XRmachine{49}%
\XRmachine{50}%
\XRmachine{55}%
\XRmachine{59}%
\XRmachine{61}%
\XRmachine{64}%
\XRmachine{68}%
\XRmachine{72}%
\XRmachine{74}%
\XRmachine{83}%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{98}%
\XRmachine{103}%
\XRmachine{106}%
\XRmachine{111}%
\XRmachine{113}%
\XRmachine{122}%
\XRmachine{127}%
\XRmachine{128}%
\XRmachine{133}%
\XRmachine{139}%
\XRmachine{141}%
\XRmachine{146}%
\XRmachine{150}%
\XRmachine{153}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{160}%
\XRmachine{162}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{170}%
\XRmachine{177}%
\XRmachine{179}%
\XRmachine{184}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{190}%
\XRmachine{195}%
\XRmachine{202}%
\XRmachine{204}%
\XRnmach{57}}%
\XRvalue{other}{%
\XRmachine{48}%
\XRmachine{49}%
\XRmachine{82}%
\XRmachine{90}%
\XRmachine{110}%
\XRmachine{142}%
\XRmachine{153}%
\XRmachine{182}%
\XRnmach{8}}%
\XRvalue{radar}{%
\XRmachine{72}%
\XRmachine{74}%
\XRmachine{90}%
\XRmachine{107}%
\XRmachine{192}%
\XRnmach{5}}%
\XRvalue{sub-mm}{%
\XRmachine{16}%
\XRmachine{192}%
\XRnmach{2}}%
}%

\XRkey{CP Lastname}{%
\XRvalue{Allen, Lori}{%
\XRmachine{91}%
\XRnmach{1}}%
\XRvalue{Aller, Hugh D.}{%
\XRmachine{150}%
\XRnmach{1}}%
\XRvalue{Anantharamaiah, K.}{%
\XRmachine{45}%
\XRnmach{1}}%
\XRvalue{Andernach, Dr. Heinz}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{Annala, A. J.}{%
\XRmachine{186}%
\XRnmach{1}}%
\XRvalue{Baan, Willem A.}{%
\XRmachine{90}%
\XRnmach{1}}%
\XRvalue{Baath, Lars B.}{%
\XRmachine{82}%
\XRnmach{1}}%
\XRvalue{Balick, Bruce}{%
\XRmachine{202}%
\XRnmach{1}}%
\XRvalue{Basart, John P.}{%
\XRmachine{152}%
\XRnmach{1}}%
\XRvalue{Batuski, David}{%
\XRmachine{109}%
\XRnmach{1}}%
\XRvalue{Beck, Dr. Sara}{%
\XRmachine{47}%
\XRnmach{1}}%
\XRvalue{Becker, Robert H.}{%
\XRmachine{197}%
\XRnmach{1}}%
\XRvalue{Bosma, A.}{%
\XRmachine{32}%
\XRnmach{1}}%
\XRvalue{Bradford, David}{%
\XRmachine{106}%
\XRnmach{1}}%
\XRvalue{Briggs, Frank}{%
\XRmachine{117}%
\XRnmach{1}}%
\XRvalue{Butler, Bryan}{%
\XRmachine{192}%
\XRnmach{1}}%
\XRvalue{Calabretta, Dr. Mark R.}{%
\XRmachine{4}%
\XRnmach{1}}%
\XRvalue{Campbell, Jonathan}{%
\XRmachine{144}%
\XRnmach{1}}%
\XRvalue{Carmody, Jack}{%
\XRmachine{110}%
\XRnmach{1}}%
\XRvalue{Catalano, Prof. Santo}{%
\XRmachine{49}%
\XRnmach{1}}%
\XRvalue{Catarzi, Marco}{%
\XRmachine{48}%
\XRnmach{1}}%
\XRvalue{Cesarsky, Diego}{%
\XRmachine{35}%
\XRnmach{1}}%
\XRvalue{Christiansen, Wayne}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{Churchwell, Edward B.}{%
\XRmachine{156}%
\XRnmach{1}}%
\XRvalue{Clegg, Andrew}{%
\XRmachine{116}%
\XRnmach{1}}%
\XRvalue{Cogan, Bruce C.}{%
\XRmachine{11}%
\XRnmach{1}}%
\XRvalue{Conner, Sam}{%
\XRmachine{95}%
\XRnmach{1}}%
\XRvalue{Conway, John}{%
\XRmachine{193}%
\XRnmach{1}}%
\XRvalue{Cornwell, Timothy J.}{%
\XRmachine{72}%
\XRnmach{1}}%
\XRvalue{Costa, Marco}{%
\XRmachine{12}%
\XRnmach{1}}%
\XRvalue{Cotton, Bill}{%
\XRmachine{64}%
\XRnmach{1}}%
\XRvalue{Crawford, Dr. David F.}{%
\XRmachine{1}%
\XRnmach{1}}%
\XRvalue{Despois, Didier}{%
\XRmachine{33}%
\XRnmach{1}}%
\XRvalue{Diamond, Phil J.}{%
\XRmachine{73}%
\XRnmach{1}}%
\XRvalue{Dulk, Dr. George A.}{%
\XRmachine{170}%
\XRnmach{1}}%
\XRvalue{Duric, Nebojsa}{%
\XRmachine{179}%
\XRnmach{1}}%
\XRvalue{Eilek, Jean}{%
\XRmachine{182}%
\XRnmach{1}}%
\XRvalue{Emerson, Darrel}{%
\XRmachine{71}%
\XRnmach{1}}%
\XRvalue{Fejes, Istvan}{%
\XRmachine{41}%
\XRnmach{1}}%
\XRvalue{Feldt, Andy}{%
\XRmachine{167}%
\XRnmach{1}}%
\XRvalue{Fich, Michel}{%
\XRmachine{19}%
\XRnmach{1}}%
\XRvalue{Florkowski, David}{%
\XRmachine{124}%
\XRnmach{1}}%
\XRvalue{Foley, Tony}{%
\XRmachine{59}%
\XRnmach{1}}%
\XRvalue{Fuller, Tom}{%
\XRmachine{194}%
\XRnmach{1}}%
\XRvalue{Gaume, Ralph}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{Ghigo, Frank}{%
\XRmachine{70}%
\XRnmach{1}}%
\XRvalue{Glendenning, Brian}{%
\XRmachine{65}%
\XRnmach{1}}%
\XRvalue{Gregory, Phil}{%
\XRmachine{23}%
\XRnmach{1}}%
\XRvalue{Grieve, Gerry}{%
\XRmachine{24}%
\XRnmach{1}}%
\XRvalue{Hamilton, Dr. P.A. (Pip)}{%
\XRmachine{13}%
\XRnmach{1}}%
\XRvalue{Hennessy, Gregory S.}{%
\XRmachine{139}%
\XRnmach{1}}%
\XRvalue{Herbertz, Ralf}{%
\XRmachine{37}%
\XRnmach{1}}%
\XRvalue{Hewitt, Jacqueline N.}{%
\XRmachine{103}%
\XRnmach{1}}%
\XRvalue{Hine, Butler}{%
\XRmachine{195}%
\XRnmach{1}}%
\XRvalue{Hines, Braden E.}{%
\XRmachine{187}%
\XRnmach{1}}%
\XRvalue{Hines, Dean C.}{%
\XRmachine{169}%
\XRnmach{1}}%
\XRvalue{Hoffman, Lyle}{%
\XRmachine{120}%
\XRnmach{1}}%
\XRvalue{Hoffman, Wilson}{%
\XRmachine{200}%
\XRnmach{1}}%
\XRvalue{Hollis, Dr. Jan M.}{%
\XRmachine{127}%
\XRnmach{1}}%
\XRvalue{Hough, David H.}{%
\XRmachine{168}%
\XRnmach{1}}%
\XRvalue{Hughes, V. A.}{%
\XRmachine{17}%
\XRnmach{1}}%
\XRvalue{Hunter, Deidre}{%
\XRmachine{178}%
\XRnmach{1}}%
\XRvalue{Irwin, Judith}{%
\XRmachine{16}%
\XRnmach{1}}%
\XRvalue{Ishiguro, Prof. Masato}{%
\XRmachine{52}%
\XRnmach{1}}%
\XRvalue{Jacobson, Abe}{%
\XRmachine{181}%
\XRnmach{1}}%
\XRvalue{Jaffe, Walter}{%
\XRmachine{55}%
\XRnmach{1}}%
\XRvalue{Jones, Dayton}{%
\XRmachine{188}%
\XRnmach{1}}%
\XRvalue{Junor, Bill}{%
\XRmachine{74}%
\XRnmach{1}}%
\XRvalue{Kaplan, Dr. George H.}{%
\XRmachine{125}%
\XRnmach{1}}%
\XRvalue{Kaufman, Michele}{%
\XRmachine{146}%
\XRnmach{1}}%
\XRvalue{Kerdraon, Alain}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{Knapp, Dr. G. R.}{%
\XRmachine{112}%
\XRnmach{1}}%
\XRvalue{Kucera, Terry A.}{%
\XRmachine{171}%
\XRnmach{1}}%
\XRvalue{Kuiper, Thomas B. H.}{%
\XRmachine{189}%
\XRnmach{1}}%
\XRvalue{Kutner, Marc L.}{%
\XRmachine{114}%
\XRnmach{1}}%
\XRvalue{Kwok, Sun}{%
\XRmachine{20}%
\XRnmach{1}}%
\XRvalue{Laing, Dr. Robert A.}{%
\XRmachine{85}%
\XRnmach{1}}%
\XRvalue{Lehar, Joseph}{%
\XRmachine{104}%
\XRnmach{1}}%
\XRvalue{Lehto, Harry}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{Leous, James}{%
\XRmachine{118}%
\XRnmach{1}}%
\XRvalue{Lonsdale, Dr. Colin J.}{%
\XRmachine{92}%
\XRnmach{1}}%
\XRvalue{Lutz, Dr. Julie}{%
\XRmachine{204}%
\XRnmach{1}}%
\XRvalue{Manager, AIPS}{%
\XRmachine{27}%
\XRmachine{31}%
\XRnmach{2}}%
\XRvalue{Meier, David L.}{%
\XRmachine{190}%
\XRnmach{1}}%
\XRvalue{Mickus, Kris}{%
\XRmachine{172}%
\XRnmach{1}}%
\XRvalue{Mizumoto, Yoshihiko}{%
\XRmachine{54}%
\XRnmach{1}}%
\XRvalue{Morris, Mark}{%
\XRmachine{184}%
\XRnmach{1}}%
\XRvalue{Muchmore, David O.}{%
\XRmachine{15}%
\XRnmach{1}}%
\XRvalue{Murphy, Pat}{%
\XRmachine{68}%
\XRnmach{1}}%
\XRvalue{Mutel, Robert}{%
\XRmachine{153}%
\XRnmach{1}}%
\XRvalue{Nanni, M.}{%
\XRmachine{50}%
\XRnmach{1}}%
\XRvalue{Neff, Susan}{%
\XRmachine{128}%
\XRnmach{1}}%
\XRvalue{Pati, A. K.}{%
\XRmachine{44}%
\XRnmach{1}}%
\XRvalue{Payne, Harry}{%
\XRmachine{134}%
\XRnmach{1}}%
\XRvalue{Pedelty, Jeffrey A.}{%
\XRmachine{133}%
\XRnmach{1}}%
\XRvalue{Peters, Bill}{%
\XRmachine{176}%
\XRnmach{1}}%
\XRvalue{Pilbratt, Goran}{%
\XRmachine{63}%
\XRnmach{1}}%
\XRvalue{Prasad, Debi}{%
\XRmachine{42}%
\XRnmach{1}}%
\XRvalue{Prestage, Richard}{%
\XRmachine{86}%
\XRnmach{1}}%
\XRvalue{Radford, Simon}{%
\XRmachine{34}%
\XRnmach{1}}%
\XRvalue{Rengarajan, Dr. T. N.}{%
\XRmachine{43}%
\XRnmach{1}}%
\XRvalue{Reynolds, Stephen P.}{%
\XRmachine{142}%
\XRnmach{1}}%
\XRvalue{Roberts, Dave}{%
\XRmachine{108}%
\XRnmach{1}}%
\XRvalue{Rosenthal, Shoshana}{%
\XRmachine{98}%
\XRnmach{1}}%
\XRvalue{Rudnick, Larry}{%
\XRmachine{158}%
\XRnmach{1}}%
\XRvalue{Rudy, Don}{%
\XRmachine{191}%
\XRnmach{1}}%
\XRvalue{Ruiz-Falco, Antonio}{%
\XRmachine{79}%
\XRnmach{1}}%
\XRvalue{Russell, Claire}{%
\XRmachine{113}%
\XRnmach{1}}%
\XRvalue{Sandqvist, Dr. Aage}{%
\XRmachine{81}%
\XRnmach{1}}%
\XRvalue{Schmidt, Juergen}{%
\XRmachine{38}%
\XRnmach{1}}%
\XRvalue{Schultz, Dan}{%
\XRmachine{157}%
\XRnmach{1}}%
\XRvalue{Shone, Dave}{%
\XRmachine{87}%
\XRnmach{1}}%
\XRvalue{Simkin, Susan M.}{%
\XRmachine{151}%
\XRnmach{1}}%
\XRvalue{Simonetti, John}{%
\XRmachine{140}%
\XRnmach{1}}%
\XRvalue{Sinha, Ramesh}{%
\XRmachine{126}%
\XRnmach{1}}%
\XRvalue{Sukumar, S.}{%
\XRmachine{162}%
\XRnmach{1}}%
\XRvalue{Tarr, Susan}{%
\XRmachine{18}%
\XRnmach{1}}%
\XRvalue{Taylor, Charles A.}{%
\XRmachine{143}%
\XRnmach{1}}%
\XRvalue{Titterington, David J.}{%
\XRmachine{83}%
\XRnmach{1}}%
\XRvalue{Trice, James}{%
\XRmachine{145}%
\XRnmach{1}}%
\XRvalue{Turner, George}{%
\XRmachine{148}%
\XRnmach{1}}%
\XRvalue{Tyson, Tony}{%
\XRmachine{111}%
\XRnmach{1}}%
\XRvalue{Vannier, Dr. Michael}{%
\XRmachine{165}%
\XRnmach{1}}%
\XRvalue{Vaughan, Dr. A.}{%
\XRmachine{3}%
\XRnmach{1}}%
\XRvalue{Velusamy, Dr. T.}{%
\XRmachine{46}%
\XRnmach{1}}%
\XRvalue{Walterbos, Rene}{%
\XRmachine{183}%
\XRnmach{1}}%
\XRvalue{Wilking, Bruce}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{Willis, Anthony G.}{%
\XRmachine{22}%
\XRnmach{1}}%
\XRvalue{Willson, Robert F.}{%
\XRmachine{105}%
\XRnmach{1}}%
\XRvalue{Windhorst, Rogier A.}{%
\XRmachine{175}%
\XRnmach{1}}%
\XRvalue{Woodward, Charles}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{Xingwu, Zheng}{%
\XRmachine{29}%
\XRnmach{1}}%
\XRvalue{Young, Erick}{%
\XRmachine{177}%
\XRnmach{1}}%
\XRvalue{Yuhai, Qiu}{%
\XRmachine{25}%
\XRnmach{1}}%
\XRvalue{Yusef-Zadeh, Farhad}{%
\XRmachine{160}%
\XRnmach{1}}%
\XRvalue{Zhang, Y. F.}{%
\XRmachine{107}%
\XRnmach{1}}%
\XRvalue{van Moorsel, Gustaaf}{%
\XRmachine{39}%
\XRnmach{1}}%
\XRvalue{van der Hulst, J.M.}{%
\XRmachine{61}%
\XRnmach{1}}%
}%

\XRkey{CPU}{%
\XRvalue{Alliant~FX/8-3CE}{%
\XRmachine{59}%
\XRmachine{122}%
\XRnmach{2}}%
\XRvalue{Alliant~FX/8-4CE}{%
\XRmachine{81}%
\XRmachine{87}%
\XRnmach{2}}%
\XRvalue{Alliant~FX/80}{%
\XRmachine{62}%
\XRnmach{1}}%
\XRvalue{Apollo~DN3000}{%
\XRmachine{95}%
\XRnmach{1}}%
\XRvalue{Apollo~DN320}{%
\XRmachine{96}%
\XRnmach{1}}%
\XRvalue{Apollo~DN660}{%
\XRmachine{97}%
\XRnmach{1}}%
\XRvalue{Convex~C-1/XL}{%
\XRmachine{82}%
\XRnmach{1}}%
\XRvalue{Convex~C-1/XP}{%
\XRmachine{38}%
\XRmachine{68}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{193}%
\XRnmach{9}}%
\XRvalue{Convex~C-201}{%
\XRmachine{50}%
\XRnmach{1}}%
\XRvalue{Convex~C-210}{%
\XRmachine{162}%
\XRnmach{1}}%
\XRvalue{Convex~C-220}{%
\XRmachine{4}%
\XRmachine{55}%
\XRnmach{2}}%
\XRvalue{Convex~C-240}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{Cray~X-MP/4}{%
\XRmachine{158}%
\XRnmach{1}}%
\XRvalue{DECstation~3100}{%
\XRmachine{56}%
\XRmachine{66}%
\XRnmach{2}}%
\XRvalue{DECstation~5000}{%
\XRmachine{175}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(2*3500;750)}{%
\XRmachine{104}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(2*uVAXII;uVAXIII)}{%
\XRmachine{37}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(3500;2*750;2*2000)}{%
\XRmachine{13}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(3600;3100;2000;5000)}{%
\XRmachine{63}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(3800;uVAXII;3100)}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(4*u3200)}{%
\XRmachine{202}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(6400;8820;8250;2*780;3*3200;3100}{%
\XRmachine{133}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(780;3*u2000)}{%
\XRmachine{35}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(780;3100)}{%
\XRmachine{144}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(8800)}{%
\XRmachine{137}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(u3900;3200)}{%
\XRmachine{33}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(uVAX3400)}{%
\XRmachine{85}%
\XRnmach{1}}%
\XRvalue{DEC~Clus.~(uVII;u3200)}{%
\XRmachine{34}%
\XRnmach{1}}%
\XRvalue{DEC~VAX4000}{%
\XRmachine{155}%
\XRnmach{1}}%
\XRvalue{DEC~VAX6310}{%
\XRmachine{79}%
\XRnmach{1}}%
\XRvalue{DEC~VAX6410}{%
\XRmachine{128}%
\XRnmach{1}}%
\XRvalue{DEC~VAX750}{%
\XRmachine{24}%
\XRmachine{46}%
\XRmachine{49}%
\XRmachine{77}%
\XRmachine{91}%
\XRmachine{116}%
\XRmachine{121}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{170}%
\XRmachine{176}%
\XRmachine{189}%
\XRnmach{14}}%
\XRvalue{DEC~VAX780}{%
\XRmachine{14}%
\XRmachine{22}%
\XRmachine{27}%
\XRmachine{43}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{64}%
\XRmachine{76}%
\XRmachine{86}%
\XRmachine{110}%
\XRmachine{156}%
\XRnmach{11}}%
\XRvalue{DEC~VAX785}{%
\XRmachine{111}%
\XRmachine{138}%
\XRmachine{188}%
\XRmachine{196}%
\XRmachine{204}%
\XRnmach{5}}%
\XRvalue{DEC~VAX8200}{%
\XRmachine{126}%
\XRnmach{1}}%
\XRvalue{DEC~VAX8350}{%
\XRmachine{129}%
\XRmachine{130}%
\XRnmach{2}}%
\XRvalue{DEC~VAX8530}{%
\XRmachine{30}%
\XRmachine{124}%
\XRmachine{125}%
\XRnmach{3}}%
\XRvalue{DEC~VAX8600}{%
\XRmachine{36}%
\XRmachine{40}%
\XRmachine{147}%
\XRmachine{166}%
\XRmachine{195}%
\XRnmach{5}}%
\XRvalue{DEC~VAX8650}{%
\XRmachine{108}%
\XRmachine{179}%
\XRnmach{2}}%
\XRvalue{DEC~VAXxxxx}{%
\XRmachine{181}%
\XRmachine{191}%
\XRnmach{2}}%
\XRvalue{DEC~uVAX3100}{%
\XRmachine{172}%
\XRnmach{1}}%
\XRvalue{DEC~uVAX3200}{%
\XRmachine{154}%
\XRmachine{180}%
\XRmachine{201}%
\XRmachine{203}%
\XRnmach{4}}%
\XRvalue{DEC~uVAX3300}{%
\XRmachine{51}%
\XRnmach{1}}%
\XRvalue{DEC~uVAX3400}{%
\XRmachine{60}%
\XRmachine{88}%
\XRnmach{2}}%
\XRvalue{DEC~uVAX3600}{%
\XRmachine{165}%
\XRnmach{1}}%
\XRvalue{DEC~uVAX~II}{%
\XRmachine{2}%
\XRmachine{26}%
\XRmachine{28}%
\XRmachine{41}%
\XRmachine{42}%
\XRmachine{80}%
\XRmachine{84}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{106}%
\XRmachine{123}%
\XRmachine{139}%
\XRmachine{142}%
\XRmachine{149}%
\XRmachine{157}%
\XRmachine{173}%
\XRmachine{192}%
\XRnmach{18}}%
\XRvalue{DEC~uVAX~II/GPX}{%
\XRmachine{31}%
\XRmachine{167}%
\XRnmach{2}}%
\XRvalue{DG~4000~Server}{%
\XRmachine{21}%
\XRnmach{1}}%
\XRvalue{Fujitsu~M380S}{%
\XRmachine{53}%
\XRnmach{1}}%
\XRvalue{Fujitsu~VP50}{%
\XRmachine{52}%
\XRnmach{1}}%
\XRvalue{IBM~RS-6000}{%
\XRmachine{65}%
\XRnmach{1}}%
\XRvalue{Masscomp~MC5500}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{SG~Iris~4D/20}{%
\XRmachine{19}%
\XRnmach{1}}%
\XRvalue{Stellar~1000}{%
\XRmachine{152}%
\XRnmach{1}}%
\XRvalue{Sun-3/160}{%
\XRmachine{177}%
\XRmachine{185}%
\XRmachine{186}%
\XRnmach{3}}%
\XRvalue{Sun-3/180}{%
\XRmachine{16}%
\XRmachine{18}%
\XRmachine{164}%
\XRnmach{3}}%
\XRvalue{Sun-3/260}{%
\XRmachine{184}%
\XRnmach{1}}%
\XRvalue{Sun-3/260C}{%
\XRmachine{54}%
\XRnmach{1}}%
\XRvalue{Sun-3/280S}{%
\XRmachine{169}%
\XRnmach{1}}%
\XRvalue{Sun-3/60}{%
\XRmachine{10}%
\XRmachine{48}%
\XRmachine{70}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{109}%
\XRmachine{115}%
\XRmachine{127}%
\XRmachine{140}%
\XRnmach{9}}%
\XRvalue{Sun-4/110}{%
\XRmachine{67}%
\XRnmach{1}}%
\XRvalue{Sun-4/260}{%
\XRmachine{20}%
\XRmachine{94}%
\XRnmach{2}}%
\XRvalue{Sun-4/280}{%
\XRmachine{145}%
\XRnmach{1}}%
\XRvalue{Sun-4/280S}{%
\XRmachine{161}%
\XRnmach{1}}%
\XRvalue{Sun-4/330}{%
\XRmachine{32}%
\XRmachine{168}%
\XRmachine{194}%
\XRnmach{3}}%
\XRvalue{Sun-4/370}{%
\XRmachine{8}%
\XRnmach{1}}%
\XRvalue{Sun-4/380}{%
\XRmachine{197}%
\XRnmach{1}}%
\XRvalue{Sun-4/390}{%
\XRmachine{183}%
\XRmachine{198}%
\XRmachine{199}%
\XRnmach{3}}%
\XRvalue{Sun-4/40}{%
\XRmachine{160}%
\XRnmach{1}}%
\XRvalue{Sun-4/490}{%
\XRmachine{163}%
\XRnmach{1}}%
\XRvalue{Sun-4/60}{%
\XRmachine{17}%
\XRmachine{23}%
\XRmachine{39}%
\XRmachine{47}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{72}%
\XRmachine{90}%
\XRmachine{93}%
\XRmachine{114}%
\XRmachine{143}%
\XRmachine{146}%
\XRmachine{148}%
\XRmachine{178}%
\XRmachine{182}%
\XRmachine{200}%
\XRnmach{16}}%
\XRvalue{Sun-4/60GX}{%
\XRmachine{9}%
\XRmachine{71}%
\XRnmach{2}}%
\XRvalue{Sun-4/65}{%
\XRmachine{3}%
\XRmachine{6}%
\XRmachine{92}%
\XRmachine{120}%
\XRmachine{187}%
\XRnmach{5}}%
\XRvalue{Sun-4/65GX}{%
\XRmachine{7}%
\XRnmach{1}}%
\XRvalue{Sun-4/75}{%
\XRmachine{12}%
\XRnmach{1}}%
\XRvalue{Sun-X/xxx}{%
\XRmachine{105}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(2*3/60;2*4/60)}{%
\XRmachine{159}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(2*4/40)}{%
\XRmachine{153}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(2*4/60)}{%
\XRmachine{83}%
\XRmachine{107}%
\XRnmach{2}}%
\XRvalue{Sun~Clus.~(2*4/60;4/390)}{%
\XRmachine{190}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(3*4/60;4/60GX)}{%
\XRmachine{5}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(3*4/65)}{%
\XRmachine{25}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(3/50;3/60G;3/110)}{%
\XRmachine{69}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4*4/65)}{%
\XRmachine{61}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/110;4/20)}{%
\XRmachine{119}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/280;3/60)}{%
\XRmachine{29}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/330;2*4/60)}{%
\XRmachine{103}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/380S;4*4/60)}{%
\XRmachine{15}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/390~server;4/100;4/370;4/65;4*}{%
\XRmachine{11}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/490~server;3/280;8*3/60)}{%
\XRmachine{136}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/490~server;4/280;10*4/60)}{%
\XRmachine{134}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/490~server;4/60)}{%
\XRmachine{1}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/490~server;6*4/60)}{%
\XRmachine{135}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(4/65;4/60)}{%
\XRmachine{118}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(5*4/60)}{%
\XRmachine{117}%
\XRnmach{1}}%
\XRvalue{Sun~Clus.~(5*4/65GX)}{%
\XRmachine{171}%
\XRnmach{1}}%
}%

\XRkey{Country}{%
\XRvalue{AUSTRALIA}{%
\XRmachine{1}%
\XRmachine{3}%
\XRmachine{4}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{13}%
\XRnmach{6}}%
\XRvalue{BRAZIL}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{CANADA}{%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{17}%
\XRmachine{18}%
\XRmachine{19}%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{24}%
\XRnmach{9}}%
\XRvalue{CHINA}{%
\XRmachine{25}%
\XRmachine{27}%
\XRmachine{29}%
\XRnmach{3}}%
\XRvalue{DENMARK}{%
\XRmachine{31}%
\XRnmach{1}}%
\XRvalue{FRANCE}{%
\XRmachine{32}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{35}%
\XRmachine{36}%
\XRnmach{5}}%
\XRvalue{GERMANY}{%
\XRmachine{37}%
\XRmachine{38}%
\XRmachine{39}%
\XRnmach{3}}%
\XRvalue{HUNGARY}{%
\XRmachine{41}%
\XRnmach{1}}%
\XRvalue{INDIA}{%
\XRmachine{42}%
\XRmachine{43}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{46}%
\XRnmach{5}}%
\XRvalue{ISRAEL}{%
\XRmachine{47}%
\XRnmach{1}}%
\XRvalue{ITALY}{%
\XRmachine{48}%
\XRmachine{49}%
\XRmachine{50}%
\XRnmach{3}}%
\XRvalue{JAPAN}{%
\XRmachine{52}%
\XRmachine{54}%
\XRnmach{2}}%
\XRvalue{NETHERLANDS}{%
\XRmachine{55}%
\XRmachine{59}%
\XRmachine{61}%
\XRmachine{63}%
\XRnmach{4}}%
\XRvalue{NRAO}{%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{70}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{74}%
\XRnmach{8}}%
\XRvalue{SPAIN}{%
\XRmachine{79}%
\XRnmach{1}}%
\XRvalue{SWEDEN}{%
\XRmachine{81}%
\XRmachine{82}%
\XRnmach{2}}%
\XRvalue{UK}{%
\XRmachine{83}%
\XRmachine{85}%
\XRmachine{86}%
\XRmachine{87}%
\XRmachine{89}%
\XRnmach{5}}%
\XRvalue{USA}{%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{95}%
\XRmachine{98}%
\XRmachine{103}%
\XRmachine{104}%
\XRmachine{105}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{116}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{120}%
\XRmachine{122}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{126}%
\XRmachine{127}%
\XRmachine{128}%
\XRmachine{133}%
\XRmachine{134}%
\XRmachine{139}%
\XRmachine{140}%
\XRmachine{141}%
\XRmachine{142}%
\XRmachine{143}%
\XRmachine{144}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{148}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{152}%
\XRmachine{153}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{158}%
\XRmachine{160}%
\XRmachine{162}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{170}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{174}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{177}%
\XRmachine{178}%
\XRmachine{179}%
\XRmachine{181}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{188}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{194}%
\XRmachine{195}%
\XRmachine{197}%
\XRmachine{200}%
\XRmachine{202}%
\XRmachine{204}%
\XRnmach{79}}%
}%

\XRkey{DISK}{%
\XRvalue{0.05}{%
\XRmachine{43}%
\XRmachine{126}%
\XRmachine{144}%
\XRnmach{3}}%
\XRvalue{0.06}{%
\XRmachine{42}%
\XRnmach{1}}%
\XRvalue{0.1}{%
\XRmachine{2}%
\XRmachine{84}%
\XRmachine{89}%
\XRmachine{104}%
\XRmachine{111}%
\XRmachine{116}%
\XRmachine{147}%
\XRmachine{173}%
\XRmachine{204}%
\XRnmach{9}}%
\XRvalue{0.128}{%
\XRmachine{195}%
\XRmachine{196}%
\XRnmach{2}}%
\XRvalue{0.18}{%
\XRmachine{27}%
\XRnmach{1}}%
\XRvalue{0.2}{%
\XRmachine{1}%
\XRmachine{14}%
\XRmachine{47}%
\XRmachine{73}%
\XRmachine{114}%
\XRmachine{168}%
\XRmachine{181}%
\XRmachine{187}%
\XRnmach{8}}%
\XRvalue{0.227}{%
\XRmachine{31}%
\XRnmach{1}}%
\XRvalue{0.25}{%
\XRmachine{85}%
\XRmachine{188}%
\XRnmach{2}}%
\XRvalue{0.256}{%
\XRmachine{86}%
\XRnmach{1}}%
\XRvalue{0.28}{%
\XRmachine{115}%
\XRmachine{186}%
\XRnmach{2}}%
\XRvalue{0.3}{%
\XRmachine{40}%
\XRmachine{44}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{66}%
\XRmachine{70}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{105}%
\XRmachine{109}%
\XRmachine{123}%
\XRnmach{13}}%
\XRvalue{0.33}{%
\XRmachine{190}%
\XRnmach{1}}%
\XRvalue{0.36}{%
\XRmachine{28}%
\XRnmach{1}}%
\XRvalue{0.37}{%
\XRmachine{157}%
\XRnmach{1}}%
\XRvalue{0.4}{%
\XRmachine{19}%
\XRmachine{24}%
\XRmachine{33}%
\XRmachine{35}%
\XRmachine{41}%
\XRmachine{60}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{140}%
\XRmachine{155}%
\XRmachine{162}%
\XRmachine{164}%
\XRnmach{15}}%
\XRvalue{0.43}{%
\XRmachine{80}%
\XRnmach{1}}%
\XRvalue{0.5}{%
\XRmachine{13}%
\XRmachine{16}%
\XRmachine{22}%
\XRmachine{34}%
\XRmachine{39}%
\XRmachine{54}%
\XRmachine{91}%
\XRmachine{102}%
\XRmachine{121}%
\XRmachine{152}%
\XRmachine{176}%
\XRmachine{179}%
\XRmachine{180}%
\XRmachine{182}%
\XRmachine{191}%
\XRnmach{15}}%
\XRvalue{0.575}{%
\XRmachine{184}%
\XRmachine{185}%
\XRnmach{2}}%
\XRvalue{0.6}{%
\XRmachine{3}%
\XRmachine{15}%
\XRmachine{26}%
\XRmachine{45}%
\XRmachine{67}%
\XRmachine{72}%
\XRmachine{78}%
\XRmachine{83}%
\XRmachine{120}%
\XRmachine{124}%
\XRmachine{143}%
\XRmachine{145}%
\XRmachine{163}%
\XRmachine{169}%
\XRmachine{178}%
\XRmachine{193}%
\XRmachine{200}%
\XRmachine{201}%
\XRmachine{203}%
\XRnmach{19}}%
\XRvalue{0.63}{%
\XRmachine{69}%
\XRnmach{1}}%
\XRvalue{0.65}{%
\XRmachine{9}%
\XRmachine{71}%
\XRnmach{2}}%
\XRvalue{0.66}{%
\XRmachine{148}%
\XRnmach{1}}%
\XRvalue{0.7}{%
\XRmachine{117}%
\XRmachine{139}%
\XRmachine{149}%
\XRmachine{167}%
\XRmachine{192}%
\XRnmach{5}}%
\XRvalue{0.75}{%
\XRmachine{101}%
\XRnmach{1}}%
\XRvalue{0.8}{%
\XRmachine{20}%
\XRmachine{21}%
\XRmachine{48}%
\XRmachine{59}%
\XRmachine{65}%
\XRmachine{112}%
\XRmachine{142}%
\XRmachine{174}%
\XRmachine{202}%
\XRnmach{9}}%
\XRvalue{0.9}{%
\XRmachine{17}%
\XRmachine{146}%
\XRmachine{189}%
\XRnmach{3}}%
\XRvalue{0.912}{%
\XRmachine{46}%
\XRnmach{1}}%
\XRvalue{0.981}{%
\XRmachine{127}%
\XRnmach{1}}%
\XRvalue{1}{%
\XRmachine{12}%
\XRmachine{51}%
\XRmachine{63}%
\XRmachine{90}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{113}%
\XRmachine{119}%
\XRmachine{128}%
\XRmachine{131}%
\XRmachine{156}%
\XRmachine{161}%
\XRmachine{165}%
\XRmachine{171}%
\XRmachine{175}%
\XRnmach{15}}%
\XRvalue{1.13}{%
\XRmachine{77}%
\XRnmach{1}}%
\XRvalue{1.2}{%
\XRmachine{23}%
\XRmachine{30}%
\XRmachine{32}%
\XRmachine{49}%
\XRmachine{81}%
\XRmachine{118}%
\XRmachine{170}%
\XRnmach{7}}%
\XRvalue{1.3}{%
\XRmachine{25}%
\XRmachine{64}%
\XRmachine{103}%
\XRmachine{150}%
\XRnmach{4}}%
\XRvalue{1.4}{%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{151}%
\XRmachine{160}%
\XRnmach{5}}%
\XRvalue{1.5}{%
\XRmachine{37}%
\XRmachine{50}%
\XRmachine{88}%
\XRmachine{153}%
\XRmachine{154}%
\XRmachine{199}%
\XRnmach{6}}%
\XRvalue{1.6}{%
\XRmachine{87}%
\XRnmach{1}}%
\XRvalue{1.7}{%
\XRmachine{82}%
\XRmachine{100}%
\XRnmach{2}}%
\XRvalue{1.73}{%
\XRmachine{133}%
\XRnmach{1}}%
\XRvalue{1.8}{%
\XRmachine{38}%
\XRmachine{177}%
\XRmachine{197}%
\XRnmach{3}}%
\XRvalue{2}{%
\XRmachine{36}%
\XRmachine{106}%
\XRmachine{110}%
\XRmachine{158}%
\XRmachine{166}%
\XRmachine{183}%
\XRmachine{194}%
\XRnmach{7}}%
\XRvalue{2.1}{%
\XRmachine{125}%
\XRnmach{1}}%
\XRvalue{2.2}{%
\XRmachine{76}%
\XRnmach{1}}%
\XRvalue{2.3}{%
\XRmachine{11}%
\XRmachine{18}%
\XRnmach{2}}%
\XRvalue{2.4}{%
\XRmachine{6}%
\XRmachine{7}%
\XRnmach{2}}%
\XRvalue{2.5}{%
\XRmachine{55}%
\XRmachine{61}%
\XRmachine{62}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{159}%
\XRnmach{6}}%
\XRvalue{2.52}{%
\XRmachine{52}%
\XRmachine{53}%
\XRnmach{2}}%
\XRvalue{3}{%
\XRmachine{79}%
\XRnmach{1}}%
\XRvalue{3.5}{%
\XRmachine{75}%
\XRnmach{1}}%
\XRvalue{4}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{4.2}{%
\XRmachine{74}%
\XRnmach{1}}%
\XRvalue{4.4}{%
\XRmachine{4}%
\XRmachine{10}%
\XRnmach{2}}%
\XRvalue{5.8}{%
\XRmachine{68}%
\XRnmach{1}}%
\XRvalue{7}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{8}{%
\XRmachine{29}%
\XRnmach{1}}%
\XRvalue{?}{%
\XRmachine{5}%
\XRmachine{8}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{132}%
\XRmachine{172}%
\XRmachine{198}%
\XRnmach{7}}%
}%

\XRkey{FRACT}{%
\XRvalue{0.1}{%
\XRmachine{70}%
\XRmachine{84}%
\XRmachine{144}%
\XRnmach{3}}%
\XRvalue{0.2}{%
\XRmachine{158}%
\XRnmach{1}}%
\XRvalue{0.3}{%
\XRmachine{32}%
\XRmachine{64}%
\XRnmach{2}}%
\XRvalue{1}{%
\XRmachine{11}%
\XRmachine{33}%
\XRmachine{41}%
\XRmachine{80}%
\XRmachine{111}%
\XRmachine{116}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{133}%
\XRmachine{145}%
\XRmachine{176}%
\XRmachine{187}%
\XRmachine{196}%
\XRmachine{202}%
\XRmachine{204}%
\XRnmach{15}}%
\XRvalue{2}{%
\XRmachine{14}%
\XRmachine{40}%
\XRmachine{85}%
\XRmachine{125}%
\XRmachine{151}%
\XRmachine{162}%
\XRnmach{6}}%
\XRvalue{3}{%
\XRmachine{13}%
\XRmachine{54}%
\XRmachine{147}%
\XRnmach{3}}%
\XRvalue{5}{%
\XRmachine{15}%
\XRmachine{34}%
\XRmachine{36}%
\XRmachine{43}%
\XRmachine{49}%
\XRmachine{63}%
\XRmachine{81}%
\XRmachine{86}%
\XRmachine{89}%
\XRmachine{91}%
\XRmachine{96}%
\XRmachine{102}%
\XRmachine{112}%
\XRmachine{126}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{139}%
\XRmachine{148}%
\XRmachine{149}%
\XRmachine{152}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{181}%
\XRmachine{186}%
\XRmachine{189}%
\XRnmach{28}}%
\XRvalue{5.5}{%
\XRmachine{35}%
\XRnmach{1}}%
\XRvalue{6}{%
\XRmachine{22}%
\XRnmach{1}}%
\XRvalue{10}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{19}%
\XRmachine{27}%
\XRmachine{29}%
\XRmachine{30}%
\XRmachine{37}%
\XRmachine{47}%
\XRmachine{67}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{76}%
\XRmachine{79}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{110}%
\XRmachine{123}%
\XRmachine{124}%
\XRmachine{128}%
\XRmachine{141}%
\XRmachine{163}%
\XRmachine{165}%
\XRmachine{170}%
\XRmachine{174}%
\XRmachine{175}%
\XRmachine{179}%
\XRmachine{188}%
\XRmachine{195}%
\XRnmach{28}}%
\XRvalue{15}{%
\XRmachine{44}%
\XRmachine{142}%
\XRmachine{155}%
\XRmachine{177}%
\XRmachine{201}%
\XRnmach{5}}%
\XRvalue{16}{%
\XRmachine{53}%
\XRnmach{1}}%
\XRvalue{20}{%
\XRmachine{12}%
\XRmachine{42}%
\XRmachine{88}%
\XRmachine{95}%
\XRmachine{97}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{118}%
\XRmachine{132}%
\XRmachine{150}%
\XRmachine{166}%
\XRmachine{200}%
\XRnmach{12}}%
\XRvalue{21}{%
\XRmachine{4}%
\XRnmach{1}}%
\XRvalue{25}{%
\XRmachine{23}%
\XRmachine{31}%
\XRmachine{45}%
\XRmachine{55}%
\XRmachine{117}%
\XRmachine{191}%
\XRmachine{198}%
\XRnmach{7}}%
\XRvalue{30}{%
\XRmachine{24}%
\XRmachine{39}%
\XRmachine{48}%
\XRmachine{50}%
\XRmachine{60}%
\XRmachine{62}%
\XRmachine{90}%
\XRmachine{104}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{120}%
\XRmachine{121}%
\XRmachine{131}%
\XRmachine{183}%
\XRmachine{203}%
\XRnmach{15}}%
\XRvalue{35}{%
\XRmachine{153}%
\XRnmach{1}}%
\XRvalue{40}{%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{59}%
\XRmachine{140}%
\XRmachine{156}%
\XRnmach{6}}%
\XRvalue{50}{%
\XRmachine{8}%
\XRmachine{38}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{105}%
\XRmachine{113}%
\XRmachine{115}%
\XRmachine{143}%
\XRmachine{154}%
\XRmachine{168}%
\XRmachine{182}%
\XRmachine{190}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{199}%
\XRnmach{15}}%
\XRvalue{58}{%
\XRmachine{52}%
\XRnmach{1}}%
\XRvalue{60}{%
\XRmachine{21}%
\XRmachine{51}%
\XRmachine{87}%
\XRmachine{114}%
\XRmachine{119}%
\XRmachine{169}%
\XRnmach{6}}%
\XRvalue{70}{%
\XRmachine{20}%
\XRmachine{92}%
\XRmachine{94}%
\XRmachine{159}%
\XRnmach{4}}%
\XRvalue{75}{%
\XRmachine{75}%
\XRmachine{194}%
\XRnmach{2}}%
\XRvalue{80}{%
\XRmachine{16}%
\XRmachine{18}%
\XRmachine{25}%
\XRmachine{46}%
\XRmachine{61}%
\XRmachine{82}%
\XRmachine{122}%
\XRmachine{127}%
\XRmachine{164}%
\XRmachine{167}%
\XRmachine{184}%
\XRmachine{185}%
\XRnmach{12}}%
\XRvalue{85}{%
\XRmachine{93}%
\XRnmach{1}}%
\XRvalue{90}{%
\XRmachine{3}%
\XRmachine{26}%
\XRmachine{28}%
\XRmachine{68}%
\XRmachine{73}%
\XRmachine{74}%
\XRmachine{77}%
\XRmachine{83}%
\XRmachine{103}%
\XRmachine{180}%
\XRmachine{197}%
\XRnmach{11}}%
\XRvalue{95}{%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{9}%
\XRmachine{78}%
\XRmachine{157}%
\XRnmach{5}}%
\XRvalue{99}{%
\XRmachine{160}%
\XRmachine{161}%
\XRnmach{2}}%
\XRvalue{100}{%
\XRmachine{5}%
\XRmachine{10}%
\XRmachine{17}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{69}%
\XRmachine{146}%
\XRmachine{171}%
\XRmachine{178}%
\XRnmach{9}}%
}%

\XRkey{MEM}{%
\XRvalue{2}{%
\XRmachine{43}%
\XRnmach{1}}%
\XRvalue{2.1}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{3}{%
\XRmachine{77}%
\XRnmach{1}}%
\XRvalue{4}{%
\XRmachine{2}%
\XRmachine{26}%
\XRmachine{44}%
\XRmachine{48}%
\XRmachine{54}%
\XRmachine{72}%
\XRmachine{76}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{115}%
\XRmachine{140}%
\XRnmach{11}}%
\XRvalue{5}{%
\XRmachine{116}%
\XRmachine{157}%
\XRnmach{2}}%
\XRvalue{6}{%
\XRmachine{14}%
\XRmachine{24}%
\XRmachine{46}%
\XRmachine{97}%
\XRnmach{4}}%
\XRvalue{6.25}{%
\XRmachine{176}%
\XRnmach{1}}%
\XRvalue{6.9}{%
\XRmachine{120}%
\XRnmach{1}}%
\XRvalue{8}{%
\XRmachine{10}%
\XRmachine{16}%
\XRmachine{20}%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{34}%
\XRmachine{35}%
\XRmachine{64}%
\XRmachine{67}%
\XRmachine{70}%
\XRmachine{71}%
\XRmachine{86}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{94}%
\XRmachine{109}%
\XRmachine{111}%
\XRmachine{114}%
\XRmachine{119}%
\XRmachine{123}%
\XRmachine{129}%
\XRmachine{131}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{170}%
\XRmachine{177}%
\XRmachine{179}%
\XRmachine{180}%
\XRmachine{186}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRmachine{200}%
\XRnmach{40}}%
\XRvalue{9}{%
\XRmachine{41}%
\XRmachine{84}%
\XRmachine{100}%
\XRmachine{149}%
\XRmachine{167}%
\XRmachine{192}%
\XRmachine{203}%
\XRnmach{7}}%
\XRvalue{10}{%
\XRmachine{42}%
\XRmachine{80}%
\XRmachine{101}%
\XRmachine{110}%
\XRmachine{142}%
\XRmachine{174}%
\XRnmach{6}}%
\XRvalue{11}{%
\XRmachine{49}%
\XRnmach{1}}%
\XRvalue{12}{%
\XRmachine{19}%
\XRmachine{25}%
\XRmachine{31}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{60}%
\XRmachine{69}%
\XRmachine{83}%
\XRmachine{117}%
\XRmachine{127}%
\XRmachine{130}%
\XRmachine{143}%
\XRmachine{147}%
\XRmachine{164}%
\XRmachine{189}%
\XRmachine{195}%
\XRmachine{196}%
\XRmachine{201}%
\XRmachine{204}%
\XRnmach{20}}%
\XRvalue{13}{%
\XRmachine{106}%
\XRmachine{139}%
\XRmachine{191}%
\XRnmach{3}}%
\XRvalue{14}{%
\XRmachine{121}%
\XRnmach{1}}%
\XRvalue{15}{%
\XRmachine{132}%
\XRnmach{1}}%
\XRvalue{16}{%
\XRmachine{1}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{9}%
\XRmachine{12}%
\XRmachine{17}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{29}%
\XRmachine{37}%
\XRmachine{39}%
\XRmachine{45}%
\XRmachine{51}%
\XRmachine{61}%
\XRmachine{66}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{93}%
\XRmachine{102}%
\XRmachine{103}%
\XRmachine{107}%
\XRmachine{118}%
\XRmachine{144}%
\XRmachine{148}%
\XRmachine{154}%
\XRmachine{158}%
\XRmachine{159}%
\XRmachine{171}%
\XRmachine{173}%
\XRmachine{178}%
\XRmachine{182}%
\XRmachine{185}%
\XRmachine{188}%
\XRmachine{202}%
\XRnmach{34}}%
\XRvalue{20}{%
\XRmachine{18}%
\XRmachine{85}%
\XRmachine{88}%
\XRmachine{126}%
\XRnmach{4}}%
\XRvalue{21}{%
\XRmachine{47}%
\XRnmach{1}}%
\XRvalue{24}{%
\XRmachine{3}%
\XRmachine{32}%
\XRmachine{53}%
\XRmachine{59}%
\XRmachine{87}%
\XRmachine{104}%
\XRmachine{134}%
\XRmachine{153}%
\XRmachine{155}%
\XRmachine{175}%
\XRnmach{10}}%
\XRvalue{25}{%
\XRmachine{13}%
\XRnmach{1}}%
\XRvalue{30}{%
\XRmachine{15}%
\XRnmach{1}}%
\XRvalue{32}{%
\XRmachine{5}%
\XRmachine{8}%
\XRmachine{11}%
\XRmachine{30}%
\XRmachine{33}%
\XRmachine{38}%
\XRmachine{52}%
\XRmachine{55}%
\XRmachine{62}%
\XRmachine{63}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{79}%
\XRmachine{82}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{108}%
\XRmachine{113}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{128}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{152}%
\XRmachine{156}%
\XRmachine{165}%
\XRmachine{181}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{190}%
\XRmachine{193}%
\XRnmach{33}}%
\XRvalue{40}{%
\XRmachine{40}%
\XRnmach{1}}%
\XRvalue{44}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{48}{%
\XRmachine{81}%
\XRmachine{122}%
\XRnmach{2}}%
\XRvalue{64}{%
\XRmachine{50}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{187}%
\XRnmach{4}}%
\XRvalue{72}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{128}{%
\XRmachine{4}%
\XRmachine{112}%
\XRmachine{141}%
\XRmachine{162}%
\XRnmach{4}}%
\XRvalue{160}{%
\XRmachine{133}%
\XRnmach{1}}%
\XRvalue{256}{%
\XRmachine{163}%
\XRnmach{1}}%
}%

\XRkey{NONRADIO}{%
\XRvalue{0}{%
\XRmachine{3}%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{10}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{13}%
\XRmachine{14}%
\XRmachine{17}%
\XRmachine{19}%
\XRmachine{21}%
\XRmachine{23}%
\XRmachine{24}%
\XRmachine{25}%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{30}%
\XRmachine{32}%
\XRmachine{34}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{38}%
\XRmachine{41}%
\XRmachine{45}%
\XRmachine{47}%
\XRmachine{52}%
\XRmachine{63}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{69}%
\XRmachine{70}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{79}%
\XRmachine{80}%
\XRmachine{81}%
\XRmachine{86}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{104}%
\XRmachine{105}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{112}%
\XRmachine{117}%
\XRmachine{120}%
\XRmachine{121}%
\XRmachine{124}%
\XRmachine{139}%
\XRmachine{141}%
\XRmachine{143}%
\XRmachine{145}%
\XRmachine{148}%
\XRmachine{149}%
\XRmachine{151}%
\XRmachine{153}%
\XRmachine{155}%
\XRmachine{158}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{165}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{178}%
\XRmachine{181}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{194}%
\XRmachine{196}%
\XRmachine{200}%
\XRmachine{201}%
\XRmachine{204}%
\XRnmach{83}}%
\XRvalue{0.5}{%
\XRmachine{75}%
\XRnmach{1}}%
\XRvalue{1}{%
\XRmachine{26}%
\XRmachine{68}%
\XRmachine{82}%
\XRmachine{195}%
\XRnmach{4}}%
\XRvalue{2}{%
\XRmachine{16}%
\XRmachine{18}%
\XRmachine{76}%
\XRmachine{87}%
\XRnmach{4}}%
\XRvalue{5}{%
\XRmachine{1}%
\XRmachine{22}%
\XRmachine{33}%
\XRmachine{48}%
\XRmachine{53}%
\XRmachine{64}%
\XRmachine{67}%
\XRmachine{74}%
\XRmachine{77}%
\XRmachine{85}%
\XRmachine{88}%
\XRmachine{103}%
\XRmachine{125}%
\XRmachine{154}%
\XRmachine{170}%
\XRmachine{188}%
\XRnmach{16}}%
\XRvalue{6}{%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRnmach{3}}%
\XRvalue{7}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{10}{%
\XRmachine{2}%
\XRmachine{4}%
\XRmachine{9}%
\XRmachine{15}%
\XRmachine{40}%
\XRmachine{43}%
\XRmachine{46}%
\XRmachine{50}%
\XRmachine{51}%
\XRmachine{55}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{60}%
\XRmachine{61}%
\XRmachine{62}%
\XRmachine{71}%
\XRmachine{83}%
\XRmachine{106}%
\XRmachine{119}%
\XRmachine{150}%
\XRmachine{156}%
\XRmachine{159}%
\XRmachine{180}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRnmach{31}}%
\XRvalue{15}{%
\XRmachine{92}%
\XRmachine{94}%
\XRmachine{98}%
\XRmachine{99}%
\XRnmach{4}}%
\XRvalue{20}{%
\XRmachine{59}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{118}%
\XRmachine{122}%
\XRmachine{127}%
\XRmachine{140}%
\XRmachine{162}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{167}%
\XRmachine{179}%
\XRnmach{13}}%
\XRvalue{25}{%
\XRmachine{113}%
\XRmachine{152}%
\XRmachine{190}%
\XRnmach{3}}%
\XRvalue{30}{%
\XRmachine{49}%
\XRmachine{93}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{146}%
\XRmachine{147}%
\XRnmach{6}}%
\XRvalue{33}{%
\XRmachine{116}%
\XRnmach{1}}%
\XRvalue{35}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{40}{%
\XRmachine{20}%
\XRmachine{35}%
\XRmachine{39}%
\XRnmach{3}}%
\XRvalue{50}{%
\XRmachine{8}%
\XRmachine{29}%
\XRmachine{72}%
\XRmachine{84}%
\XRmachine{91}%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{133}%
\XRnmach{8}}%
\XRvalue{55}{%
\XRmachine{44}%
\XRnmach{1}}%
\XRvalue{60}{%
\XRmachine{132}%
\XRnmach{1}}%
\XRvalue{70}{%
\XRmachine{128}%
\XRnmach{1}}%
\XRvalue{75}{%
\XRmachine{123}%
\XRnmach{1}}%
\XRvalue{80}{%
\XRmachine{31}%
\XRmachine{129}%
\XRmachine{142}%
\XRnmach{3}}%
\XRvalue{90}{%
\XRmachine{54}%
\XRmachine{131}%
\XRnmach{2}}%
\XRvalue{100}{%
\XRmachine{42}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{126}%
\XRmachine{130}%
\XRmachine{144}%
\XRmachine{157}%
\XRmachine{177}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{189}%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{13}}%
}%

\XRkey{Net-Service}{%
\XRvalue{ARPA/Internet-service}{%
\XRmachine{20}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{69}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{96}%
\XRmachine{104}%
\XRmachine{105}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{108}%
\XRmachine{111}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{118}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{129}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{140}%
\XRmachine{141}%
\XRmachine{143}%
\XRmachine{144}%
\XRmachine{149}%
\XRmachine{153}%
\XRmachine{154}%
\XRmachine{157}%
\XRmachine{158}%
\XRmachine{159}%
\XRmachine{161}%
\XRmachine{163}%
\XRmachine{163}%
\XRmachine{166}%
\XRmachine{168}%
\XRmachine{170}%
\XRmachine{176}%
\XRmachine{177}%
\XRmachine{178}%
\XRmachine{182}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{187}%
\XRmachine{188}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{194}%
\XRmachine{195}%
\XRmachine{198}%
\XRmachine{201}%
\XRnmach{61}}%
\XRvalue{BITNET/EARN-service}{%
\XRmachine{16}%
\XRmachine{18}%
\XRmachine{21}%
\XRmachine{24}%
\XRmachine{25}%
\XRmachine{32}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{35}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{37}%
\XRmachine{40}%
\XRmachine{48}%
\XRmachine{48}%
\XRmachine{50}%
\XRmachine{56}%
\XRmachine{60}%
\XRmachine{62}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{69}%
\XRmachine{71}%
\XRmachine{73}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{91}%
\XRmachine{99}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{110}%
\XRmachine{114}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{121}%
\XRmachine{128}%
\XRmachine{142}%
\XRmachine{143}%
\XRmachine{143}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{149}%
\XRmachine{151}%
\XRmachine{152}%
\XRmachine{157}%
\XRmachine{167}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{180}%
\XRmachine{185}%
\XRmachine{195}%
\XRmachine{201}%
\XRmachine{203}%
\XRmachine{205}%
\XRnmach{58}}%
\XRvalue{SPAN/HEPNET-service}{%
\XRmachine{5}%
\XRmachine{12}%
\XRmachine{15}%
\XRmachine{35}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{40}%
\XRmachine{49}%
\XRmachine{50}%
\XRmachine{51}%
\XRmachine{60}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{80}%
\XRmachine{90}%
\XRmachine{107}%
\XRmachine{127}%
\XRmachine{128}%
\XRmachine{129}%
\XRmachine{129}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{135}%
\XRmachine{145}%
\XRmachine{149}%
\XRmachine{151}%
\XRmachine{154}%
\XRmachine{157}%
\XRmachine{158}%
\XRmachine{161}%
\XRmachine{168}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{173}%
\XRmachine{173}%
\XRmachine{179}%
\XRmachine{185}%
\XRmachine{189}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{196}%
\XRnmach{48}}%
\XRvalue{UUCP-service}{%
\XRmachine{39}%
\XRnmach{1}}%
\XRvalue{service-on-at-least-one-net}{%
\XRmachine{1}%
\XRmachine{3}%
\XRmachine{4}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{13}%
\XRmachine{14}%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{17}%
\XRmachine{18}%
\XRmachine{19}%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{24}%
\XRmachine{31}%
\XRmachine{32}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{35}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{38}%
\XRmachine{39}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{47}%
\XRmachine{48}%
\XRmachine{49}%
\XRmachine{50}%
\XRmachine{54}%
\XRmachine{55}%
\XRmachine{59}%
\XRmachine{61}%
\XRmachine{63}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{70}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{74}%
\XRmachine{79}%
\XRmachine{81}%
\XRmachine{82}%
\XRmachine{83}%
\XRmachine{85}%
\XRmachine{86}%
\XRmachine{87}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{95}%
\XRmachine{98}%
\XRmachine{103}%
\XRmachine{104}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{116}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{120}%
\XRmachine{122}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{126}%
\XRmachine{127}%
\XRmachine{128}%
\XRmachine{133}%
\XRmachine{134}%
\XRmachine{139}%
\XRmachine{140}%
\XRmachine{141}%
\XRmachine{142}%
\XRmachine{143}%
\XRmachine{144}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{148}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{152}%
\XRmachine{153}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{158}%
\XRmachine{160}%
\XRmachine{162}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{168}%
\XRmachine{169}%
\XRmachine{170}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{174}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{177}%
\XRmachine{178}%
\XRmachine{179}%
\XRmachine{181}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{188}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{194}%
\XRmachine{195}%
\XRmachine{197}%
\XRmachine{200}%
\XRmachine{202}%
\XRmachine{204}%
\XRnmach{130}}%
}%

\XRkey{OS}{%
\XRvalue{Fujitsu}{%
\XRmachine{52}%
\XRmachine{53}%
\XRnmach{2}}%
\XRvalue{UNIX~4.x}{%
\XRmachine{111}%
\XRnmach{1}}%
\XRvalue{UNIX~AIX}{%
\XRmachine{65}%
\XRnmach{1}}%
\XRvalue{UNIX~Concentrix~5.5}{%
\XRmachine{59}%
\XRmachine{62}%
\XRnmach{2}}%
\XRvalue{UNIX~Concentrix~5.5.02}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{UNIX~Concentrix~5.5.03}{%
\XRmachine{87}%
\XRnmach{1}}%
\XRvalue{UNIX~Concentrix~x.x}{%
\XRmachine{81}%
\XRnmach{1}}%
\XRvalue{UNIX~Convex~6.2}{%
\XRmachine{112}%
\XRnmach{1}}%
\XRvalue{UNIX~Convex~7.1}{%
\XRmachine{193}%
\XRnmach{1}}%
\XRvalue{UNIX~Convex~8.1}{%
\XRmachine{4}%
\XRmachine{38}%
\XRmachine{50}%
\XRmachine{55}%
\XRmachine{68}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{113}%
\XRmachine{162}%
\XRnmach{11}}%
\XRvalue{UNIX~Convex~9.0}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{UNIX~Convex~x.x}{%
\XRmachine{82}%
\XRnmach{1}}%
\XRvalue{UNIX~DG~4.3}{%
\XRmachine{21}%
\XRnmach{1}}%
\XRvalue{UNIX~Irix~System~V~3.3.2}{%
\XRmachine{19}%
\XRnmach{1}}%
\XRvalue{UNIX~Stellix~System~V}{%
\XRmachine{152}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~3.5}{%
\XRmachine{115}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~4.0}{%
\XRmachine{10}%
\XRmachine{29}%
\XRmachine{83}%
\XRmachine{169}%
\XRnmach{4}}%
\XRvalue{UNIX~Sun~4.0.1}{%
\XRmachine{20}%
\XRmachine{119}%
\XRmachine{184}%
\XRmachine{185}%
\XRnmach{4}}%
\XRvalue{UNIX~Sun~4.0.2}{%
\XRmachine{16}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~4.0.3}{%
\XRmachine{23}%
\XRmachine{32}%
\XRmachine{47}%
\XRmachine{70}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{109}%
\XRmachine{114}%
\XRmachine{117}%
\XRmachine{140}%
\XRmachine{148}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{194}%
\XRnmach{15}}%
\XRvalue{UNIX~Sun~4.0.3c}{%
\XRmachine{6}%
\XRmachine{9}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{107}%
\XRmachine{118}%
\XRmachine{159}%
\XRnmach{7}}%
\XRvalue{UNIX~Sun~4.1}{%
\XRmachine{1}%
\XRmachine{3}%
\XRmachine{5}%
\XRmachine{8}%
\XRmachine{11}%
\XRmachine{15}%
\XRmachine{25}%
\XRmachine{48}%
\XRmachine{61}%
\XRmachine{67}%
\XRmachine{69}%
\XRmachine{71}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{90}%
\XRmachine{103}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{120}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{153}%
\XRmachine{163}%
\XRmachine{168}%
\XRmachine{178}%
\XRmachine{183}%
\XRmachine{187}%
\XRmachine{190}%
\XRmachine{200}%
\XRnmach{33}}%
\XRvalue{UNIX~Sun~4.1.1}{%
\XRmachine{12}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~4.2}{%
\XRmachine{54}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~4.4.1}{%
\XRmachine{171}%
\XRnmach{1}}%
\XRvalue{UNIX~Sun~x.x}{%
\XRmachine{7}%
\XRmachine{17}%
\XRmachine{18}%
\XRmachine{39}%
\XRmachine{105}%
\XRmachine{127}%
\XRmachine{143}%
\XRmachine{164}%
\XRmachine{177}%
\XRmachine{182}%
\XRmachine{186}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRnmach{14}}%
\XRvalue{UNIX~Ultrix~4.0}{%
\XRmachine{63}%
\XRmachine{66}%
\XRmachine{175}%
\XRnmach{3}}%
\XRvalue{UNIX~UniCOS~5.1}{%
\XRmachine{158}%
\XRnmach{1}}%
\XRvalue{UNIX~x.x}{%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{174}%
\XRnmach{4}}%
\XRvalue{VMS~4.2}{%
\XRmachine{26}%
\XRmachine{45}%
\XRmachine{46}%
\XRnmach{3}}%
\XRvalue{VMS~4.4}{%
\XRmachine{110}%
\XRmachine{204}%
\XRnmach{2}}%
\XRvalue{VMS~4.5}{%
\XRmachine{104}%
\XRmachine{157}%
\XRnmach{2}}%
\XRvalue{VMS~4.6}{%
\XRmachine{14}%
\XRmachine{56}%
\XRmachine{64}%
\XRmachine{196}%
\XRnmach{4}}%
\XRvalue{VMS~4.7}{%
\XRmachine{22}%
\XRmachine{30}%
\XRmachine{34}%
\XRmachine{40}%
\XRmachine{44}%
\XRmachine{76}%
\XRmachine{77}%
\XRmachine{91}%
\XRmachine{125}%
\XRnmach{9}}%
\XRvalue{VMS~5.0}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{VMS~5.0-2}{%
\XRmachine{189}%
\XRnmach{1}}%
\XRvalue{VMS~5.1}{%
\XRmachine{192}%
\XRnmach{1}}%
\XRvalue{VMS~5.1-1}{%
\XRmachine{121}%
\XRnmach{1}}%
\XRvalue{VMS~5.2}{%
\XRmachine{13}%
\XRmachine{36}%
\XRmachine{80}%
\XRmachine{133}%
\XRmachine{176}%
\XRmachine{201}%
\XRnmach{6}}%
\XRvalue{VMS~5.3}{%
\XRmachine{37}%
\XRmachine{51}%
\XRmachine{60}%
\XRmachine{84}%
\XRmachine{88}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{128}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{142}%
\XRmachine{155}%
\XRmachine{165}%
\XRmachine{167}%
\XRmachine{172}%
\XRmachine{173}%
\XRnmach{21}}%
\XRvalue{VMS~5.3-1}{%
\XRmachine{2}%
\XRmachine{85}%
\XRmachine{89}%
\XRmachine{106}%
\XRmachine{123}%
\XRmachine{147}%
\XRmachine{154}%
\XRmachine{156}%
\XRmachine{181}%
\XRnmach{9}}%
\XRvalue{VMS~5.3a}{%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{2}}%
\XRvalue{VMS~5.4}{%
\XRmachine{33}%
\XRmachine{63}%
\XRmachine{79}%
\XRmachine{124}%
\XRmachine{144}%
\XRmachine{149}%
\XRnmach{6}}%
\XRvalue{VMS~x.x}{%
\XRmachine{24}%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{31}%
\XRmachine{35}%
\XRmachine{41}%
\XRmachine{42}%
\XRmachine{43}%
\XRmachine{49}%
\XRmachine{86}%
\XRmachine{108}%
\XRmachine{116}%
\XRmachine{126}%
\XRmachine{139}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{170}%
\XRmachine{179}%
\XRmachine{180}%
\XRmachine{188}%
\XRmachine{191}%
\XRmachine{195}%
\XRnmach{22}}%
}%

\XRkey{PLOT}{%
\XRvalue{?}{%
\XRmachine{3}%
\XRmachine{12}%
\XRmachine{31}%
\XRmachine{37}%
\XRmachine{39}%
\XRmachine{42}%
\XRmachine{47}%
\XRmachine{52}%
\XRmachine{56}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{83}%
\XRmachine{98}%
\XRmachine{103}%
\XRmachine{110}%
\XRmachine{127}%
\XRmachine{141}%
\XRmachine{152}%
\XRmachine{157}%
\XRmachine{175}%
\XRmachine{178}%
\XRmachine{181}%
\XRmachine{182}%
\XRmachine{191}%
\XRmachine{204}%
\XRnmach{25}}%
\XRvalue{Apple~Laserwriter}{%
\XRmachine{11}%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{48}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{81}%
\XRmachine{113}%
\XRmachine{117}%
\XRmachine{133}%
\XRmachine{158}%
\XRmachine{159}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{162}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{186}%
\XRmachine{193}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{28}}%
\XRvalue{Apple~Laserwriter+}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{34}%
\XRmachine{49}%
\XRmachine{168}%
\XRmachine{189}%
\XRnmach{6}}%
\XRvalue{Apple~Laserwriter~II}{%
\XRmachine{17}%
\XRmachine{25}%
\XRmachine{33}%
\XRmachine{54}%
\XRmachine{91}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{169}%
\XRmachine{190}%
\XRmachine{192}%
\XRnmach{12}}%
\XRvalue{Apple~Laserwriter~IIx}{%
\XRmachine{183}%
\XRnmach{1}}%
\XRvalue{Apple~Laserwriter~II~NT}{%
\XRmachine{13}%
\XRnmach{1}}%
\XRvalue{Apple~Laserwriter~II~NTX}{%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRnmach{3}}%
\XRvalue{Apple~NT..}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{Benson~1113}{%
\XRmachine{33}%
\XRnmach{1}}%
\XRvalue{CANPB~II}{%
\XRmachine{27}%
\XRmachine{28}%
\XRnmach{2}}%
\XRvalue{Calcomp}{%
\XRmachine{43}%
\XRnmach{1}}%
\XRvalue{Calcomp~81}{%
\XRmachine{60}%
\XRnmach{1}}%
\XRvalue{Canon~LN03}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{Canon~LP68~II}{%
\XRmachine{86}%
\XRnmach{1}}%
\XRvalue{Canon~Mark~II}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{Canon~laser}{%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{88}%
\XRnmach{3}}%
\XRvalue{Citoh}{%
\XRmachine{126}%
\XRnmach{1}}%
\XRvalue{DEC~LN01}{%
\XRmachine{108}%
\XRnmach{1}}%
\XRvalue{DEC~LN03}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{DEC~LN03+}{%
\XRmachine{79}%
\XRmachine{80}%
\XRmachine{156}%
\XRnmach{3}}%
\XRvalue{DEC~LN03+~(Postscript)}{%
\XRmachine{32}%
\XRmachine{142}%
\XRnmach{2}}%
\XRvalue{DEC~LN03R}{%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{156}%
\XRnmach{3}}%
\XRvalue{DEC~Scriptprinter}{%
\XRmachine{63}%
\XRnmach{1}}%
\XRvalue{Dataproducts~Laserwriter}{%
\XRmachine{4}%
\XRmachine{5}%
\XRmachine{8}%
\XRmachine{10}%
\XRnmach{4}}%
\XRvalue{Digital~(across~network)}{%
\XRmachine{32}%
\XRnmach{1}}%
\XRvalue{Facom~6715E}{%
\XRmachine{53}%
\XRnmach{1}}%
\XRvalue{HI~DMP29}{%
\XRmachine{165}%
\XRnmach{1}}%
\XRvalue{HP~7475}{%
\XRmachine{149}%
\XRmachine{155}%
\XRnmach{2}}%
\XRvalue{HP~7475A}{%
\XRmachine{91}%
\XRmachine{165}%
\XRnmach{2}}%
\XRvalue{HP~7475B}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{HP~7550}{%
\XRmachine{23}%
\XRnmach{1}}%
\XRvalue{HP~7550A}{%
\XRmachine{34}%
\XRmachine{150}%
\XRnmach{2}}%
\XRvalue{HP~GL}{%
\XRmachine{148}%
\XRnmach{1}}%
\XRvalue{HP~LaserJet}{%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRnmach{3}}%
\XRvalue{HP~LaserJet+}{%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{174}%
\XRnmach{4}}%
\XRvalue{HP~LaserJet~(w/PS)}{%
\XRmachine{109}%
\XRnmach{1}}%
\XRvalue{HP~LaserJet~II}{%
\XRmachine{41}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{105}%
\XRmachine{120}%
\XRmachine{121}%
\XRmachine{153}%
\XRmachine{154}%
\XRmachine{155}%
\XRmachine{179}%
\XRmachine{180}%
\XRnmach{12}}%
\XRvalue{HP~LaserJet~III}{%
\XRmachine{21}%
\XRmachine{55}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{171}%
\XRnmach{6}}%
\XRvalue{HP~LaserJet~III~(w/PS)}{%
\XRmachine{187}%
\XRnmach{1}}%
\XRvalue{HP~LaserJet~IIP}{%
\XRmachine{45}%
\XRmachine{46}%
\XRnmach{2}}%
\XRvalue{IBM~4019}{%
\XRmachine{74}%
\XRmachine{75}%
\XRnmach{2}}%
\XRvalue{Imagen~2308S}{%
\XRmachine{201}%
\XRnmach{1}}%
\XRvalue{Imagen~8/300}{%
\XRmachine{112}%
\XRmachine{188}%
\XRmachine{195}%
\XRmachine{196}%
\XRnmach{4}}%
\XRvalue{Imagen~Innovator}{%
\XRmachine{18}%
\XRmachine{150}%
\XRnmach{2}}%
\XRvalue{Imagen~x/yyy}{%
\XRmachine{111}%
\XRnmach{1}}%
\XRvalue{KS~LP28XT}{%
\XRmachine{30}%
\XRnmach{1}}%
\XRvalue{LBP-8II}{%
\XRmachine{26}%
\XRnmach{1}}%
\XRvalue{Mitsubishi~Color}{%
\XRmachine{4}%
\XRmachine{5}%
\XRmachine{8}%
\XRmachine{10}%
\XRnmach{4}}%
\XRvalue{NEC~LC-800}{%
\XRmachine{70}%
\XRnmach{1}}%
\XRvalue{NEC~LC890~(Postscript)}{%
\XRmachine{140}%
\XRnmach{1}}%
\XRvalue{NEC~Silentwriter}{%
\XRmachine{4}%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{10}%
\XRnmach{7}}%
\XRvalue{Omnilaser~(Postscript)}{%
\XRmachine{114}%
\XRnmach{1}}%
\XRvalue{Omnilaser~2108}{%
\XRmachine{29}%
\XRmachine{84}%
\XRnmach{2}}%
\XRvalue{Pericom~Video~Printer}{%
\XRmachine{139}%
\XRnmach{1}}%
\XRvalue{Postscript}{%
\XRmachine{51}%
\XRmachine{70}%
\XRmachine{128}%
\XRmachine{132}%
\XRmachine{143}%
\XRmachine{148}%
\XRmachine{149}%
\XRmachine{194}%
\XRnmach{8}}%
\XRvalue{Printronix~300}{%
\XRmachine{49}%
\XRmachine{156}%
\XRnmach{2}}%
\XRvalue{Printronix~4160}{%
\XRmachine{46}%
\XRnmach{1}}%
\XRvalue{Printronix~600}{%
\XRmachine{44}%
\XRmachine{45}%
\XRnmach{2}}%
\XRvalue{Printronix~MVP}{%
\XRmachine{188}%
\XRnmach{1}}%
\XRvalue{Printronix~xxx}{%
\XRmachine{34}%
\XRmachine{85}%
\XRmachine{108}%
\XRmachine{112}%
\XRmachine{151}%
\XRmachine{193}%
\XRnmach{6}}%
\XRvalue{QMS~1200}{%
\XRmachine{137}%
\XRmachine{138}%
\XRnmach{2}}%
\XRvalue{QMS~1500}{%
\XRmachine{72}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{76}%
\XRnmach{4}}%
\XRvalue{QMS~1500~PS}{%
\XRmachine{74}%
\XRmachine{75}%
\XRnmach{2}}%
\XRvalue{QMS~2400}{%
\XRmachine{129}%
\XRnmach{1}}%
\XRvalue{QMS~800}{%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{24}%
\XRmachine{38}%
\XRmachine{40}%
\XRmachine{46}%
\XRmachine{50}%
\XRmachine{59}%
\XRmachine{60}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{67}%
\XRmachine{68}%
\XRmachine{69}%
\XRmachine{71}%
\XRmachine{77}%
\XRmachine{82}%
\XRmachine{90}%
\XRmachine{99}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{116}%
\XRmachine{128}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{144}%
\XRmachine{170}%
\XRmachine{179}%
\XRmachine{180}%
\XRnmach{33}}%
\XRvalue{QMS~800C}{%
\XRmachine{139}%
\XRnmach{1}}%
\XRvalue{QMS~810}{%
\XRmachine{107}%
\XRmachine{144}%
\XRmachine{200}%
\XRnmach{3}}%
\XRvalue{QMS~Lasergrafix}{%
\XRmachine{104}%
\XRmachine{176}%
\XRnmach{2}}%
\XRvalue{QMS~PS-800}{%
\XRmachine{65}%
\XRmachine{66}%
\XRmachine{67}%
\XRmachine{68}%
\XRmachine{69}%
\XRnmach{5}}%
\XRvalue{QMS~PS-800~II}{%
\XRmachine{167}%
\XRnmach{1}}%
\XRvalue{QMS~PS-810}{%
\XRmachine{19}%
\XRmachine{61}%
\XRmachine{62}%
\XRnmach{3}}%
\XRvalue{QMS~PS-820}{%
\XRmachine{106}%
\XRmachine{107}%
\XRnmach{2}}%
\XRvalue{QMS~PS-xxx}{%
\XRmachine{90}%
\XRnmach{1}}%
\XRvalue{QMS~Talaris~1200}{%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{2}}%
\XRvalue{QMS~Talaris~1590}{%
\XRmachine{124}%
\XRnmach{1}}%
\XRvalue{QMS~Talaris~1590-T}{%
\XRmachine{106}%
\XRmachine{122}%
\XRmachine{123}%
\XRmachine{125}%
\XRmachine{133}%
\XRmachine{147}%
\XRnmach{6}}%
\XRvalue{QMS~Talaris~800}{%
\XRmachine{126}%
\XRmachine{130}%
\XRmachine{172}%
\XRmachine{173}%
\XRnmach{4}}%
\XRvalue{QMS~Talaris~xxx}{%
\XRmachine{151}%
\XRmachine{172}%
\XRmachine{173}%
\XRnmach{3}}%
\XRvalue{SPARCprinter}{%
\XRmachine{171}%
\XRnmach{1}}%
\XRvalue{TI~2105~(Postscript)}{%
\XRmachine{115}%
\XRnmach{1}}%
\XRvalue{TI~2108}{%
\XRmachine{82}%
\XRnmach{1}}%
\XRvalue{TI~2108~(Postscript)}{%
\XRmachine{177}%
\XRnmach{1}}%
\XRvalue{TI~2115~laser}{%
\XRmachine{192}%
\XRnmach{1}}%
\XRvalue{TI~8250}{%
\XRmachine{114}%
\XRnmach{1}}%
\XRvalue{Tektronix~4010}{%
\XRmachine{189}%
\XRnmach{1}}%
\XRvalue{Versatec~1200}{%
\XRmachine{22}%
\XRnmach{1}}%
\XRvalue{Versatec~1200A}{%
\XRmachine{64}%
\XRmachine{68}%
\XRmachine{69}%
\XRnmach{3}}%
\XRvalue{Versatec~V80}{%
\XRmachine{35}%
\XRmachine{36}%
\XRmachine{131}%
\XRnmach{3}}%
\XRvalue{Versatec~xxxx}{%
\XRmachine{40}%
\XRmachine{86}%
\XRmachine{112}%
\XRmachine{195}%
\XRmachine{196}%
\XRnmach{5}}%
\XRvalue{Zeta}{%
\XRmachine{148}%
\XRnmach{1}}%
}%

\XRkey{SITE}{%
\XRvalue{Arcetri~Obs.}{%
\XRmachine{48}%
\XRnmach{1}}%
\XRvalue{Arecibo~Obs.}{%
\XRmachine{90}%
\XRnmach{1}}%
\XRvalue{Arizona~St.~U.}{%
\XRmachine{175}%
\XRnmach{1}}%
\XRvalue{Australia~Telescope}{%
\XRmachine{4}%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{10}%
\XRnmach{7}}%
\XRvalue{Beijing~Obs.}{%
\XRmachine{25}%
\XRmachine{26}%
\XRnmach{2}}%
\XRvalue{Beijing~U.~Geophys.}{%
\XRmachine{27}%
\XRmachine{28}%
\XRnmach{2}}%
\XRvalue{Bell~Labs.~Murray~Hill}{%
\XRmachine{111}%
\XRnmach{1}}%
\XRvalue{Bologna~IRA}{%
\XRmachine{50}%
\XRnmach{1}}%
\XRvalue{Bombay~IR}{%
\XRmachine{43}%
\XRnmach{1}}%
\XRvalue{Boston~U.}{%
\XRmachine{106}%
\XRmachine{107}%
\XRnmach{2}}%
\XRvalue{Brandeis~U.}{%
\XRmachine{108}%
\XRnmach{1}}%
\XRvalue{CIT~Astronomy}{%
\XRmachine{193}%
\XRnmach{1}}%
\XRvalue{CIT~Plan.~Sci.}{%
\XRmachine{192}%
\XRnmach{1}}%
\XRvalue{Columbia~U.~Astron.}{%
\XRmachine{113}%
\XRnmach{1}}%
\XRvalue{Cornell~U.}{%
\XRmachine{116}%
\XRnmach{1}}%
\XRvalue{DRAO~Penticton}{%
\XRmachine{22}%
\XRnmach{1}}%
\XRvalue{ESA/ESTEC-SA}{%
\XRmachine{63}%
\XRnmach{1}}%
\XRvalue{ESO~Garching}{%
\XRmachine{39}%
\XRmachine{40}%
\XRnmach{2}}%
\XRvalue{Fomi~Sat.~Geod.~Obs.}{%
\XRmachine{41}%
\XRnmach{1}}%
\XRvalue{HIA~Ottawa}{%
\XRmachine{16}%
\XRnmach{1}}%
\XRvalue{Haystack~Obs.}{%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRnmach{3}}%
\XRvalue{IIA~Bangalore}{%
\XRmachine{44}%
\XRnmach{1}}%
\XRvalue{INPE-DAS~Brazil}{%
\XRmachine{14}%
\XRnmach{1}}%
\XRvalue{IRAM}{%
\XRmachine{34}%
\XRnmach{1}}%
\XRvalue{Indiana~U.}{%
\XRmachine{148}%
\XRmachine{149}%
\XRnmach{2}}%
\XRvalue{Inst.~Ap.~Paris}{%
\XRmachine{35}%
\XRnmach{1}}%
\XRvalue{Inst.~Ast.~Andalucia}{%
\XRmachine{79}%
\XRmachine{80}%
\XRnmach{2}}%
\XRvalue{Iowa~St.~U.}{%
\XRmachine{152}%
\XRnmach{1}}%
\XRvalue{JPL~Optical~Interferometry}{%
\XRmachine{187}%
\XRnmach{1}}%
\XRvalue{JPL~Planetary~Astronomy}{%
\XRmachine{191}%
\XRnmach{1}}%
\XRvalue{JPL~RAG}{%
\XRmachine{189}%
\XRnmach{1}}%
\XRvalue{JPL~Section~335}{%
\XRmachine{190}%
\XRnmach{1}}%
\XRvalue{Jack~Carmody}{%
\XRmachine{110}%
\XRnmach{1}}%
\XRvalue{Jodrell~Bank}{%
\XRmachine{87}%
\XRmachine{88}%
\XRnmach{2}}%
\XRvalue{Kapteyn~Institute}{%
\XRmachine{61}%
\XRmachine{62}%
\XRnmach{2}}%
\XRvalue{Kobe~U.~Gamma~Ray}{%
\XRmachine{54}%
\XRnmach{1}}%
\XRvalue{LANL}{%
\XRmachine{181}%
\XRnmach{1}}%
\XRvalue{LLNL-IGPP}{%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRnmach{3}}%
\XRvalue{Lafayette~Coll.}{%
\XRmachine{120}%
\XRmachine{121}%
\XRnmach{2}}%
\XRvalue{Leiden~Obs.}{%
\XRmachine{55}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRnmach{4}}%
\XRvalue{Lowell~Obs.}{%
\XRmachine{178}%
\XRnmach{1}}%
\XRvalue{MIT~Phys.}{%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{103}%
\XRnmach{4}}%
\XRvalue{MIT~Radio~Astron.}{%
\XRmachine{104}%
\XRnmach{1}}%
\XRvalue{MPIfR~Bonn}{%
\XRmachine{38}%
\XRnmach{1}}%
\XRvalue{MRAO}{%
\XRmachine{83}%
\XRmachine{84}%
\XRnmach{2}}%
\XRvalue{MacQuarie~U.}{%
\XRmachine{3}%
\XRnmach{1}}%
\XRvalue{Mallinckrodt~Inst.~Radiology}{%
\XRmachine{165}%
\XRnmach{1}}%
\XRvalue{Marseille~Obs.}{%
\XRmachine{32}%
\XRnmach{1}}%
\XRvalue{Michigan~St.~U.}{%
\XRmachine{151}%
\XRnmach{1}}%
\XRvalue{Mt.~Stromlo~Obs.}{%
\XRmachine{11}%
\XRnmach{1}}%
\XRvalue{NASA-Ames}{%
\XRmachine{195}%
\XRmachine{196}%
\XRnmach{2}}%
\XRvalue{NASA-GSFC~Ast/Solar}{%
\XRmachine{128}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRnmach{5}}%
\XRvalue{NASA-GSFC~SDCDCL}{%
\XRmachine{127}%
\XRnmach{1}}%
\XRvalue{NASA-JPL~XENOS}{%
\XRmachine{188}%
\XRnmach{1}}%
\XRvalue{NASA-Marshall~SFC}{%
\XRmachine{144}%
\XRnmach{1}}%
\XRvalue{NASA/GSFC~SDCD}{%
\XRmachine{133}%
\XRnmach{1}}%
\XRvalue{NC~St.~U.~Phys.}{%
\XRmachine{142}%
\XRnmach{1}}%
\XRvalue{NFRA~Dwingeloo}{%
\XRmachine{59}%
\XRmachine{60}%
\XRnmach{2}}%
\XRvalue{NOTO~IRA}{%
\XRmachine{51}%
\XRnmach{1}}%
\XRvalue{NRAO~AOC~SPARC2}{%
\XRmachine{72}%
\XRnmach{1}}%
\XRvalue{NRAO~CHOLLA}{%
\XRmachine{75}%
\XRnmach{1}}%
\XRvalue{NRAO~COLOBUS}{%
\XRmachine{66}%
\XRnmach{1}}%
\XRvalue{NRAO~CVAX}{%
\XRmachine{64}%
\XRnmach{1}}%
\XRvalue{NRAO~CV~AIPS~GROUP}{%
\XRmachine{69}%
\XRnmach{1}}%
\XRvalue{NRAO~GB~SUN}{%
\XRmachine{70}%
\XRnmach{1}}%
\XRvalue{NRAO~KILLIANS}{%
\XRmachine{71}%
\XRnmach{1}}%
\XRvalue{NRAO~KONG}{%
\XRmachine{73}%
\XRnmach{1}}%
\XRvalue{NRAO~LEMUR}{%
\XRmachine{65}%
\XRnmach{1}}%
\XRvalue{NRAO~MANDRILL}{%
\XRmachine{67}%
\XRnmach{1}}%
\XRvalue{NRAO~NEGRA}{%
\XRmachine{78}%
\XRnmach{1}}%
\XRvalue{NRAO~NRAO1}{%
\XRmachine{68}%
\XRnmach{1}}%
\XRvalue{NRAO~OUTBAX}{%
\XRmachine{77}%
\XRnmach{1}}%
\XRvalue{NRAO~VAX1}{%
\XRmachine{76}%
\XRnmach{1}}%
\XRvalue{NRAO~YUCCA}{%
\XRmachine{74}%
\XRnmach{1}}%
\XRvalue{Nanjing~U.~CSTDAC}{%
\XRmachine{29}%
\XRmachine{30}%
\XRnmach{2}}%
\XRvalue{Nav.~Res.~Lab.~Radio}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{Nav.~Res.~Lab.~X-Ray}{%
\XRmachine{123}%
\XRnmach{1}}%
\XRvalue{New~Mexico~St.~U.}{%
\XRmachine{183}%
\XRnmach{1}}%
\XRvalue{New~Mexico~Tech~AOC}{%
\XRmachine{182}%
\XRnmach{1}}%
\XRvalue{Nobeyama~Radio~Obs.}{%
\XRmachine{52}%
\XRmachine{53}%
\XRnmach{2}}%
\XRvalue{Northwestern~U.}{%
\XRmachine{160}%
\XRmachine{161}%
\XRnmach{2}}%
\XRvalue{Obs.~de~Meudon}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{Ohio~St.~U.}{%
\XRmachine{146}%
\XRmachine{147}%
\XRnmach{2}}%
\XRvalue{Onsala~Space~Obs.}{%
\XRmachine{82}%
\XRnmach{1}}%
\XRvalue{Ooty~Obs.}{%
\XRmachine{46}%
\XRnmach{1}}%
\XRvalue{PRL~Ahmedabad}{%
\XRmachine{42}%
\XRnmach{1}}%
\XRvalue{Penn~State~U.}{%
\XRmachine{118}%
\XRmachine{119}%
\XRnmach{2}}%
\XRvalue{Phys.~Inst.~Koeln}{%
\XRmachine{37}%
\XRnmach{1}}%
\XRvalue{Princeton~U.~Obs.}{%
\XRmachine{112}%
\XRnmach{1}}%
\XRvalue{Queen's~U.}{%
\XRmachine{17}%
\XRnmach{1}}%
\XRvalue{R.~Greenwich~Obs.}{%
\XRmachine{85}%
\XRnmach{1}}%
\XRvalue{R.~Obs.~Edinburgh}{%
\XRmachine{86}%
\XRnmach{1}}%
\XRvalue{Raman~Res.~Inst.}{%
\XRmachine{45}%
\XRnmach{1}}%
\XRvalue{Rensselaer~P.~Inst.}{%
\XRmachine{114}%
\XRmachine{115}%
\XRnmach{2}}%
\XRvalue{S.~T.~Sci.~Inst.}{%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{137}%
\XRmachine{138}%
\XRnmach{5}}%
\XRvalue{SAO~Cambridge}{%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRnmach{5}}%
\XRvalue{ST~Systems~Corp.}{%
\XRmachine{126}%
\XRnmach{1}}%
\XRvalue{Steward~Obs.}{%
\XRmachine{176}%
\XRmachine{177}%
\XRnmach{2}}%
\XRvalue{Stockholm~Obs.}{%
\XRmachine{81}%
\XRnmach{1}}%
\XRvalue{Tech.~U.~Denmark}{%
\XRmachine{31}%
\XRnmach{1}}%
\XRvalue{Tel~Aviv~U.}{%
\XRmachine{47}%
\XRnmach{1}}%
\XRvalue{Trinity~U.}{%
\XRmachine{168}%
\XRnmach{1}}%
\XRvalue{Tufts~U.}{%
\XRmachine{105}%
\XRnmach{1}}%
\XRvalue{U.S.~Naval~Obs.}{%
\XRmachine{124}%
\XRmachine{125}%
\XRnmach{2}}%
\XRvalue{U.~Bordeaux}{%
\XRmachine{33}%
\XRnmach{1}}%
\XRvalue{U.~Calgary}{%
\XRmachine{20}%
\XRmachine{21}%
\XRnmach{2}}%
\XRvalue{U.~Catania}{%
\XRmachine{49}%
\XRnmach{1}}%
\XRvalue{U.~Colorado~APAS}{%
\XRmachine{170}%
\XRmachine{171}%
\XRnmach{2}}%
\XRvalue{U.~Colorado~CASA}{%
\XRmachine{172}%
\XRmachine{173}%
\XRnmach{2}}%
\XRvalue{U.~Florida}{%
\XRmachine{143}%
\XRnmach{1}}%
\XRvalue{U.~Illinois~Astron.}{%
\XRmachine{163}%
\XRmachine{164}%
\XRnmach{2}}%
\XRvalue{U.~Illinois~CSO}{%
\XRmachine{162}%
\XRnmach{1}}%
\XRvalue{U.~Iowa}{%
\XRmachine{153}%
\XRmachine{154}%
\XRmachine{155}%
\XRnmach{3}}%
\XRvalue{U.~Ky.~Astron.}{%
\XRmachine{145}%
\XRnmach{1}}%
\XRvalue{U.~Maine~Astron.}{%
\XRmachine{109}%
\XRnmach{1}}%
\XRvalue{U.~Mass.}{%
\XRmachine{91}%
\XRnmach{1}}%
\XRvalue{U.~Michigan}{%
\XRmachine{150}%
\XRnmach{1}}%
\XRvalue{U.~Minnesota}{%
\XRmachine{158}%
\XRmachine{159}%
\XRnmach{2}}%
\XRvalue{U.~Missouri}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{U.~Montreal}{%
\XRmachine{15}%
\XRnmach{1}}%
\XRvalue{U.~N.~Carolina}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{U.~New~Mexico}{%
\XRmachine{179}%
\XRmachine{180}%
\XRnmach{2}}%
\XRvalue{U.~Oklahoma}{%
\XRmachine{167}%
\XRnmach{1}}%
\XRvalue{U.~Oxford~Astrophys.}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{U.~Pittsburgh~Astron.}{%
\XRmachine{117}%
\XRnmach{1}}%
\XRvalue{U.~Sydney}{%
\XRmachine{1}%
\XRmachine{2}%
\XRnmach{2}}%
\XRvalue{U.~Tasmania~Phys.}{%
\XRmachine{13}%
\XRnmach{1}}%
\XRvalue{U.~Texas~Austin}{%
\XRmachine{169}%
\XRnmach{1}}%
\XRvalue{U.~Toronto}{%
\XRmachine{18}%
\XRnmach{1}}%
\XRvalue{U.~Virginia}{%
\XRmachine{139}%
\XRnmach{1}}%
\XRvalue{U.~W.~Aust.~Gravity~Lab}{%
\XRmachine{12}%
\XRnmach{1}}%
\XRvalue{U.~Washington}{%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{2}}%
\XRvalue{U.~Waterloo~Phys.}{%
\XRmachine{19}%
\XRnmach{1}}%
\XRvalue{U.~Wisc.~Space~Phys.}{%
\XRmachine{157}%
\XRnmach{1}}%
\XRvalue{U.~Wisc.~Washburn~Obs.}{%
\XRmachine{156}%
\XRnmach{1}}%
\XRvalue{U.~Wyoming~IR~Obs.}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{UBC~Astron.}{%
\XRmachine{24}%
\XRnmach{1}}%
\XRvalue{UBC~Phys.}{%
\XRmachine{23}%
\XRnmach{1}}%
\XRvalue{UC~Berkeley}{%
\XRmachine{200}%
\XRmachine{201}%
\XRnmach{2}}%
\XRvalue{UC~Los~Angeles}{%
\XRmachine{184}%
\XRmachine{185}%
\XRnmach{2}}%
\XRvalue{UC~Santa~Barbara}{%
\XRmachine{194}%
\XRnmach{1}}%
\XRvalue{USC~Psychology}{%
\XRmachine{186}%
\XRnmach{1}}%
\XRvalue{Va.~Tech.~Astron.}{%
\XRmachine{140}%
\XRnmach{1}}%
\XRvalue{Washington~St.~U.}{%
\XRmachine{204}%
\XRnmach{1}}%
}%

\XRkey{TAPE}{%
\XRvalue{0.25"~cartridge}{%
\XRmachine{114}%
\XRnmach{1}}%
\XRvalue{1600bpi}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{4}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{10}%
\XRmachine{11}%
\XRmachine{13}%
\XRmachine{14}%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{19}%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{26}%
\XRmachine{29}%
\XRmachine{30}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{38}%
\XRmachine{39}%
\XRmachine{40}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{46}%
\XRmachine{50}%
\XRmachine{55}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{59}%
\XRmachine{60}%
\XRmachine{62}%
\XRmachine{63}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{76}%
\XRmachine{77}%
\XRmachine{80}%
\XRmachine{81}%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{88}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{111}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{121}%
\XRmachine{122}%
\XRmachine{123}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{128}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{133}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{141}%
\XRmachine{142}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{152}%
\XRmachine{155}%
\XRmachine{156}%
\XRmachine{159}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{162}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{169}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{175}%
\XRmachine{176}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{194}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRmachine{200}%
\XRmachine{201}%
\XRmachine{202}%
\XRmachine{204}%
\XRnmach{118}}%
\XRvalue{3200bpi}{%
\XRmachine{125}%
\XRmachine{175}%
\XRnmach{2}}%
\XRvalue{6250bpi}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{4}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{10}%
\XRmachine{11}%
\XRmachine{14}%
\XRmachine{15}%
\XRmachine{16}%
\XRmachine{19}%
\XRmachine{20}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{29}%
\XRmachine{30}%
\XRmachine{33}%
\XRmachine{34}%
\XRmachine{36}%
\XRmachine{37}%
\XRmachine{38}%
\XRmachine{39}%
\XRmachine{40}%
\XRmachine{44}%
\XRmachine{45}%
\XRmachine{48}%
\XRmachine{50}%
\XRmachine{51}%
\XRmachine{55}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{59}%
\XRmachine{60}%
\XRmachine{62}%
\XRmachine{63}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{70}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{76}%
\XRmachine{77}%
\XRmachine{79}%
\XRmachine{81}%
\XRmachine{83}%
\XRmachine{84}%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{88}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{91}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{102}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{112}%
\XRmachine{113}%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{122}%
\XRmachine{123}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{128}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{133}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{141}%
\XRmachine{144}%
\XRmachine{145}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{152}%
\XRmachine{155}%
\XRmachine{156}%
\XRmachine{157}%
\XRmachine{158}%
\XRmachine{159}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{162}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{169}%
\XRmachine{171}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{175}%
\XRmachine{178}%
\XRmachine{181}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{189}%
\XRmachine{190}%
\XRmachine{191}%
\XRmachine{192}%
\XRmachine{193}%
\XRmachine{194}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRmachine{200}%
\XRmachine{201}%
\XRmachine{202}%
\XRnmach{128}}%
\XRvalue{6250~bpi~across~network}{%
\XRmachine{32}%
\XRnmach{1}}%
\XRvalue{800bpi}{%
\XRmachine{1}%
\XRmachine{2}%
\XRmachine{11}%
\XRmachine{20}%
\XRmachine{26}%
\XRmachine{36}%
\XRmachine{50}%
\XRmachine{55}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{59}%
\XRmachine{60}%
\XRmachine{64}%
\XRmachine{65}%
\XRmachine{68}%
\XRmachine{81}%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{88}%
\XRmachine{91}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{123}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{156}%
\XRmachine{166}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{175}%
\XRmachine{189}%
\XRmachine{192}%
\XRmachine{201}%
\XRmachine{202}%
\XRnmach{38}}%
\XRvalue{8mm}{%
\XRmachine{124}%
\XRnmach{1}}%
\XRvalue{?}{%
\XRmachine{18}%
\XRmachine{21}%
\XRmachine{24}%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{31}%
\XRmachine{35}%
\XRmachine{41}%
\XRmachine{42}%
\XRmachine{43}%
\XRmachine{49}%
\XRmachine{52}%
\XRmachine{53}%
\XRmachine{61}%
\XRmachine{72}%
\XRmachine{73}%
\XRmachine{78}%
\XRmachine{82}%
\XRmachine{86}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{104}%
\XRmachine{108}%
\XRmachine{116}%
\XRmachine{126}%
\XRmachine{139}%
\XRmachine{150}%
\XRmachine{151}%
\XRmachine{170}%
\XRmachine{174}%
\XRmachine{177}%
\XRmachine{179}%
\XRmachine{180}%
\XRmachine{182}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{188}%
\XRmachine{195}%
\XRmachine{196}%
\XRmachine{203}%
\XRnmach{41}}%
\XRvalue{CD-ROM}{%
\XRmachine{171}%
\XRnmach{1}}%
\XRvalue{CIFER}{%
\XRmachine{89}%
\XRnmach{1}}%
\XRvalue{DAT}{%
\XRmachine{62}%
\XRmachine{118}%
\XRmachine{119}%
\XRnmach{3}}%
\XRvalue{Exabyte}{%
\XRmachine{1}%
\XRmachine{3}%
\XRmachine{4}%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{9}%
\XRmachine{10}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{22}%
\XRmachine{23}%
\XRmachine{32}%
\XRmachine{39}%
\XRmachine{47}%
\XRmachine{65}%
\XRmachine{83}%
\XRmachine{84}%
\XRmachine{89}%
\XRmachine{103}%
\XRmachine{113}%
\XRmachine{117}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{120}%
\XRmachine{127}%
\XRmachine{128}%
\XRmachine{129}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRmachine{133}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{140}%
\XRmachine{154}%
\XRmachine{158}%
\XRmachine{159}%
\XRmachine{171}%
\XRmachine{183}%
\XRmachine{193}%
\XRmachine{200}%
\XRnmach{43}}%
\XRvalue{Kennedy~9610}{%
\XRmachine{166}%
\XRnmach{1}}%
\XRvalue{QIC11}{%
\XRmachine{67}%
\XRmachine{69}%
\XRmachine{127}%
\XRnmach{3}}%
\XRvalue{QIC150}{%
\XRmachine{4}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{10}%
\XRmachine{25}%
\XRmachine{71}%
\XRmachine{168}%
\XRmachine{190}%
\XRmachine{194}%
\XRnmach{11}}%
\XRvalue{QIC24}{%
\XRmachine{1}%
\XRmachine{11}%
\XRmachine{17}%
\XRmachine{32}%
\XRmachine{54}%
\XRmachine{67}%
\XRmachine{69}%
\XRmachine{70}%
\XRmachine{71}%
\XRmachine{83}%
\XRmachine{105}%
\XRmachine{109}%
\XRmachine{118}%
\XRmachine{119}%
\XRmachine{120}%
\XRmachine{127}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{140}%
\XRmachine{143}%
\XRmachine{148}%
\XRmachine{149}%
\XRmachine{153}%
\XRmachine{168}%
\XRmachine{190}%
\XRmachine{194}%
\XRnmach{27}}%
\XRvalue{TK50}{%
\XRmachine{66}%
\XRmachine{106}%
\XRmachine{142}%
\XRmachine{154}%
\XRmachine{167}%
\XRnmach{5}}%
\XRvalue{TK70}{%
\XRmachine{63}%
\XRnmach{1}}%
}%

\XRkey{TV}{%
\XRvalue{?}{%
\XRmachine{4}%
\XRmachine{14}%
\XRmachine{37}%
\XRmachine{41}%
\XRmachine{43}%
\XRmachine{45}%
\XRmachine{52}%
\XRmachine{55}%
\XRmachine{79}%
\XRmachine{80}%
\XRmachine{98}%
\XRmachine{102}%
\XRmachine{104}%
\XRmachine{108}%
\XRmachine{110}%
\XRmachine{116}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{179}%
\XRmachine{181}%
\XRmachine{188}%
\XRmachine{189}%
\XRnmach{22}}%
\XRvalue{AED~512}{%
\XRmachine{2}%
\XRnmach{1}}%
\XRvalue{AGS}{%
\XRmachine{27}%
\XRmachine{28}%
\XRnmach{2}}%
\XRvalue{Alliant~(X-Windows)}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{Amiga~500}{%
\XRmachine{121}%
\XRnmach{1}}%
\XRvalue{Apollo~DN3000~(1-bit)}{%
\XRmachine{95}%
\XRnmach{1}}%
\XRvalue{Apollo~DN320~(1-bit)}{%
\XRmachine{96}%
\XRnmach{1}}%
\XRvalue{Apollo~DN560}{%
\XRmachine{87}%
\XRnmach{1}}%
\XRvalue{Apollo~DN660~(8-bit)}{%
\XRmachine{97}%
\XRnmach{1}}%
\XRvalue{COMTAL~Vision~One}{%
\XRmachine{44}%
\XRnmach{1}}%
\XRvalue{DECstation}{%
\XRmachine{56}%
\XRmachine{107}%
\XRnmach{2}}%
\XRvalue{DECstations}{%
\XRmachine{62}%
\XRnmach{1}}%
\XRvalue{DECstation~(SAO~Image)}{%
\XRmachine{175}%
\XRnmach{1}}%
\XRvalue{DECstation~(XAS)}{%
\XRmachine{59}%
\XRnmach{1}}%
\XRvalue{DECstation~color~(8-bit)}{%
\XRmachine{66}%
\XRnmach{1}}%
\XRvalue{DEC~(X-Windows)}{%
\XRmachine{13}%
\XRmachine{154}%
\XRmachine{172}%
\XRnmach{3}}%
\XRvalue{DEC~(X-Windows~planned)}{%
\XRmachine{144}%
\XRnmach{1}}%
\XRvalue{DEC~VAXstation}{%
\XRmachine{33}%
\XRmachine{34}%
\XRnmach{2}}%
\XRvalue{DEC~VAXstation~3100}{%
\XRmachine{156}%
\XRnmach{1}}%
\XRvalue{DEC~uVAX~GPX}{%
\XRmachine{100}%
\XRmachine{101}%
\XRnmach{2}}%
\XRvalue{DG~System~Monitor~(Hitachi)}{%
\XRmachine{21}%
\XRnmach{1}}%
\XRvalue{Data~Translation~AT2651}{%
\XRmachine{149}%
\XRnmach{1}}%
\XRvalue{Digisolve~Ikon}{%
\XRmachine{84}%
\XRmachine{85}%
\XRmachine{87}%
\XRmachine{88}%
\XRmachine{89}%
\XRnmach{5}}%
\XRvalue{Gould}{%
\XRmachine{63}%
\XRnmach{1}}%
\XRvalue{Gould~85}{%
\XRmachine{40}%
\XRmachine{60}%
\XRnmach{2}}%
\XRvalue{Gould~8500}{%
\XRmachine{128}%
\XRmachine{130}%
\XRmachine{131}%
\XRmachine{132}%
\XRnmach{4}}%
\XRvalue{Gould~9500}{%
\XRmachine{129}%
\XRnmach{1}}%
\XRvalue{Gould~FD5000~(dead)}{%
\XRmachine{91}%
\XRnmach{1}}%
\XRvalue{Gould~IP8400}{%
\XRmachine{106}%
\XRnmach{1}}%
\XRvalue{Gould~IP8500}{%
\XRmachine{137}%
\XRmachine{138}%
\XRnmach{2}}%
\XRvalue{Gould~IP9000}{%
\XRmachine{123}%
\XRnmach{1}}%
\XRvalue{Graphica~I5048}{%
\XRmachine{53}%
\XRnmach{1}}%
\XRvalue{Grinnell~GMR270}{%
\XRmachine{156}%
\XRnmach{1}}%
\XRvalue{Grinnell~GMRxxx}{%
\XRmachine{151}%
\XRnmach{1}}%
\XRvalue{Grinnell~xxxx}{%
\XRmachine{170}%
\XRnmach{1}}%
\XRvalue{IAC~QD512}{%
\XRmachine{149}%
\XRnmach{1}}%
\XRvalue{IBM~color~(8-bit)}{%
\XRmachine{65}%
\XRnmach{1}}%
\XRvalue{IIS}{%
\XRmachine{99}%
\XRnmach{1}}%
\XRvalue{IIS~70}{%
\XRmachine{24}%
\XRmachine{74}%
\XRmachine{75}%
\XRmachine{76}%
\XRmachine{77}%
\XRnmach{5}}%
\XRvalue{IIS~70E}{%
\XRmachine{68}%
\XRmachine{204}%
\XRnmach{2}}%
\XRvalue{IIS~75}{%
\XRmachine{22}%
\XRmachine{46}%
\XRmachine{82}%
\XRmachine{111}%
\XRmachine{113}%
\XRmachine{131}%
\XRmachine{133}%
\XRmachine{155}%
\XRmachine{165}%
\XRmachine{180}%
\XRmachine{202}%
\XRmachine{203}%
\XRnmach{12}}%
\XRvalue{IIS~IVAS}{%
\XRmachine{15}%
\XRmachine{20}%
\XRmachine{50}%
\XRmachine{68}%
\XRmachine{75}%
\XRmachine{124}%
\XRmachine{125}%
\XRmachine{126}%
\XRmachine{133}%
\XRmachine{139}%
\XRmachine{142}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{165}%
\XRmachine{166}%
\XRmachine{167}%
\XRmachine{173}%
\XRmachine{193}%
\XRmachine{195}%
\XRmachine{196}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRmachine{201}%
\XRnmach{24}}%
\XRvalue{Imaging~Technology~V200}{%
\XRmachine{31}%
\XRnmach{1}}%
\XRvalue{Masscomp~GA1000}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{PC-386~(X-Windows)}{%
\XRmachine{122}%
\XRnmach{1}}%
\XRvalue{Peritek~VCH-Q}{%
\XRmachine{192}%
\XRnmach{1}}%
\XRvalue{Peritek~VCK}{%
\XRmachine{157}%
\XRnmach{1}}%
\XRvalue{Peritek~VCX}{%
\XRmachine{191}%
\XRnmach{1}}%
\XRvalue{RCI~TRAPIX~5500}{%
\XRmachine{150}%
\XRnmach{1}}%
\XRvalue{Ramtek~9460*}{%
\XRmachine{49}%
\XRnmach{1}}%
\XRvalue{Raster~Technology~1/80}{%
\XRmachine{81}%
\XRnmach{1}}%
\XRvalue{SGI~4D/20}{%
\XRmachine{19}%
\XRnmach{1}}%
\XRvalue{Sigma~ARGS~7000}{%
\XRmachine{34}%
\XRnmach{1}}%
\XRvalue{Sigma~ARGS~xxxx}{%
\XRmachine{26}%
\XRmachine{35}%
\XRmachine{86}%
\XRmachine{88}%
\XRnmach{4}}%
\XRvalue{Sigmex~7000}{%
\XRmachine{36}%
\XRnmach{1}}%
\XRvalue{Sony~Trinitron}{%
\XRmachine{168}%
\XRmachine{171}%
\XRnmach{2}}%
\XRvalue{Stellar}{%
\XRmachine{152}%
\XRnmach{1}}%
\XRvalue{Sun}{%
\XRmachine{29}%
\XRmachine{105}%
\XRmachine{158}%
\XRmachine{159}%
\XRnmach{4}}%
\XRvalue{Sun-3}{%
\XRmachine{10}%
\XRmachine{16}%
\XRmachine{48}%
\XRmachine{54}%
\XRmachine{109}%
\XRmachine{115}%
\XRmachine{127}%
\XRmachine{140}%
\XRmachine{185}%
\XRmachine{186}%
\XRnmach{10}}%
\XRvalue{Sun-3/260}{%
\XRmachine{141}%
\XRnmach{1}}%
\XRvalue{Sun-3~(X-Windows)}{%
\XRmachine{73}%
\XRnmach{1}}%
\XRvalue{Sun-3~color}{%
\XRmachine{18}%
\XRmachine{69}%
\XRmachine{70}%
\XRmachine{78}%
\XRmachine{112}%
\XRmachine{169}%
\XRmachine{177}%
\XRmachine{184}%
\XRnmach{8}}%
\XRvalue{Sun-3~mono}{%
\XRmachine{69}%
\XRnmach{1}}%
\XRvalue{Sun-4}{%
\XRmachine{5}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{11}%
\XRmachine{12}%
\XRmachine{17}%
\XRmachine{32}%
\XRmachine{39}%
\XRmachine{47}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{61}%
\XRmachine{62}%
\XRmachine{72}%
\XRmachine{87}%
\XRmachine{90}%
\XRmachine{107}%
\XRmachine{117}%
\XRmachine{120}%
\XRmachine{143}%
\XRmachine{145}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{168}%
\XRmachine{182}%
\XRnmach{27}}%
\XRvalue{Sun-4/110}{%
\XRmachine{119}%
\XRnmach{1}}%
\XRvalue{Sun-4/65}{%
\XRmachine{38}%
\XRnmach{1}}%
\XRvalue{Sun-4~(Openlook~(XAS))}{%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRnmach{3}}%
\XRvalue{Sun-4~(Sunview)}{%
\XRmachine{1}%
\XRmachine{83}%
\XRmachine{92}%
\XRmachine{93}%
\XRmachine{94}%
\XRmachine{103}%
\XRmachine{114}%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRmachine{178}%
\XRmachine{190}%
\XRmachine{194}%
\XRnmach{13}}%
\XRvalue{Sun-4~(X-Windows)}{%
\XRmachine{122}%
\XRmachine{148}%
\XRmachine{153}%
\XRmachine{187}%
\XRmachine{194}%
\XRnmach{5}}%
\XRvalue{Sun-4~color}{%
\XRmachine{3}%
\XRmachine{23}%
\XRmachine{25}%
\XRmachine{67}%
\XRmachine{71}%
\XRmachine{118}%
\XRmachine{183}%
\XRmachine{200}%
\XRnmach{8}}%
\XRvalue{TEK~4014}{%
\XRmachine{125}%
\XRnmach{1}}%
\XRvalue{Vibhyor-2001}{%
\XRmachine{42}%
\XRnmach{1}}%
\XRvalue{X-Windows}{%
\XRmachine{133}%
\XRmachine{149}%
\XRnmach{2}}%
\XRvalue{none}{%
\XRmachine{30}%
\XRmachine{51}%
\XRmachine{64}%
\XRmachine{162}%
\XRmachine{176}%
\XRnmach{5}}%
}%

\XRkey{VLA}{%
\XRvalue{0}{%
\XRmachine{2}%
\XRmachine{3}%
\XRmachine{6}%
\XRmachine{7}%
\XRmachine{8}%
\XRmachine{9}%
\XRmachine{10}%
\XRmachine{12}%
\XRmachine{13}%
\XRmachine{25}%
\XRmachine{29}%
\XRmachine{37}%
\XRmachine{41}%
\XRmachine{42}%
\XRmachine{44}%
\XRmachine{47}%
\XRmachine{54}%
\XRmachine{63}%
\XRmachine{66}%
\XRmachine{78}%
\XRmachine{84}%
\XRmachine{102}%
\XRmachine{110}%
\XRmachine{111}%
\XRmachine{126}%
\XRmachine{130}%
\XRmachine{143}%
\XRmachine{144}%
\XRmachine{145}%
\XRmachine{157}%
\XRmachine{165}%
\XRmachine{177}%
\XRmachine{181}%
\XRmachine{186}%
\XRmachine{187}%
\XRmachine{189}%
\XRmachine{202}%
\XRmachine{203}%
\XRmachine{204}%
\XRnmach{39}}%
\XRvalue{1}{%
\XRmachine{176}%
\XRnmach{1}}%
\XRvalue{3}{%
\XRmachine{174}%
\XRnmach{1}}%
\XRvalue{5}{%
\XRmachine{1}%
\XRmachine{148}%
\XRmachine{149}%
\XRnmach{3}}%
\XRvalue{8}{%
\XRmachine{52}%
\XRmachine{53}%
\XRnmach{2}}%
\XRvalue{10}{%
\XRmachine{27}%
\XRmachine{28}%
\XRmachine{30}%
\XRmachine{69}%
\XRmachine{70}%
\XRmachine{73}%
\XRmachine{82}%
\XRmachine{131}%
\XRmachine{188}%
\XRmachine{191}%
\XRnmach{10}}%
\XRvalue{15}{%
\XRmachine{193}%
\XRnmach{1}}%
\XRvalue{20}{%
\XRmachine{5}%
\XRmachine{31}%
\XRmachine{129}%
\XRmachine{142}%
\XRnmach{4}}%
\XRvalue{25}{%
\XRmachine{114}%
\XRmachine{115}%
\XRmachine{123}%
\XRmachine{190}%
\XRnmach{4}}%
\XRvalue{30}{%
\XRmachine{4}%
\XRmachine{106}%
\XRmachine{107}%
\XRmachine{128}%
\XRnmach{4}}%
\XRvalue{33}{%
\XRmachine{116}%
\XRnmach{1}}%
\XRvalue{35}{%
\XRmachine{26}%
\XRnmach{1}}%
\XRvalue{40}{%
\XRmachine{98}%
\XRmachine{99}%
\XRmachine{100}%
\XRmachine{101}%
\XRmachine{132}%
\XRnmach{5}}%
\XRvalue{45}{%
\XRmachine{87}%
\XRmachine{88}%
\XRnmach{2}}%
\XRvalue{50}{%
\XRmachine{23}%
\XRmachine{33}%
\XRmachine{46}%
\XRmachine{51}%
\XRmachine{60}%
\XRmachine{64}%
\XRmachine{72}%
\XRmachine{91}%
\XRmachine{117}%
\XRmachine{133}%
\XRmachine{140}%
\XRmachine{166}%
\XRmachine{192}%
\XRnmach{13}}%
\XRvalue{60}{%
\XRmachine{20}%
\XRmachine{35}%
\XRmachine{39}%
\XRnmach{3}}%
\XRvalue{65}{%
\XRmachine{153}%
\XRnmach{1}}%
\XRvalue{70}{%
\XRmachine{49}%
\XRmachine{50}%
\XRmachine{93}%
\XRmachine{137}%
\XRmachine{138}%
\XRmachine{146}%
\XRmachine{147}%
\XRmachine{168}%
\XRnmach{8}}%
\XRvalue{75}{%
\XRmachine{15}%
\XRmachine{59}%
\XRmachine{113}%
\XRmachine{152}%
\XRmachine{154}%
\XRnmach{5}}%
\XRvalue{80}{%
\XRmachine{38}%
\XRmachine{56}%
\XRmachine{57}%
\XRmachine{58}%
\XRmachine{61}%
\XRmachine{62}%
\XRmachine{65}%
\XRmachine{67}%
\XRmachine{68}%
\XRmachine{83}%
\XRmachine{95}%
\XRmachine{96}%
\XRmachine{97}%
\XRmachine{118}%
\XRmachine{122}%
\XRmachine{127}%
\XRmachine{159}%
\XRmachine{162}%
\XRmachine{163}%
\XRmachine{164}%
\XRmachine{167}%
\XRmachine{171}%
\XRmachine{179}%
\XRnmach{23}}%
\XRvalue{85}{%
\XRmachine{22}%
\XRmachine{92}%
\XRmachine{94}%
\XRnmach{3}}%
\XRvalue{90}{%
\XRmachine{36}%
\XRmachine{40}%
\XRmachine{43}%
\XRmachine{55}%
\XRmachine{71}%
\XRmachine{85}%
\XRmachine{119}%
\XRmachine{150}%
\XRmachine{155}%
\XRmachine{156}%
\XRmachine{180}%
\XRmachine{182}%
\XRmachine{183}%
\XRmachine{184}%
\XRmachine{185}%
\XRmachine{197}%
\XRmachine{198}%
\XRmachine{199}%
\XRnmach{18}}%
\XRvalue{94}{%
\XRmachine{134}%
\XRmachine{135}%
\XRmachine{136}%
\XRnmach{3}}%
\XRvalue{95}{%
\XRmachine{48}%
\XRmachine{74}%
\XRmachine{76}%
\XRmachine{77}%
\XRmachine{103}%
\XRmachine{125}%
\XRmachine{170}%
\XRmachine{200}%
\XRmachine{201}%
\XRnmach{9}}%
\XRvalue{98}{%
\XRmachine{16}%
\XRmachine{18}%
\XRnmach{2}}%
\XRvalue{99}{%
\XRmachine{75}%
\XRmachine{195}%
\XRnmach{2}}%
\XRvalue{100}{%
\XRmachine{11}%
\XRmachine{14}%
\XRmachine{17}%
\XRmachine{19}%
\XRmachine{21}%
\XRmachine{24}%
\XRmachine{32}%
\XRmachine{34}%
\XRmachine{45}%
\XRmachine{79}%
\XRmachine{80}%
\XRmachine{81}%
\XRmachine{86}%
\XRmachine{89}%
\XRmachine{90}%
\XRmachine{104}%
\XRmachine{105}%
\XRmachine{108}%
\XRmachine{109}%
\XRmachine{112}%
\XRmachine{120}%
\XRmachine{121}%
\XRmachine{124}%
\XRmachine{139}%
\XRmachine{141}%
\XRmachine{151}%
\XRmachine{158}%
\XRmachine{160}%
\XRmachine{161}%
\XRmachine{169}%
\XRmachine{172}%
\XRmachine{173}%
\XRmachine{175}%
\XRmachine{178}%
\XRmachine{194}%
\XRmachine{196}%
\XRnmach{36}}%
}%

\endcrossref

\vfill

\eject

\end
