%-----------------------------------------------------------------------
%;  Copyright (C) 1995
%;  Associated Universities, Inc. Washington DC, USA.
%;
%;  This program is free software; you can redistribute it and/or
%;  modify it under the terms of the GNU General Public License as
%;  published by the Free Software Foundation; either version 2 of
%;  the License, or (at your option) any later version.
%;
%;  This program is distributed in the hope that it will be useful,
%;  but WITHOUT ANY WARRANTY; without even the implied warranty of
%;  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%;  GNU General Public License for more details.
%;
%;  You should have received a copy of the GNU General Public
%;  License along with this program; if not, write to the Free
%;  Software Foundation, Inc., 675 Massachusetts Ave, Cambridge,
%;  MA 02139, USA.
%;
%;  Correspondence concerning AIPS should be addressed as follows:
%;          Internet email: aipsmail@nrao.edu.
%;          Postal address: AIPS Project Office
%;                          National Radio Astronomy Observatory
%;                          520 Edgemont Road
%;                          Charlottesville, VA 22903-2475 USA
%-----------------------------------------------------------------------
\input al82.mac
\input al8pt.mac
\letterbegin {IX} {4} {October 15, 1989}

\subtitf{Lead article}

\subtit{Next article}

\subtit{User Agreement}

   Starting with the 15OCT89 release of \AIPS\ all user sites will
need to provide the NRAO with a signed ``User Agreement'' form.  This
``User Agreement'' will be a no-cost item for sites engaged in basic
research in astronomy.  The need for an \AIPS\ ``user agreement'' has
arisen for several reasons.  The most important is that we want all
\AIPS\ sites to obtain their copies of \AIPS\ from the NRAO and
thereby to be made aware of the restrictions that apply to their use
of the code and to our support of it.  For sites doing astronomical
research, these restrictions are only to maintain the proprietary
nature of the code and to direct third parties who wish to receive the
code to the \hbox{NRAO}.  Once properly signed, a User Agreement
remains valid for 5 years and does not need to be renewed before this
time.  A copy of this agreement is printed in the back of this
\AIPS Letter.  The agreement should be signed by an individual in a
position to obligate the user group to follow the agreement.  In most
cases this will be a department chairman or an administrative officer.
Mail signed forms to Amy Shepherd, NRAO, Edgemont Road,
Charlottesville, VA 22903-2475

\subtit{Personnel}
\subtit{Document and Software Distribution by AIPSSERV}

   The NRAO maintains a mail-based file server, AIPSSERV, which is
available for use by \AIPS\ users to fetch files from {\tt CVAX}.
Detailed instructions for using this facility may be obtained by
sending an E-mail message containing the single word ``{\tt help}'' to
one of the following addresses:\hfil\break
{\tt aipsserv@nrao.edu}, {\tt aipsserv@nrao.bitnet}, {\tt
...!uunet!nrao1!aipsserv} or {\tt 6654::aipsserv}.

\subtit{Summary of Changes:  15 July 1989 --- 15 October 1989}
\smallhead{Changes of Interest to Users: 15OCT89}
\smallhead{Changes of Interest to Users: 15JAN90}
\smallhead{Changes of Interest to Programmers: 15OCT89}
\smallhead{Changes of Interest to Programmers: 15JAN90}

\vfill\eject
%\pgskip
\input order.tex
\end
