% Old, historical, AIPS User Agreement.  Now a curiosity.  Here's how
% we do it now:
%-----------------------------------------------------------------------
%;  Copyright (C) 1995
%;  Associated Universities, Inc. Washington DC, USA.
%;
%;  This program is free software; you can redistribute it and/or
%;  modify it under the terms of the GNU General Public License as
%;  published by the Free Software Foundation; either version 2 of
%;  the License, or (at your option) any later version.
%;
%;  This program is distributed in the hope that it will be useful,
%;  but WITHOUT ANY WARRANTY; without even the implied warranty of
%;  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%;  GNU General Public License for more details.
%;
%;  You should have received a copy of the GNU General Public
%;  License along with this program; if not, write to the Free
%;  Software Foundation, Inc., 675 Massachusetts Ave, Cambridge,
%;  MA 02139, USA.
%;
%;  Correspondence concerning AIPS should be addressed as follows:
%;          Internet email: aipsmail@nrao.edu.
%;          Postal address: AIPS Project Office
%;                          National Radio Astronomy Observatory
%;                          520 Edgemont Road
%;                          Charlottesville, VA 22903-2475 USA
%-----------------------------------------------------------------------

\def\bbar{\vbox{\vfill\hrule width 2.5in}}
\centerline{ THE FORMER AIPS USER AGREEMENT}
\centerline{(Now relegated to a historical document)}
\vskip 0.20in
\centerline{ASSOCIATED UNIVERSITIES, INC.}
\vskip 0.1in
\centerline{NATIONAL RADIO ASTRONOMY OBSERVATORY}
\vskip 0.2in
\centerline {BASIC RESEARCH IN ASTRONOMY} \vskip 0.1in

{\bf
  THIS DOCUMENT IS NO LONGER A VALID WAY OF OBTAINING A LICENSE TO RUN
  AIPS.  THE GNU GENERAL PUBLIC LICENSE HAS SUPERCEDED IT.  }
\vskip 0.2in

\vbox{\hbox{\qquad THIS AGREEMENT, made this \vbox{\vfill\hrule width 0.25in}
day of \vbox{\vfill\hrule width 1.25in}, 19\vbox{\vfill\hrule width 0.25in},
between ASSOCIATED}
\vskip 0.1in\hbox{UNIVERSITIES, INC./NATIONAL RADIO ASTRONOMY OBSERVATORY,
Charlottesville, Virginia,}
\vskip 0.1in\hbox{hereinafter called the
Developer; and \vbox{\vfill\hrule width 4.1in}}
\vskip 0.1in\hbox {hereinafter called the User.}}
\vskip 0.1in

      WHEREAS, the Developer has developed an Astronomical Image
Processing System (AIPS) under agreement with the National Science
Foundation; and
\vskip 0.2in

     WHEREAS, the Developer desires to promote basic research,
particularly in astronomy, throughout the world; and
\vskip 0.2in

     WHEREAS, the Developer has determined that distribution of AIPS
to User will further such research.
\vskip 0.2in

NOW, THEREFORE, in consideration of the mutual covenants and
agreements herein contained, the parties agree as follows:
\vskip 0.2in

\item{1.} For no charge, Developer hereby grants to the User a
non-exclusive right to use the Astronomical Image Processing System
(AIPS).\parskip 5pt

\item{2.}  User acknowledges that the AIPS software/system, including
all documentation, will be used primarily by User for basic research.

\item{3.}  The Developer agrees that it will make available to User
updates and modifications to AIPS; however, Developer is under no
obligation to do so.

\item{4.}  The AIPS software/documentation shall not be transfered,
sold, or assigned to a third party.

\item{5.}  AIPS shall not be copied or duplicated except as
follows:

\itemitem{a.} AIPS may be copied for backup purposes.

\itemitem{b.} AIPS may be duplicated and run on all systems supported
by the User's AIPS Manager, who will be the designated operational
contact person between the User and Developer, provided that its use
on all such systems is in accordance with the dictates of this agreement.

\item{6.}  The obligation to use, install, and run AIPS shall be the
sole responsibility of the User.  It IS UNDERSTOOD THAT THE DEVELOPER
MAKES NO WARRANTIES, CLAIMS, OR OTHER ASSERTIONS, IMPLIED OR
OTHERWISE, CONCERNING THE PERFORMANCE OR INTEGRITY OF AIPS OR THE
SUITABILITY OF AIPS FOR A PARTICULAR PURPOSE.


\item{7.}  User acknowledges that any data, tapes, or containers
received from the Developer are and shall remain Developer's property
and shall be returned upon request of the Developer, when updated
materials are requested by User or at termination of this agreement.


\item{8.}  This Agreement shall remain in effect for a period of five
(5) years from the date hereof.\parskip 0pt
\vskip 0.75in

\vfil\eject

\noindent {YOUR INSTITUTION:} \hskip 1.72in {AUI/NRAO:} \par
\vskip 0.2in
\hskip 2.94in {ASSOCIATED UNIVERSITIES, INC./}
\par \noindent \vskip 2pt \hskip 2.94in {NATIONAL RADIO ASTRONOMY OBSERVATORY\hfill}
\par \noindent \vskip 2pt \hskip 2.94in {520 EDGEMONT ROAD\hfill}
\par \noindent \vskip 2pt \hskip 2.94in {CHARLOTTESVILLE, VA 22903-2475\hfill}
\par \vskip 0.1in
\noindent\bbar\hfil\bbar \par
\noindent{Institution  Name and Address}
\par \vskip 0.5in
\noindent\bbar\hfil\bbar\par\noindent
\noindent{Signature}\hskip 2.66in {Signature}\par\noindent
\vskip 0.5in
\noindent\bbar\hfil\bbar\par\noindent
\noindent{Print Name}\hskip 2.52in {Print Name} \par\noindent
\vskip 0.5in
\noindent\bbar\hfil\bbar\par\noindent
\noindent{Title}\hskip 2.95in{Title}\par\noindent
\vskip 0.5in
\noindent\bbar\hfil\bbar\par\noindent
\noindent {Date} \hskip 2.9in {Date} \par
\vskip 1.5in
\vfill
\noindent{\it Please indicate technical contact person (i.e. AIPS
Manager):}\par\noindent
\vskip 0.35in
\bbar
\vskip 1.2in

\rightline{Rev. 06/02/94}
\eject
\end


